---
title: "NB_EnvData_CollinearTests"
author: "Amy Zyck"
date: '2024-05-02'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Before moving on to outlier detection programs for the EecSeq SNP dataset, I'm going to test the environmental variables for collinearity. I have a large dataset with 40 environmental variables. I'd like to explore the data and identify variables that are collinear as these can affect downstream analysis. 

Dormann et al. 2013 is a nice review of methods for identifying and working with collinear predictor variables. I'd like to work through a few of these approaches to see which works best for my data. 

```{r}
install.packages("psych")
install.packages("mvnormtest")
install.packages("ggfortify")
```

```{r}
library(psych)    # Used to investigate correlations among predictors
library(vegan)    # Used to run RDA
# library(packfor)  # Used for forward selection to ID significant predictor variables 
library(adegenet) # Used for DAPC 
library(dplyr)
library(ggplot2)
library(ggfortify)
#library(factoextra)
```

```{r}
# Subsetting the strata file to just include the environmental variables
# Also shortening the column names
# Transposing dataframe under new name - need this for running and plotting PCA 
strata6_full <- read.csv("strata6.csv", header=TRUE)
env_data_pop <- strata6_full %>% distinct(Population, .keep_all = TRUE)
env_data_pop <- env_data_pop[,5:44]
colnames(env_data_pop) <- c("SE","T_Avg","T_Min","T_Max","Sal_Avg","Sal_Min","Sal_Max","pH_Avg","DO_Avg",	"May_T_Avg","May_T_Min","May_T_Max","May_Sal_Avg","May_Sal_Min","May_Sal_Max","May_pH_Avg","May_DO_Avg",	"Jun_T_Avg","Jun_T_Min","Jun_T_Max","Jun_Sal_Avg","Jun_Sal_Min","Jun_Sal_Max","Jun_pH_Avg","Jun_DO_Avg","Jul_T_Avg",	"Jul_T_Min","Jul_T_Max","Jul_Sal_Avg","Jul_Sal_Min","Jul_Sal_Max","Jul_pH_Avg","Jul_DO_Avg","Aug_T_Avg","Aug_T_Min",	"Aug_T_Max","Aug_pH_Avg","Sept_T_Avg","Sept_T_Min","Sept_T_Max")
env_data_bypop <- t(env_data_pop)
```

```{r}
# removing se from data frame because it is an outlier 
env_data_pop_nose <- env_data_pop[,2:40]
```

```{r}
# standardizing dataset
std_env <- scale(env_data_pop)
std_env_t <- t(std_env)
```

```{r}
# Pearson Correlation coefficient 
env_corr <- abs(cor(env_data_pop))

env_corr_noabs <- cor(env_data_pop)

env_corr_noabs_nose <- cor(env_data_pop[,-1])

env_corr_nose <- abs(cor(env_data_pop[,-1]))

# Converting to a .csv file so I can explore in Excel 
write.table(env_corr, "~/NB_capture_both/NB_ddhaplo/NB_OutlierDetect_both/NB_OutlierDetect_6pops/env_corr_6popsfull.csv", sep="\t", quote=F, row.names=T)
```

```{r}
# We can visually look for correlations between variables:
heatmap(abs(cor(strata6_full[,5:44])), 
        # Compute pearson correlation (note they are absolute values)
        col = rev(heat.colors(6)), symm = TRUE, margins = c(6,6), scale = "none",  
        Colv = NA, Rowv = NA)
legend("topright", 
       title = "Absolute Pearson R",
       legend =  round(seq(0,1, length.out = 6),1),
       y.intersp = 0.7, bty = "n",
       fill = rev(heat.colors(6)))
```

# Collinearity Method 1 

```{r} 
# Running PCA on correlation matrix of environmental variables (excluding SE as an outlier)
env_pca_corr <- prcomp(env_corr_noabs_nose)
```

```{r}
summary(env_pca_corr)
```

```{r}
# Saving loadings from PCA as dataframe 
corr_pca_loadings <- as.data.frame(env_pca_corr$rotation)
```

```{r}
# Adding populations as rownames to data set for PCA 
row.names(env_data_pop) <- c("BAR", "BIS", "GB", "KIC", "MCD", "PVD")
```

```{r}
# Running PCA on the env dataset (with SE excluded)
env_pca_col <- prcomp(env_data_pop[,2:40], center = TRUE, scale = TRUE, retx = TRUE)
```

```{r}
# summary(env_pca)
summary(env_pca_col)
```

The first 3 PCs explain a majority of the variance in the environmental data

Based on the loadings from the PCA on the correlation matrix and using the PCA collinearity method from Dormann et al. (loadings >0.32 in each PC are likely collinear), three clusters of variables in PC2, 4 and 5 were identified. I only focused on the first 5 PCs as these explain 100% of the variation. This aligns with the output from the PCA run directly on the 6 observations per variable (variables as columns). This second PCA also identified a cluster in PC3. Since I have two separate PCA analyses from both the correlation matrix (recommended by Dormann et al.) and the direct variables (centered and scaled) identifying the same clusters, I am going to take these clusters and identify the variable closest to the cluster centroid. This variable will be used to represent the cluster moving forward.

Determining variable closest to centroid of each cluster  

```{r}
# PC2 cluster 

# Extract the loadings for the principal component (PC) that the two variables cluster on:
pc2_loadings <- env_pca_col$rotation[, 2]

# Calculate the centroid of the cluster:
cluster_centroid_pc2 <- mean(pc2_loadings[c("T_Avg", "May_T_Avg")])

# Calculate the distance between each variable and the centroid:
distance_TAvg <- abs(pc2_loadings["T_Avg"] - cluster_centroid_pc2)
distance_MayTAvg <- abs(pc2_loadings["May_T_Avg"] - cluster_centroid_pc2)

if (distance_TAvg < distance_MayTAvg) {
  print("Variable 1 is closer to the centroid")
} else {
  print("Variable 2 is closer to the centroid")
}
```

The difference between T_Avg and centroid is 0.006971257 whereas the difference between May_T_Avg and centroid is 0.006971312, so T_Avg will represent this cluster. 

```{r}
# PC3 cluster 

# Extract the loadings for the principal component (PC) that the two variables cluster on:
pc3_loadings <- env_pca_col$rotation[, 3]

# Calculate the centroid of the cluster:
cluster_centroid_pc3 <- mean(pc3_loadings[c("pH_Avg", "May_pH_Avg")])

# Calculate the distance between each variable and the centroid:
distance_pHAvg <- abs(pc3_loadings["pH_Avg"] - cluster_centroid_pc3)
distance_MaypHAvg <- abs(pc3_loadings["May_pH_Avg"] - cluster_centroid_pc3)

if (distance_pHAvg < distance_MaypHAvg) {
  print("Variable 1 is closer to the centroid")
} else {
  print("Variable 2 is closer to the centroid")
}
```

The difference between pH_Avg and centroid is 0.011648774 whereas the difference between May_pH_Avg and centroid is 0.01164869, so May_pH_Avg will represent this cluster.

```{r}
# PC4 cluster 

# Extract the loadings for the principal component (PC) that the two variables cluster on:
pc4_loadings <- env_pca_col$rotation[, 4]

# Calculate the centroid of the cluster:
cluster_centroid_pc4 <- mean(pc4_loadings[c("May_Sal_Max", "Jul_DO_Avg", "Sept_T_Avg")])

# Calculate the distance between each variable and the centroid:
distances <- c(
  abs(pc4_loadings["May_Sal_Max"] - cluster_centroid_pc4),
  abs(pc4_loadings["Jul_DO_Avg"] - cluster_centroid_pc4),
  abs(pc4_loadings["Sept_T_Avg"] - cluster_centroid_pc4)
)

# Find the index of the variable with the minimum distance
closest_var_index <- which.min(distances)

# Create a vector of variable names
var_names <- c("Variable 1", "Variable 2", "Variable 3")

# Print the result
cat(var_names[closest_var_index], "is closest to the centroid\n")

# If you want to see the distances for all variables:
for (i in 1:3) {
  cat(var_names[i], "distance:", distances[i], "\n")
}
```

Sept_T_Avg will represent this cluster

```{r}
# PC5 cluster 

# Extract the loadings for the principal component (PC) that the two variables cluster on:
pc5_loadings <- env_pca_col$rotation[, 5]

# Calculate the centroid of the cluster:
cluster_centroid_pc5 <- mean(pc5_loadings[c("Jul_T_Min", "Aug_T_Min")])

# Calculate the distance between each variable and the centroid:
distance_JulTMin <- abs(pc5_loadings["Jul_T_Min"] - cluster_centroid_pc5)
distance_AugTMin <- abs(pc5_loadings["Aug_T_Min"] - cluster_centroid_pc5)

if (distance_JulTMin < distance_AugTMin) {
  print("Variable 1 is closer to the centroid")
} else {
  print("Variable 2 is closer to the centroid")
}
```

Aug_T_Min will represent this cluster. 

```{r}
# Prepping the environmental data

# Excluding variables from clusters represented by other variables from PCA collinearity analysis
drop <- c("May_Temp_Avg", "pH_Avg","May_Salinity_Max", "July_DO_Avg","July_Temp_Min")

strata6_pca_coll_method <- strata6_full[,!(names(strata6_full) %in% drop)]

strata6_pca_coll_method

write.table(strata6_pca_coll_method, "/home/azyck/NB_capture_both/NB_ddhaplo/NB_ddhaplo_working/NB_OutlierDetection_working/strata6_pcacoll_35var", sep="\t", quote=F, row.names=F)
```


# Collinearity Method 2 

```{r}
# Running PCA on variables (rows are variables columns are observations)
env_pca_row <- prcomp(env_data_bypop[2:40,], scale = TRUE, retx = TRUE)
```

```{r}
summary(env_pca_row)
```

```{r}
# Plotting PCA to examine clustering of variables
autoplot(env_pca_row, data = env_data_bypop[2:40,], shape = FALSE, label = TRUE)
```

```{r}
summary(env_pca_row)
```

```{r}
# Plotting PCA with variables as columns
autoplot(env_pca_col, data = env_data_pop[,2:40], shape = FALSE, label = TRUE)
```

```{r}
summary(env_pca_col)
```

It looks like the first 3 PCs explain a majority of the variance. I will save these and include them as new "environmental variables". 

```{r}
# Save scores for PCs 1-3 to new variable
pc_scores_env <- env_pca_col$x
pc_scores_3 <- pc_scores_env[,1:3]
pc_scores_3 <- as.data.frame(pc_scores_3)
```

```{r}
# Plotting env_pca_col results with variables as loadings 
autoplot(env_pca_col, data = env_data_pop[,2:40], shape = FALSE, label = TRUE, loadings = TRUE, loadings.label = TRUE, loadings.label.size = 3)
```

```{r}
# Compute k-means with k = 8
set.seed(123)
env.km <- kmeans(scale(env_data_bypop[2:40,]), 8)
# K-means clusters showing the group of each individuals
env.km$cluster
```

```{r}
# Plot k-means results 
cluster_data_all <- data.frame(PC1 = pc_scores_envall[,1], PC2 = pc_scores_envall[,2], Cluster = as.factor(env.km$cluster))
ggplot(cluster_data_all, aes(x = PC1, y = PC2, color = Cluster)) +
  geom_point() +
  geom_text(aes(label = rownames(env_data_bypop[2:40,])))+
  theme_minimal() +
  labs(title = "K-means Clustering of All Env Variables")
```

Within these 8 clusters, I'm going to identify the variable closest to the centroid. This will represent all variables of the cluster. 

```{r}
find_closest_to_cluster_center <- function(pca_result, kmeans_result) {
  # Extract the principal component scores
  pc_scores <- pca_result$x
  
  # Get the number of clusters
  n_clusters <- length(unique(kmeans_result$cluster))
  
  # Initialize a list to store results
  closest_variables <- list()
  
  for (i in 1:n_clusters) {
    # Get the indices of variables in this cluster
    cluster_indices <- which(kmeans_result$cluster == i)
    
    # Calculate the centroid of the cluster based on PC scores
    cluster_center <- colMeans(pc_scores[cluster_indices, , drop = FALSE])
    
    # Calculate distances from each variable in the cluster to the cluster center
    distances <- apply(pc_scores[cluster_indices, , drop = FALSE], 1, function(x) sum((x - cluster_center)^2))
    
    # Find the variable with the minimum distance
    closest_var_index <- cluster_indices[which.min(distances)]
    closest_var_name <- rownames(pc_scores)[closest_var_index]
    
    # Store the result
    closest_variables[[paste0("Cluster_", i)]] <- closest_var_name
  }
  
  return(closest_variables)
}
```

```{r}
# Run the function 
closest_vars <- find_closest_to_cluster_center(env_pca_row, env.km)

# Print the results
print(closest_vars)
```


Performing PCA on the pH and DO variables 

```{r}
# Specifying the columns to keep 
keep_DOpH <- c("pH_Avg","DO_Avg","May_pH_Avg","May_DO_Avg","Jun_pH_Avg","Jun_DO_Avg","Jul_pH_Avg","Jul_DO_Avg","Aug_pH_Avg")
```

```{r}
# Specifying the columns to keep 
keep_pH <- c("pH_Avg","May_pH_Avg","Jun_pH_Avg","Jul_pH_Avg","Aug_pH_Avg")
```

```{r}
pH_DO <- env_data_pop[keep_DOpH]
pH_DO_t <- t(pH_DO)
```

```{r}
pH <- env_data_pop[keep_pH]
pH_t <- t(pH)
```

```{r}
# Running PCA on transposed pH and DO dataset
pHDO_pca_t <- prcomp(pH_DO_t, scale = TRUE)
```

```{r}
# Running PCA on transposed pH and DO dataset
pH_pca_t <- prcomp(pH_t, scale = TRUE)
```

```{r}
# Running PCA on pH and DO dataset
pHDO_pca <- prcomp(pH_DO, scale = TRUE)
```

```{r}
pHDO_loadings <- pHDO_pca$rotation
```

```{r}
# Plotting PCA 
autoplot(pHDO_pca_t, data = pH_DO_t, shape = FALSE, label = TRUE)
```

```{r}
# Plotting PCA 
autoplot(pH_pca_t, data = pH_t, shape = FALSE, label = TRUE)
```

```{r}
# Compute k-means with k = 8
set.seed(123)
pHDO.km <- kmeans(scale(pH_DO_t), 4)
# K-means clusters showing the group of each individuals
pHDO.km$cluster
```

```{r}
# Run the function 
closest_vars_pHDO <- find_closest_to_cluster_center(pHDO_pca_t, pHDO.km)

# Print the results
print(closest_vars_pHDO)
```


Performing PCA on the Temperature variables

```{r}
# Specifying the columns to keep 
keep_T <- c("T_Avg","T_Min","T_Max","May_T_Avg","May_T_Min","May_T_Max","Jun_T_Avg","Jun_T_Min","Jun_T_Max","Jul_T_Avg","Jul_T_Min","Jul_T_Max","Aug_T_Avg","Aug_T_Min","Aug_T_Max","Sept_T_Avg","Sept_T_Min","Sept_T_Max")
```

```{r}
T_env <- env_data_pop[keep_T]
T_env_t <- t(T_env)
```

```{r}
# Running PCA on Temp dataset
T_pca <- prcomp(T_env, scale = TRUE)
```

```{r}
# Running PCA on transposed Temp dataset
T_pca_t <- prcomp(T_env_t, scale = TRUE)
```

```{r}
temp_loadings <- T_pca$rotation
```

```{r}
# Plotting PCA 
autoplot(T_pca, data = T_env, shape = FALSE, label = TRUE)
```

```{r}
# Plotting PCA of transposed data 
autoplot(T_pca_t, data = T_env_t, shape = FALSE, label = TRUE)
```

```{r}
# Compute k-means with k = 8
set.seed(123)
Temp.km <- kmeans(scale(T_env_t), 3)
# K-means clusters showing the group of each individuals
Temp.km$cluster
```

```{r}
library(ggplot2)
cluster_data_temp <- data.frame(PC1 = pc_scores_temp[,1], PC2 = pc_scores_temp[,2], Cluster = as.factor(Temp.km$cluster))
ggplot(cluster_data_temp, aes(x = PC1, y = PC2, color = Cluster)) +
  geom_point() +
  geom_text(aes(label = rownames(T_env_t)))+
  theme_minimal() +
  labs(title = "K-means Clustering of Temperature Variables")
```


Based on PCA analysis and kmeans clustering of the environmental variables, I'm going to create a reduced environmental dataset that includes Sewage Effluent (not collinear), T_Min, Jun_Sal_Max, T_Max, Jul_T_Min, Jul_T_Avg, Jun_Sal_Avg, Jun_Sal_Min,  pH_Avg, Jun_DO_Avg, and scores for PCs 1-3. 

```{r}
# Saving these variables to a keep variable
env_data_pca_9var_tokeep <- c("Individual", "Population", "Latitude", "Longitude", "SewageEffluent","May_Temp_Min","June_Salinity_Max","Temp_Max", "July_Temp_Min", "July_Temp_Avg","June_Salinity_Avg","June_Salinity_Min","pH_Avg")
```

```{r}
# saving env dataset with just variables from keep list 
env_data_9var <- strata6_full[env_data_pca_9var_tokeep]
```

The strata file includes all individuals per population. Right now, the PC scores are saved as one value per population, so I need to duplicate each population's PC score 10 times (10 individuals per pop) before I can bind it with the environmental strata info.  

```{r}
env_pc_scores_strata <- pc_scores_3 %>%
  slice(rep(1:n(), each = 10)) %>%
  mutate(duplicate_id = rep(1:10, times = nrow(pc_scores_3))
```

```{r}
pc_scores_strata <- env_pc_scores_strata[,1:3]
rownames(pc_scores_strata) <- NULL
```

```{r}
# Merging the reduced env dataset with the pc scores
env_data_12var <- cbind(env_data_9var, pc_scores_strata)
```

```{r}
write.table(env_data_12var, "/home/azyck/NB_capture_both/NB_ddhaplo/NB_ddhaplo_working/NB_PopSeaGen_Analysis_working/strata6_red_12var", sep="\t", quote=F, row.names=F)
```

