---
title: "NB Population Genomic Analysis on Non-Haplotig Masked All SNPs"
author: "Amy Zyck"
date: "7/12/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Population analyses on All SNP dataset 

I first created a new directory `NB_PopGen_All` within `PATH: /home/azyck/NB_capture/NB_ddocent`. I linked `SNP.TRSdp5g5mafMIp9g9dnDNAmaf052A4.recode.vcf` to this working directory. The RMarkdown file should be saved in this same working directory. 

```{r}
#Loading all the necessary packages
library(adegenet)
library(vcfR)
library("radiator") # Conversion from vcf to a lot of other formats
library("dplyr")
library("hierfstat")
library("ggplot2") #For plotting
library("reshape2") #For plotting
library("plyr")
library("cowplot") #For plotting manuscript figs
library(PCAviz)  #Visualizing output of PCA
library("stringr")
library("bigsnpr") # package for Linkage Disequilibrium pruning
```

# All SNPs

## Making files

**Make genind object**

`SNP.TRSdp5g5mafMIp9g9dnDNAmaf052A4.recode.vcf` contains all SNPs (both outlier and neutral SNPs) for populations PVD, GB, BIS, and NAR. Steps for generating this VCF file are located in [EecSeq_Cvirginica_dDocent.md](https://github.com/amyzyck/EecSeq_NB_EasternOyster/blob/master/Analysis/EecSeq_Cvirginica_dDocent.md) and [EecSeq_Cvirginica_Filtering.md](https://github.com/amyzyck/EecSeq_NB_EasternOyster/blob/master/Analysis/EecSeq_Cvirginica_Filtering.md). Population NIN was removed from the VCF file following steps in [EecSeq_Cvirginica_OutlierDetection.md](https://github.com/amyzyck/EecSeq_NB_EasternOyster/blob/master/Analysis/EecSeq_Cvirginica_OutlierDetection.md).

`strata` contains population, environmental, and library information for each sample - can be accessed [here](https://github.com/amyzyck/EecSeq_NB_EasternOyster/blob/master/Analysis/PopGen_SeaGen_Analyses/strata).

```{r}
my_vcf <- read.vcfR("SNP.TRSdp5g5mafMIp9g9dnDNAmaf052A4.recode.vcf")
strata <- read.table("strata", header=TRUE)
```

```{r}
rad.filt <- vcfR2genind(my_vcf, strata = strata, pop = c(rep("BIS", 10),rep("GB", 10),rep("NAR", 10), rep("PVD",10)))
```

```{r}
rad.filt
```

```{r}
#Providing population names for plotting
pop_order <- c("BIS","GB","NAR","PVD")
```

**Read in the other info from .strata file and extract information such as locality, latitude, and longitude.**
```{r}
info <- as.data.frame(read.table("strata",header = T,sep = "\t",stringsAsFactors = F))
mystrats <- as.data.frame(matrix(nrow = length(indNames(rad.filt)),ncol=10))
colnames(mystrats) <- c("Population","Latitude","Longitude","Distance","SE","Temperature","Salinity","pH","Chlorophylla","DO")
just.strats <- select(info,c("Population"))
stratted.filt <- strata(rad.filt, formula= Population, combine = TRUE,just.strats)
stratted.filt@other <- select(info, Latitude,Longitude,Distance,SE,Temperature,Salinity,pH,Chlorophylla,DO)
```

```{r}
stratted.filt
```

**Repeat for quasi-independent set of SNPs**

Steps for filtering are documented for OutFLANK in [EecSeq_Cvirginica_OutlierDetection.md](https://github.com/amyzyck/EecSeq_NB_EasternOyster/blob/master/Analysis/EecSeq_Cvirginica_OutlierDetection.md). However, I am repeating the steps here. 

```{r}
my_vcf <- read.vcfR("SNP.TRSdp5g5mafMIp9g9dnDNAmaf052A4.recode.vcf")

geno <- extract.gt(my_vcf) # Character matrix containing the genotypes
position <- getPOS(my_vcf) # Positions in bp
chromosome <- getCHROM(my_vcf) # Chromosome information

G <- matrix(NA, nrow = nrow(geno), ncol = ncol(geno))

G[geno %in% c("0/0", "0|0")] <- 0
G[geno  %in% c("0/1", "1/0", "1|0", "0|1")] <- 1
G[geno %in% c("1/1", "1|1")] <- 2

# NA should be replaced with “9” to work with the functions in the OutFLANK package
G[is.na(G)] <- 9
```

Chromosomes need to be of class integer for this to work.
```{r}
# Visualizing chromosomes
chrom_unique <- unique(chromosome)
print(chrom_unique)
```

```{r}
# Removing "NC_" from the chromosome name so it can be converted to an integer
chrom_new <- chromosome%>%str_replace("NC_","")

# Converting the character string into an integer
chrom1 <- as.integer(chrom_new)
```

`chromosome` and `position` need to be sorted for imputation to work.

```{r}
chrom_sort <- sort(chrom1)
pos_sort <- sort(position)
```

_Note: This filtering program does not allow for missing genotype values._

I can either remove the missing genotype values or impute missing genotypes. I'm going to try both and see how the two methods differ during the trimming step. 

### Remove missing genotypes

```{r}
# removing missing data 
G_miss <- matrix(NA, nrow = nrow(geno), ncol = ncol(geno))

G_miss[geno %in% c("0/0", "0|0")] <- 0
G_miss[geno  %in% c("0/1", "1/0", "1|0", "0|1")] <- 1
G_miss[geno %in% c("1/1", "1|1")] <- 2

# removing missing data
G_miss[na.omit(G_miss)]
```

```{r}
# Converting the genotype matrix to class FBM.code256 using the matrix where missing data was removed
G1_miss <- add_code256(big_copy(t(G_miss),type="raw"),code=bigsnpr:::CODE_012)
```

```{r}
newpc_miss <- snp_autoSVD(G1_miss,infos.chr =chrom_sort,infos.pos = pos_sort)
which_pruned_miss <- attr(newpc_miss, which="subset") # Indexes of remaining SNPS after pruning
length(which_pruned_miss)
```

### Using `snp_fastImputeSimple` to impute missing genotypes

This requires the genotpe matrix to be converted to class FBM.code256.

> A reference class for storing and accessing up to 256 arbitrary different values using a Filebacked Big Matrix of type unsigned char.

```{r}
# Converting the genotype matrix to class FBM.code256
G1 <- add_code256(big_copy(t(G),type="raw"),code=bigsnpr:::CODE_012)
```

The chunk below imputs missing genotypes in order for the following code to run. Fast imputation via mode, mean, sampling according to allele frequencies, or 0.

Method argument: Either "random" (sampling according to allele frequencies), "mean0" (rounded mean), "mean2" (rounded mean to 2 decimal places), "mode" (most frequent call).

Depending on the method I choose, the # of SNPs kept after running the following chunk changes, so I'm not sure which method to use. 

```{r}
# Imputing missing genotypes
G_missSimple <- snp_fastImputeSimple(G1,method = "mode") # Using mode or no method call results in the same number of SNPs in the following code 
```

Truncated SVD while limiting LD.

```{r}
newpc<-snp_autoSVD(G_missSimple,infos.chr =chrom_sort,infos.pos = pos_sort)
which_pruned <- attr(newpc, which="subset") # Indexes of remaining SNPS after pruning
length(which_pruned)
```

Either removing missing data or imputation results in a similar number of trimmed SNPs, so I will continue forward with missing data removed.

```{r}
invisible(lapply(which_pruned_miss, write, "pruned_data.txt", append=TRUE))
```

**In terminal**

```
$ mawk '!/#/' SNP.TRSdp5g5mafMIp9g9dnDNAmaf052A4.recode.vcf | cut -f1,2 > totalloci
$ NUM=(`cat totalloci | wc -l`)
$ paste <(seq 1 $NUM) totalloci > loci.plus.index
$ cat pruned_data.txt | parallel "grep -w ^{} loci.plus.index" | cut -f2,3> pruned_data.loci.txt

$ head pruned_data.loci.txt

output:
NC_035780.1	573785
NC_035780.1	573788
NC_035780.1	574229
NC_035780.1	574307
NC_035780.1	574313
NC_035780.1	574321
NC_035780.1	574444
NC_035780.1	574580
NC_035780.1	574616
NC_035780.1	574657
```

Create VCF file with just the pruned_data loci 
```
$ vcftools --vcf SNP.TRSdp5g5mafMIp9g9dnDNAmaf052A4.recode.vcf --recode --recode-INFO-all --positions pruned_data.loci.txt --out pruned_data

output:
fter filtering, kept 40 out of 40 Individuals
Outputting VCF file...
After filtering, kept 14969 out of a possible 74520 Sites
Run Time = 3.00 seconds
```

```{r}
my_vcf_u <- read.vcfR("pruned_data.recode.vcf")
strata<- read.table("strata", header=TRUE)
```

```{r}
rad.u <- vcfR2genind(my_vcf_u, strata = strata, pop = c(rep("BIS", 10),rep("GB", 10),rep("NAR", 10), rep("PVD",10)))
```

```{r}
rad.u
```


```{r}
stratted.u <- strata(rad.u, formula= Population, combine = TRUE,just.strats)
stratted.u@other <- select(info, Latitude,Longitude, Distance,SE, Temperature,Salinity,pH,Chlorophylla,DO)
```

```{r}
stratted.u
```

**Make hierfstat object**
```{r}
hf.filt <- genind2hierfstat(rad.filt, pop = c(rep("BIS", 10),rep("GB", 10),rep("NAR", 10), rep("PVD",10))) 
```

```{r}
hf.u <- genind2hierfstat(rad.u, pop = c(rep("BIS", 10),rep("GB", 10),rep("NAR", 10), rep("PVD",10)))
hf.u <- hf.u$hierfstat.no.imputation
```


## Estimating effective migration surfaces [(EEMS)](https://github.com/dipetkov/eems)

The program `runeems_snps` implements the EEMS method for analyzing spatial population structure. This version uses the pairwise genetic dissimilarity matrix computed from SNP data.

Here is the code used to run `runeems_snps` in both RStudio and command-line. Detailed steps for running this program can be accessed [here](https://github.com/dipetkov/eems/tree/master/runeems_snps). Input files are created in RStudio or manually (see below) and saved to the directory where `runeems_snps` will be run. The program is then run in the command-line. The outputs are then plotted in RStudio. 

`runeems_snps` requires three data input files that have the same file name but different extension. The description below assumes that datapath is the full path + the file name (but without the extension).

__1. datapath.diffs__

`datapath.diffs` is the matrix of average pairwise genetic dissimilarities. This can be computed with bed2diffs from genetic data in plink binary format.

The dissimilarity matrix is nonnegative, symmetric, with 0s on the main diagonal. These conditions are necessary but not sufficient for diffs to be a valid dissimilarity matrix. Mathematically, diffs should be conditionally negative definite.

**Steps for generating `datapath.diffs` are completed in RStudio.**
```{r}
# V1 methtod to get diffs matrix, preferred
bed2diffs_v1 <- function(Geno) {
  nIndiv <- nrow(Geno)
  nSites <- ncol(Geno)
  Diffs <- matrix(0, nIndiv, nIndiv)
  
  for (i in seq(nIndiv - 1)) {
    for (j in seq(i + 1, nIndiv)) {
      x <- Geno[i, ]
      y <- Geno[j, ]
      Diffs[i, j] <- mean((x - y)^2, na.rm = TRUE)
      Diffs[j, i] <- Diffs[i, j]
    }
  }
  Diffs
}
```

```{r}
# V2 method to get .diffs matrix, only if V1 doesn't work
bed2diffs_v2 <- function(Geno) {
  nIndiv <- nrow(Geno)
  nSites <- ncol(Geno)
  Miss <- is.na(Geno)
  ## Impute NAs with the column means (= twice the allele frequencies)
  Mean <- matrix(colMeans(Geno, na.rm = TRUE), ## a row of means
                 nrow = nIndiv, ncol = nSites, byrow = TRUE) ## a matrix with nIndiv identical rows of means
  Mean[Miss == 0] <- 0 ## Set the means that correspond to observed genotypes to 0
  Geno[Miss == 1] <- 0 ## Set the missing genotypes to 0 (used to be NA) 
  Geno <- Geno + Mean
  ## Compute similarities
  Sim <- Geno %*% t(Geno) / nSites
  SelfSim <- diag(Sim) ## self-similarities
  vector1s <- rep(1, nIndiv) ## vector of 1s
  ## This chunk generates a `diffs` matrix
  Diffs <- SelfSim %*% t(vector1s) + vector1s %*% t(SelfSim) - 2 * Sim
  Diffs
}
```

```{r}
geno <- stratted.filt@tab
```

```{r}
# Get rid of non-biallelic loci
multi.loci <- names(which(stratted.filt@loc.n.all != 2))
multi.cols <- which(grepl(paste0("^", multi.loci, "\\.\\d+$", collapse = "|"), colnames(geno)))
if (length(multi.cols)) geno <- geno[, - multi.cols]
nloci <- dim(geno)[2] / 2
dim(geno)
```

```{r}
stopifnot(identical(stratted.filt@type, 'codom'))
```

```{r}
# bed2diffs functions  
diffs.v1 <- bed2diffs_v1(geno)
diffs.v2 <- bed2diffs_v2(geno)
# Round to 6 digits
diffs.v1 <- round(diffs.v1, digits = 6)
diffs.v2 <- round(diffs.v2, digits = 6)
```

**Check that the dissimilarity matrix has one positive eigenvalue and nIndiv-1 negative eigenvalues, as required by a full-rank Euclidean distance matrix. If the V1 method does not make a Euclidean matrix, you must use V2.**

```{r}
tail(sort(round(eigen(diffs.v1)$values, digits = 2)))
```

```{r}
tail(sort(round(eigen(diffs.v2)$values, digits = 2)))
```

```{r}
# Set suffix for EEMS input files
suf <- "neutraldata-filt"
```

```{r}
# This saves the file to directory 
write.table(diffs.v1, paste(suf,".v1.diffs",sep=""), 
            col.names = FALSE, row.names = FALSE, quote = FALSE)
```


__2. datapath.coord__

`datapath.coord` are the sample coordinates, two coordinates per sample, one sample per line. The sampling locations should be given in the same order as the rows and columns of the dissimilarity matrix.

**Steps for generating `datapath.coord` are completed in RStudio.**
```{r}
## Get gps coordinates from previously created info matrix
xOR.info <- dplyr::filter(info)
gps_matrix <- select(xOR.info,c("Longitude","Latitude"))
```

```{r}
#write .coord file
write.table(gps_matrix, paste(suf,".v1.coord",sep=""),col.names = FALSE, row.names = FALSE,quote = FALSE)
```

__3. datapath.outer__

`datapath.outer` are the habitat coordinates, as a sequence of vertices that form a closed polygon. The habitat vertices should be listed counterclockwise and the first vertex should also be the last vertex, so that the outline is a closed ring. Otherwise, EEMS attempts to "correct" the polygon and prints a warning message.

**`datapath.outer` is created manually in Excel,based on site coordinates gathered from Google Maps, copied into terminal using `nano`, and saved as the file with the appropriate extension.** 


**`runeems_snps` is then run in command-line following the steps documented in [NB_EEMS_OutlierHap.md](https://github.com/amyzyck/EecSeq_NB_EasternOyster/blob/master/Analysis/PopGen_SeaGen_Analyses/EEMS/NB_EEMS_OutlierHap.md) 

**Back in RStudio to plot `runeems_snps` outputs**

```{r}
# Install rEEMSplots
library(rEEMSplots)
```

```{r} 
# Plotting EEMS after running runeems_snps
path = "./NB_EEMS_Neutral/"
dirs = c(paste0(path,"neutraldata-D200-chain1"), paste0(path,"neutraldata-D300-chain1"), paste0(path,"neutraldata-D600-chain1"))
```

```{r}
eems.plots(mcmcpath = c(paste0(path,"./neutraldata-D200-chain1"), paste0(path,"neutraldata-D300-chain1"), paste0(path,"neutraldata-D600-chain1")), plotpath = paste0(path,"neutraldata-All-plots"),
           longlat = T,add.grid=F,add.outline = T,add.demes = T,
           projection.in = "+proj=longlat +datum=WGS84",projection.out = "+proj=merc +datum=WGS84",
           add.map = T,add.abline = T, add.r.squared = T)
```

The plots are saved to the same directory where `runeems_snps` was run. 

## Pairwise Fst

```{r}
fst.mat <- pairwise.WCfst(hf.filt)
```

```{r}
gindF.fst.mat.triN <- as.matrix(fst.mat)
colnames(gindF.fst.mat.triN) <- pop_order
rownames(gindF.fst.mat.triN) <- pop_order
```

```{r}
meltedN <- melt(gindF.fst.mat.triN, na.rm =TRUE)
round(gindF.fst.mat.triN,4)
```

```{r}
summary(meltedN$value)
```


```{r}
#Plotting Pairwise fst
neutral <- ggplot(data = meltedN, aes(Var2, Var1, fill = value))+ geom_tile(color = "white")+ 
  scale_fill_gradient(low = "white", high = "red", name="FST")  +
  ggtitle(expression(atop("Pairwise FST, WC (1984) All SNPs", atop(italic("N = 40, L = 75,402"), ""))))+
  labs( x = "Sampling Site", y = "Sampling Site") + 
  theme(axis.text.x = element_text(angle = 45, vjust = 1, size = 12, hjust = 1),axis.text.y = element_text(size = 12)) + 
  theme(axis.title = element_text(size = 12),legend.text = element_text(size =14), legend.title = element_text(size =14)) +
  theme(plot.title = element_text(size = 14)) +
coord_fixed()
neutral
```

## Genetic diversity (observed and expected heterozygosity)

```{r}
comb <- summary(stratted.filt)
names(comb)
```

```{r}
plot(comb$Hobs, xlab="Loci number", ylab="Observed Heterozygosity", 
     main="Observed heterozygosity per locus")
```


```{r}
plot(comb$Hobs,comb$Hexp, xlab="Hobs", ylab="Hexp", 
     main="Expected heterozygosity as a function of observed heterozygosity per locus")
```


```{r}
bartlett.test(list(comb$Hexp, comb$Hobs)) # a test : H0: Hexp = Hobs
```

**_Significant difference between Observed and expected heterzygosity_**.  

```{r}
basicstat <- basic.stats(hf.filt, diploid = TRUE, digits = 3)
```

```{r}
as.data.frame(basicstat$overall)
```

```{r}
# get bootstrap confidence values for Fis
boot <- boot.ppfis(hf.filt,nboot = 1000)
boot5 <- boot.ppfis(hf.filt,nboot = 1000,quant = 0.5)
```

```{r}
# add latitude for each population
latitude = c(41.545, 41.654, 41.505, 41.816)
```

```{r}
# add longitude for each population
longitude = c(-71.431, -71.445, -71.453, -71.391)
```

```{r}
# add distance for each population
distance = c(4.76, 0.47, 15.41, 1.49)
```

```{r}
# add sewage effluent for each population
sewage = c(8.82, 14.60, 2.03, 59.86)
```

```{r}
# add temperature for each population
temperature = c(23, 24, 25, 23)
```

```{r}
# add salinity for each population
salinity = c(30, 28, 18, 25)
```

```{r}
# add pH for each population
pH = c(7.9, 7.4, 7.6, 7.4)
```

```{r}
# add Chlorophylla for each population
Chlor_a = c(4.9, 18.8, 4.6, 8.1)
```

```{r}
# add DO for each population
DO = c(8.2, 5.7, 7, 4.9)
```

```{r}
# combine all pop statistics
colnames(basicstat$Ho) <- pop_order
Ho <- colMeans(basicstat$Ho,na.rm = T)
He <- colMeans(basicstat$Hs,na.rm = T)
Fis<- boot5$fis.ci$ll
y <- cbind(pop_order,Ho,He,Fis,boot$fis.ci,latitude, longitude, distance,sewage, temperature, salinity, pH, Chlor_a, DO)
y
```

```{r}
summary(He)
```

```{r}
summary(Fis)
```

```{r}
# Plot He vs Latitude
R2 = round(summary(lm(y$He ~ y$latitude))$r.squared, 4)             
ggplot(y, aes(x = latitude, y = He)) + geom_point(aes(shape=pop_order, color=pop_order), size = 3) + 
  scale_shape_manual(values = c(15,18,17,16)) +
geom_smooth(method=lm) + 
ggtitle("Expected heterozygosity vs Latitude, Neutral SNPs") +
annotate(geom = "text", label=paste("italic(R^2)==",R2), x=41.55, y=0.2635, parse=T) +
scale_x_continuous()
```

```{r}
# Plot Ho vs Latitude
R2 = round(summary(lm(y$Ho ~ y$latitude))$r.squared, 4)             
ggplot(y, aes(x = latitude, y = Ho)) + geom_point(aes(shape=pop_order, color=pop_order), size = 3) + 
  scale_shape_manual(values = c(15,18,17,16)) + 
geom_smooth(method=lm) + 
ggtitle("Observed heterozygosity vs Latitude, Neutral SNPs") +
annotate(geom = "text", label=paste("italic(R^2)==",R2), x=41.55, y=0.2635, parse=T) +
scale_x_continuous()
```

```{r}
#Plot He vs Longitude
R3 = round(summary(lm(y$He ~ y$longitude))$r.squared, 4)             
ggplot(y, aes(x = longitude, y = He)) + geom_point(aes(shape=pop_order, color=pop_order), size = 3) + 
  scale_shape_manual(values = c(15,18,17,16)) +
geom_smooth(method=lm) + 
ggtitle("Expected heterozygosity vs Longitude, Neutral SNPs") +
annotate(geom = "text", label=paste("italic(R^2)==",R3), x=-71.45, y=0.2635, parse=T) +
scale_x_reverse()
```

```{r}
#Plot Ho vs Longitude
R3 = round(summary(lm(y$Ho ~ y$longitude))$r.squared, 4)             
ggplot(y, aes(x = longitude, y = Ho)) + geom_point(aes(shape=pop_order, color=pop_order), size = 3) + 
  scale_shape_manual(values = c(15,18,17,16)) +
geom_smooth(method=lm) + 
ggtitle("Observed heterozygosity vs Longitude, Neutral SNPs") +
annotate(geom = "text", label=paste("italic(R^2)==",R3), x=-71.45, y=0.2635, parse=T) +
scale_x_reverse()
```

```{r}
#Plot He vs Distance from sewage outflow 
R4 = round(summary(lm(y$He ~ y$distance))$r.squared, 4)             
ggplot(y, aes(x = distance, y = He)) + geom_point(aes(shape=pop_order, color=pop_order), size = 3) + 
  scale_shape_manual(values = c(15,18,17,16)) + 
geom_smooth(method=lm) + 
ggtitle("Expected heterozygosity vs Distance, Neutral SNPs") +
annotate(geom = "text", label=paste("italic(R^2)==",R4), x=2, y=0.2685, parse=T) +
scale_x_reverse()
```

```{r}
#Plot Ho vs Distance from sewage outflow 
R4 = round(summary(lm(y$Ho ~ y$distance))$r.squared, 4)             
ggplot(y, aes(x = distance, y = Ho)) + geom_point(aes(shape=pop_order, color=pop_order), size = 3) + 
  scale_shape_manual(values = c(15,18,17,16)) + 
geom_smooth(method=lm) + 
ggtitle("Observed heterozygosity vs Distance, Neutral SNPs") +
annotate(geom = "text", label=paste("italic(R^2)==",R4), x=2, y=0.2635, parse=T) +
scale_x_reverse()
```

```{r}
#Plot He vs Sewage Effluent 
R4 = round(summary(lm(y$He ~ y$sewage))$r.squared, 4)             
ggplot(y, aes(x = sewage, y = He)) + geom_point(aes(shape=pop_order, color=pop_order), size = 3) + 
  scale_shape_manual(values = c(15,18,17,16)) + 
geom_smooth(method=lm) + 
ggtitle("Expected heterozygosity vs Sewage Effluent, Neutral SNPs") +
annotate(geom = "text", label=paste("italic(R^2)==",R4), x=20, y=0.265, parse=T) +
scale_x_reverse()
```

```{r}
#Plot Ho vs Sewage Effluent 
R4 = round(summary(lm(y$Ho ~ y$sewage))$r.squared, 4)             
ggplot(y, aes(x = sewage, y = Ho)) + geom_point(aes(shape=pop_order, color=pop_order), size = 3) + 
  scale_shape_manual(values = c(15,18,17,16)) + 
geom_smooth(method=lm) + 
ggtitle("Observed heterozygosity vs Sewage Effluent, Neutral SNPs") +
annotate(geom = "text", label=paste("italic(R^2)==",R4), x=20, y=0.2635, parse=T) +
scale_x_reverse()
```

```{r}
#Plot He vs Temperature 
R4 = round(summary(lm(y$He ~ y$temperature))$r.squared, 4)             
ggplot(y, aes(x = temperature, y = He)) + geom_point(aes(shape=pop_order, color=pop_order), size = 3) + 
  scale_shape_manual(values = c(15,18,17,16)) +
geom_smooth(method=lm) + 
ggtitle("Expected heterozygosity vs Temperature, Neutral SNPs") +
annotate(geom = "text", label=paste("italic(R^2)==",R4), x=23.25, y=0.2635, parse=T) +
scale_x_reverse()
```

```{r}
#Plot Ho vs Temperature 
R4 = round(summary(lm(y$Ho ~ y$temperature))$r.squared, 4)             
ggplot(y, aes(x = temperature, y = Ho)) + geom_point(aes(shape=pop_order, color=pop_order), size = 3) + 
  scale_shape_manual(values = c(15,18,17,16)) + 
geom_smooth(method=lm) + 
ggtitle("Observed heterozygosity vs Temperature, Neutral SNPs") +
annotate(geom = "text", label=paste("italic(R^2)==",R4), x=23.25, y=0.2635, parse=T) +
scale_x_reverse()
```

```{r}
#Plot He vs Salinity 
R4 = round(summary(lm(y$He ~ y$salinity))$r.squared, 4)             
ggplot(y, aes(x = salinity, y = He)) + geom_point(aes(shape=pop_order, color=pop_order), size = 3) + 
  scale_shape_manual(values = c(15,18,17,16)) + 
geom_smooth(method=lm) + 
ggtitle("Expected heterozygosity vs Salinity, Neutral SNPs") +
annotate(geom = "text", label=paste("italic(R^2)==",R4), x=23.25, y=0.2675, parse=T) +
scale_x_reverse()
```

```{r}
#Plot Ho vs Salinity 
R4 = round(summary(lm(y$Ho ~ y$salinity))$r.squared, 4)             
ggplot(y, aes(x = salinity, y = Ho)) + geom_point(aes(shape=pop_order, color=pop_order), size = 3) + 
  scale_shape_manual(values = c(15,18,17,16)) + 
geom_smooth(method=lm) + 
ggtitle("Observed heterozygosity vs Salinity, Neutral SNPs") +
annotate(geom = "text", label=paste("italic(R^2)==",R4), x=23.25, y=0.2635, parse=T) +
scale_x_reverse()
```

```{r}
#Plot He vs pH
R4 = round(summary(lm(y$He ~ y$pH))$r.squared, 4)             
ggplot(y, aes(x = pH, y = He)) + geom_point(aes(shape=pop_order, color=pop_order), size = 3) + 
  scale_shape_manual(values = c(15,18,17,16)) +
geom_smooth(method=lm) + 
ggtitle("Expected heterozygosity vs pH, Neutral SNPs") +
annotate(geom = "text", label=paste("italic(R^2)==",R4), x=7.45, y=0.2635, parse=T) +
scale_x_reverse()
```

```{r}
#Plot Ho vs pH
R4 = round(summary(lm(y$Ho ~ y$pH))$r.squared, 4)             
ggplot(y, aes(x = pH, y = Ho)) + geom_point(aes(shape=pop_order, color=pop_order), size = 3) + 
  scale_shape_manual(values = c(15,18,17,16)) + 
geom_smooth(method=lm) + 
ggtitle("Observed heterozygosity vs pH, Neutral SNPs") +
annotate(geom = "text", label=paste("italic(R^2)==",R4), x=7.45, y=0.2635, parse=T) +
scale_x_reverse()
```

```{r}
#Plot He vs Chlorophyll a 
R4 = round(summary(lm(y$He ~ y$Chlor_a))$r.squared, 4)             
ggplot(y, aes(x = Chlor_a, y = He)) + geom_point(aes(shape=pop_order, color=pop_order), size = 3) + 
  scale_shape_manual(values = c(15,18,17,16)) + 
geom_smooth(method=lm) + 
ggtitle("Expected heterozygosity vs Chlorophyll a, Neutral SNPs") +
annotate(geom = "text", label=paste("italic(R^2)==",R4), x=6, y=0.2665, parse=T) +
scale_x_reverse()
```

```{r}
#Plot Ho vs Chlorophyll a 
R4 = round(summary(lm(y$Ho ~ y$Chlor_a))$r.squared, 4)             
ggplot(y, aes(x = Chlor_a, y = Ho)) + geom_point(aes(shape=pop_order, color=pop_order), size = 3) + 
  scale_shape_manual(values = c(15,18,17,16)) + 
geom_smooth(method=lm) + 
ggtitle("Observed heterozygosity vs Chlorophyll a, Neutral SNPs") +
annotate(geom = "text", label=paste("italic(R^2)==",R4), x=6, y=0.2665, parse=T) +
scale_x_reverse()
```

```{r}
#Plot He vs Dissolved Oxygen 
R4 = round(summary(lm(y$He ~ y$DO))$r.squared, 4)             
ggplot(y, aes(x = DO, y = He)) + geom_point(aes(shape=pop_order, color=pop_order), size = 3) + 
  scale_shape_manual(values = c(15,18,17,16)) + 
geom_smooth(method=lm) + 
ggtitle("Expected heterozygosity vs Dissolved Oxygen, Neutral SNPs") +
annotate(geom = "text", label=paste("italic(R^2)==",R4), x=5.5, y=0.2635, parse=T) +
scale_x_reverse()
```

```{r}
#Plot Ho vs Dissolved Oxygen 
R4 = round(summary(lm(y$Ho ~ y$DO))$r.squared, 4)             
ggplot(y, aes(x = DO, y = Ho)) + geom_point(aes(shape=pop_order, color=pop_order), size = 3) + 
  scale_shape_manual(values = c(15,18,17,16)) + 
geom_smooth(method=lm) + 
ggtitle("Observed heterozygosity vs Dissolved Oxygen, Neutral SNPs") +
annotate(geom = "text", label=paste("italic(R^2)==",R4), x=5.5, y=0.2635, parse=T) +
scale_x_reverse()
```


## PCA 

```{r}
X <- tab(stratted.u, freq = TRUE, NA.method = "mean")
pca1 <- dudi.pca(X, scale = FALSE, scannf = FALSE, nf = 3)
barplot(pca1$eig[1:50], main = "PCA eigenvalues", col = heat.colors(50))
s.class(pca1$li, pop(stratted.filt))
title("PCA of neutral SNPs\naxes 1-2")
add.scatter.eig(pca1$eig[1:20], 3,1,2)
col <- funky(15) 
s.class(pca1$li, pop(stratted.filt),xax=1,yax=2, col=col, axesell=FALSE, cstar=0, cpoint=3, grid=FALSE)
```


## Discriminant Analysis of Principal Components (DAPC)

The first line of the code block below is interactive. In the console, it will tell you:  
1. Choose the number PCs to retain (>=1): 
    I picked 1 
2. Choose the number discriminant functions to retain (>=1):
    I picked 2

```{r}
grp <- find.clusters(stratted.u, n.pca = 1, choose.n.clust = FALSE)
table(pop(stratted.u), grp$grp)
table.value(table(pop(stratted.u), grp$grp), col.lab=paste("inf", 1:2), row.lab=paste("ori", 1:4))
```

Again, the first line of the code block below is interactive. In the console, it will tell you:  
1. Choose the number PCs to retain (>=1): 
    I picked 1 
2. Choose the number discriminant functions to retain (>=1):
    I picked 1

```{r}
dapc1 <- dapc(stratted.u, grp$grp, n.pca = 1, n.da= 1)
scatter(dapc1,col=col,bg="white", solid=1)
```

Again, the first line of the code block below is interactive. In the console, it will tell you:  
1. Choose the number PCs to retain (>=1): 
    I picked 1 
2. Choose the number discriminant functions to retain (>=1):
    I picked 1

```{r}
contrib <- loadingplot(dapc1$var.contr, axis=1, thres=.01, lab.jitter=1)
contrib
setPop(rad.filt) <- ~Library
dapc1 <- dapc(stratted.u, pop(stratted.u), n.pca = 1, n.da = 1)
contrib <- loadingplot(dapc1$var.contr, axis=1, thres=.05, lab.jitter=1)
```

```{r}
#Structure Like
compoplot(dapc1, posi="bottomright",txt.leg=paste("Cluster", 1:4), lab="", ncol=1, xlab="individuals")
temp <- which(apply(dapc1$posterior,1, function(e) all(e<0.9)))
compoplot(dapc1, subset=temp, posi="bottomright", txt.leg=paste("Cluster", 1:4), ncol=2)
```


### PCAviz

```{r}
NA.afDraw<- function(ind){
  ind.mat <- ind@tab
  new.mat <- ind.mat
  af = colSums(ind.mat[,seq(1,ncol(ind.mat)-1,2)],na.rm = TRUE)/
      (2*apply(ind.mat[,seq(1,ncol(ind.mat)-1,2)],2,function(x) sum(!is.na(x))))
  af.Draw <- function(geno, af){
     new <- function(geno,af){
        if(is.na(geno)){
        newA = rbinom(1,2,af)
        }
        else {newA <- geno}
        return(newA)
   }
  new.row <- mapply(geno,af,FUN = new)
  return(new.row)}
  
  new.mat[,seq(1,ncol(ind.mat)-1,2)] <- t(apply(ind.mat[,seq(1,ncol(ind.mat)-1,2)],1,af.Draw,af))
  new.mat[,seq(2,ncol(ind.mat),2)] <- 2-new.mat[,seq(1,ncol(ind.mat)-1,2)]
  new.ind <- ind
  new.ind@tab <- new.mat
  return(new.ind)
}
```

```{r}
u.na <- NA.afDraw(stratted.u)
```

```{r}
pca <- dudi.pca(u.na,center=TRUE,scale=TRUE,scannf = F, nf = 30)
```

```{r}
col18 <- funky(length(unique(u.na@strata$Population)))
#Colors that match the neutral Structure results
col6 <-  c("firebrick","dodgerblue","midnightblue","salmon")
par(mfrow=c(2,2))
s.class(pca$li, strata(u.na)$Population,xax=1,yax=2, 
        sub = "neutralloci,PC 1-2, 40 individuals, 14,880 loci, maf 5.0%",
        possub = "topleft",col=transp(col6,.6),axesell=FALSE,
        cstar=0, cpoint=3, grid=FALSE, cellipse = 0)
s.class(pca$li, strata(u.na)$Population,xax=1,yax=3, 
        sub = "neutralloci,PC 1-3, 40 individuals, 14,880 loci, maf 5.0%",
        possub = "topleft",col=transp(col6,.6),axesell=FALSE,
        cstar=0, cpoint=3, grid=FALSE, cellipse = 0)
s.class(pca$li, strata(u.na)$Population,xax=1,yax=4, 
        sub = "neutralloci,PC 1-4, 40 individuals, 14,880 loci, maf 5.0%",
        possub = "topleft",col=transp(col6,.6),axesell=FALSE,
        cstar=0, cpoint=3, grid=FALSE, cellipse = 0)
s.class(pca$li, strata(u.na)$Population,xax=1,yax=5, 
        sub = "neutralloci,PC 1-5, 40 individuals, 14,880 loci, maf 5.0%",
        possub = "topleft",col=transp(col6,.6),axesell=FALSE,
        cstar=0, cpoint=3, grid=FALSE, cellipse = 0)
```

```{r}
s.label(pca$li, xax=1,yax=5, 
        sub = "neutralloci,PC 1-2, 40 individuals, 14,880 loci, maf 5.0%",
        possub = "topleft")
```


```{r}
eig.perc <- 100*pca$eig/sum(pca$eig)
head(eig.perc)
```

```{r}
li <-pca$li
c1 <- pca$c1
#Create dataframe of info like latitude and population for each individual
info_mat <- as.data.frame(cbind(u.na@strata, u.na@other$Latitude, u.na@other$Longitude, u.na@other$Distance, u.na@other$Temperature, u.na@other$Salinity, u.na@other$pH, u.na@other$Chlorophylla, u.na@other$DO))
colnames(info_mat) <- c("Population","Latitude", "Longitude", "Distance","Temperature","Salinity","pH","Chlorophylla","DO")
colnames(c1) <- colnames(li)
#create pcaviz object
pviz <- pcaviz(x=li,rotation=c1,dat=info_mat)
```

```{r}
p = list(size=2) 
plot(pviz,color = "Distance", draw.points = T, group.summary.labels = F, draw.pc.axes = T, geom.point.params = p)
```

```{r}
# Plot against distance 
p = list(size=6)
plot(pviz, coords = c("Axis1","Distance"),
     show.legend = T,color = "Population",colors = col18,
     draw.points =T,group.summary.labels = F,geom.point.params= list(size = 6, alpha = 0.9),geom.point.summary.params = list(size =6,alpha = 0.4))
```

```{r}
plot(pviz,coords = c("Axis1","Axis2","Axis3","Axis4","Axis5","Distance"),group = NULL,color= "Population",colors = col18,draw.points = T,scale.pc.axes = 0.6, show.legend=F)
```

```{r}
p = list(size=6)
A2 <- plot(pviz,coords = c("Axis2","Distance"),group="Population",
     show.legend = F,color = "Population",colors = col6,draw.points =T,
     group.summary.labels = F,draw.linear.fit = F, coord_fixed=T)
A3 <- plot(pviz,coords = c("Axis3","Distance"),group="Population",
     show.legend = F,color = "Population",colors = col6,
     draw.points =T,group.summary.labels = F,draw.linear.fit = F)
A4 <- plot(pviz,coords = c("Axis4","Distance"),group="Population",
     show.legend = F,color = "Population",colors = col6,
     draw.points =T,group.summary.labels = F,draw.linear.fit = F)
A5 <- plot(pviz,coords = c("Axis5","Distance"),group="Population",
     show.legend = T,color = "Population",colors = col6,
     draw.points =T,group.summary.labels = F,draw.linear.fit = F)
plot_grid(A2,A3,A4,A5, labels = c('A','B','C','D'))
```

# Seascape Redundancy Analysis

This code follows that documented by [Tom Jenkins](https://github.com/Tom-Jenkins/seascape_rda_tutorial).

## [Prepare genetic data for redundancy analysis](https://github.com/Tom-Jenkins/seascape_rda_tutorial/blob/master/1.Prepare_genetic_data/1.prepare_genetic_data.R).

#### Notes before execution:
1. Make sure all required R packages are installed.
2. Set working directory to the location of this R script.

```{r}
# Load packages
library(adegenet)
library(poppr)
library(dplyr)
library(reshape2)
library(ggplot2)
library(vcfR)
```


```{r}
# Explore data
rad.filt
nLoc(rad.filt) # number of loci
nPop(rad.filt) # number of sites
nInd(rad.filt) # number of individuals
summary(rad.filt$pop) # sample size
```

```{r}
# Calculate allele frequencies for each site
allele_freqs = data.frame(rraf(rad.filt, by_pop = TRUE, correction = FALSE), check.names = FALSE)
```

```{r}
# Keep only the first of the two alleles for each SNP (since p=1-q).
allele_freqs = allele_freqs[, seq(1, dim(allele_freqs)[2], 2)]
```

```{r}
# Export allele frequencies
write.csv(allele_freqs, file = "all_allele_freqs.csv", row.names = TRUE)
```

#--------------#
#
# Calculate minor allele frequencies
#
#--------------#

```{r}
# Separate genind object by site
site_list = seppop(rad.filt)
names(site_list)
```

```{r}
# Calculate the minor allele frequency for each site
maf_list = lapply(site_list, FUN = minorAllele)
```

```{r}
# Convert list to dataframe
maf = as.data.frame(maf_list) %>% t() %>% as.data.frame()
```

```{r}
# Export minor allele frequencies
write.csv(maf, file = "minor_allele_freqs.csv", row.names = TRUE)
```

# ----------------- #
#
# Visualise allele frequencies
#
# ----------------- #

```{r}
# Add site labels
allele_freqs$site = rownames(allele_freqs)
```

```{r}
# Function to add regional labels to dataframe
addregion = function(x){
  # If pop label is present function will output the region
  if(x=="BIS") y = " Bissel Cove "
  if(x=="GB") y = " Greenwich Bay "
  if(x=="NAR") y = " Narrow River "
  if(x=="PVD") y = " Bold Point Park "
  return(y)
}
```

```{r}
# Add regional labels
allele_freqs$region = sapply(rownames(allele_freqs), addregion)
```

```{r}
# Convert dataframe to long format
allele_freqs.long = melt(allele_freqs, id.vars=c("site","region"))
```

```{r}
# Define order of facets using the levels argument in factor
unique(allele_freqs.long$site)
site_order =  c("BIS","GB","NAR","PVD")
allele_freqs.long$site_ord = factor(allele_freqs.long$site, levels = site_order)
```

```{r}
# Define region order
region_order = c(" Bissel Cove "," Greenwich Bay "," Narrow River ", " Bold Point Park ")
allele_freqs.long$region = factor(allele_freqs.long$region, levels = region_order)
```

```{r}
# Create colour scheme
# blue=#377EB8, green=#7FC97F, orange=#FDB462, red=#E31A1C
col_scheme = c("#7FC97F","#377EB8","#FDB462","#E31A1C")
```


A subset of the putatively outlier loci was selected, spanning all 4 outlier detection programs. 

```{r}
# Vector of outlier SNP loci to subset
desired_loci = c("NC_035780.1_2492392","NC_035780.1_58431616","NC_035780.1_2492005","NC_035782.1_39657559","NC_035783.1_56550670","NC_035784.1_65251317","NC_035785.1_32466402","NC_035786.1_26380600","NC_035788.1_60313741")
desired_loci_ID = sapply(paste(desired_loci, "..", sep = ""),
                         grep,
                         levels(allele_freqs.long$variable),
                         value = TRUE) %>% as.vector()
```

```{r}
# Subset dataset to plot desired SNP loci
allele_freqs.sub = allele_freqs.long %>% filter(variable %in% desired_loci_ID)
```

```{r}
# ggplot2 theme
ggtheme = theme(
  axis.text.x = element_blank(),
  axis.text.y = element_text(colour="black", size=6),
  axis.title = element_text(colour="black", size=15),
  panel.background = element_rect(fill="white"),
  panel.grid.minor = element_blank(),
  panel.grid.major = element_blank(),
  panel.border = element_rect(colour="black", fill=NA, size=0.5),
  plot.title = element_text(hjust = 0.5, size=18),
  legend.title = element_blank(),
  legend.text = element_text(size=10),
  legend.position = "top",
  legend.justification = "centre",
  # facet labels
  strip.text = element_text(colour="black", size=8)
)
```

```{r}
# Plot barplot
ggplot(data = allele_freqs.sub, aes(x = site_ord, y = value, fill = region))+
  geom_bar(stat = "identity", colour = "black", size = 0.3)+
  facet_wrap(~variable, scales = "free")+
  scale_y_continuous(limits = c(0,1), expand = c(0,0))+
  scale_fill_manual(values = col_scheme)+
  ylab("Allele frequency")+
  xlab("Site")+
  ggtheme
ggsave("allele_freq.png", width=10, height=8, dpi=300)
ggsave("allele_freq.pdf", width=10, height=8)
```


## [Prepare environmental data for redundancy analysis](https://github.com/Tom-Jenkins/seascape_rda_tutorial/blob/master/3.Prepare_environmental_data/3.prepare_env_data.R).

#### Environmental variables:
- Distance from sewage effluent source (km)
- Sewage Effluent (PW stats)
- Mean temperature (deg C)
- Mean Salinity (psu)
- Mean pH
- Mean Chlorophyll-a (ug/L)
- Mean Dissolved Oxygen (mg/L)

Environmental data for each population is saved in a `strata_pop` file that can be accessed [here](https://github.com/amyzyck/EecSeq_NB_EasternOyster/blob/master/Analysis/PopGen_SeaGen_Analyses/strata_pop)

```{r}
# All environmental data was previously saved in strata file 
strata_pop <- read.table("strata_pop", header=TRUE)
strata_pop
```

```{r}
# Export data as a csv file
write.csv(strata_pop, file="environmental_data.csv", row.names = FALSE)
```

I also prepared spatial data for the redundancy analysis which is documented [here](https://github.com/amyzyck/EecSeq_NB_EasternOyster/blob/master/Analysis/PopGen_SeaGen_Analyses/RedundancyAnalysis/PrepSpatialData.Rmd).

Allele frequency, environmental, and spatial csv files are saved to your working directory and must be imported into the Rscript to run the redundancy analysis. Documentation of the RDA can be accessed [here](https://github.com/amyzyck/EecSeq_NB_EasternOyster/blob/master/Analysis/PopGen_SeaGen_Analyses/RedundancyAnalysis/RDA_Outlier_Hap.Rmd). 

RDA and pRDA output plots can be viewed [here](https://github.com/amyzyck/EecSeq_NB_EasternOyster/tree/master/Output/RedundancyAnalysis/NonHapMasked_All).