---
title: "NB_PopSeaGenAnalysis_OutlierSNPs"
author: "Amy Zyck"
date: '2024-07-22'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Population and Seascape Genomics Analyses on Outlier SNP data sets 

Outlier SNPs were identified across four different detection programs documented [here](https://github.com/amyzyck/EecSeq_NB_EasternOyster/blob/master/Analysis/Analysis_Final/NB_OutlierDetection_Final.Rmd). There are seven different outlier sets: 

1.  Original full set of outlier SNPs - 70 loci (outlierloci6p9g9_red.recode.vcf)
2.  pcadapt outliers only - 56 loci (outlierloci6p9g9_db.recode.vcf)
3.  pcadapt outliers inside inversions - 34 loci (outliers6p9g9_db_invert.recode.vcf)
4.  pcadapt outliers outside inversions - 22 loci (outliers6p9g9_db_noinvert.recode.vcf)
5.  LFMM outliers only - 14 loci (outlierloci6p9g9_lfmm.recode.vcf)
6.  LFMM outliers inside inversions - 6 loci (outliers6p9g9_lfmm_invert.recode.vcf)
7.  LFMM outliers outside inversions - 8 loci (outliers6p9g9_lfmm_noinvert.recode.vcf)
8.  RDA outliers - 291 loci (outliers_rda.recode.vcf)
9.  New full set of outlier SNPs with RDA added (outlierloci_new.recode.vcf) 

On each outlier SNP set, I will be measuring population pairwise Fst (population differentiation) and heterozygosity (genetic diversity). For the  pcadapt outlier SNP sets, I will also run a redundancy analysis to investigate associations with environmental predictors. 


### Combine all outlier loci into one file

To make this easier:

```
$ cat outlier_pcadapt_thinned500.loci.txt outlier_lfmm2_6p9g9_red.loci.txt outliers_rda_sal_min.loci.txt > all.new.outliers
$ cat all.new.outliers | sort | uniq | wc -l

    output:
    361
```

```
$ vcftools --vcf SNP.TRS6newdp10mafp9g9nDNAmaf052A.recode.vcf --recode --recode-INFO-all --positions all.new.outliers --out outlierloci_new

After filtering, kept 60 out of 60 Individuals
Outputting VCF file...
After filtering, kept 361 out of a possible 22874 Sites
Run Time = 0.00 seconds
```


## Organizing vcf files

In terminal:

In `NB_ddhaplo_working` directory, make a new directory for Pop and Sea Gen analyses

```
$ mkdir NB_PopSeaGen_Analysis_working 
$ cd NB_PopSeaGen_Analysis_working
```

Make a sub directory for the outlier SNP sets and link vcf files 

```
$ mkdir Outlier_SNPs_6p9g9
$ cd Outlier_SNPs_6p9g9

$ ln -s ../../NB_OutlierDetection_working/outlierloci6p9g9* .
$ ln -s ../../NB_OutlierDetection_working/outliers6p9g9* .
```

The markdown file will be organized by analysis type. I'll repeat the steps for each analysis across the four different outlier SNP sets, following this order and naming scheme:

1. Full set of outlier SNPs - full
2.  pcadapt outliers only - db
3.  pcadapt outliers inside inversions - db_invert
4.  pcadapt outliers outside inversions - db_noinvert
5.  lfmm outliers only - lfmm
6.  lfmm outliers inside inversions - lfmm_invert
7.  lfmm outliers outside inversions - lfmm_noinvert
8.  RDA outliers - rda
9.  New set of outlier SNPs - all

This organization will allow me to more easily compare results across files within each analysis. 

```{r}
#Loading all the necessary packages
library(adegenet)
library(vcfR)
library("radiator") # Conversion from vcf to a lot of other formats
library("dplyr")
library("hierfstat")
library("ggplot2") #For plotting
library("reshape2") #For plotting
#library("plyr")
#library("cowplot") #For plotting manuscript figs
library(PCAviz)  #Visualizing output of PCA
#library("stringr")
library(tidyverse)
library("poppr")
library(eulerr)
```

## Making files

```{r}
# Setting up strata file - population info and environmental data 
strata6 <- read.table("../strata6_red_12var", header=TRUE)

# Making new version of strata file with one row per population
strata_6pops <- strata6 %>% distinct(Population, .keep_all = TRUE)
```

```{r}
#Providing population names for plotting
pop_order_6pops <- c("BAR","BIS","GB","KIC","MCD","PVD")
```

**Make genind object**

Full SNP set 

```{r}
# VCF file with full SNP dataset
my_vcf_out_full <- read.vcfR("outlierloci6p9g9_red.recode.vcf")
```

```{r}
rad_out_full.filt <- vcfR2genind(my_vcf_out_full, strata = strata6, pop = c(rep("BAR", 10),rep("BIS", 10),rep("GB", 10),rep("KIC", 10),rep("MCD", 10), rep("PVD",10)))

rad_out_full.filt
```

```{r}
info_out_full <- as.data.frame(read.table("../strata6_red_12var",header = T,sep = "\t",stringsAsFactors = F))
mystrats_out_full <- as.data.frame(matrix(nrow = length(indNames(rad_out_full.filt)),ncol=16))
just_out_full.strats <- select(info_out_full,c("Population"))
stratted_out_full.filt <- strata(rad_out_full.filt, formula= Population, combine = TRUE,just_out_full.strats)
stratted_out_full.filt@other <- info_out_full[,4:17]

stratted_out_full.filt
```

**Make hierfstat object**

```{r}
hf_out_full.filt <- genind2hierfstat(rad_out_full.filt, pop = c(rep("BAR", 10),rep("BIS", 10),rep("GB", 10),rep("KIC", 10),rep("MCD", 10), rep("PVD",10)))
```


pcadapt (db) SNP set 

```{r}
# VCF file with db SNP dataset
my_vcf_out_db <- read.vcfR("outlierloci6p9g9_db.recode.vcf")
```

```{r}
rad_out_db.filt <- vcfR2genind(my_vcf_out_db, strata = strata6, pop = c(rep("BAR", 10),rep("BIS", 10),rep("GB", 10),rep("KIC", 10),rep("MCD", 10), rep("PVD",10)))

rad_out_db.filt
```

```{r}
info_out_db <- as.data.frame(read.table("../strata6_red_12var",header = T,sep = "\t",stringsAsFactors = F))
mystrats_out_db <- as.data.frame(matrix(nrow = length(indNames(rad_out_db.filt)),ncol=16))
just_out_db.strats <- select(info_out_db,c("Population"))
stratted_out_db.filt <- strata(rad_out_db.filt, formula= Population, combine = TRUE,just_out_db.strats)
stratted_out_db.filt@other <- info_out_db[,4:17]

stratted_out_db.filt
```

**Make hierfstat object**

```{r}
hf_out_db.filt <- genind2hierfstat(rad_out_db.filt, pop = c(rep("BAR", 10),rep("BIS", 10),rep("GB", 10),rep("KIC", 10),rep("MCD", 10), rep("PVD",10)))
```


pcadapt inside inversion (db_invert) SNP set 

```{r}
# VCF file with db_invert SNP dataset
my_vcf_out_db_invert <- read.vcfR("outliers6p9g9_db_invert.recode.vcf")
```

```{r}
rad_out_db_invert.filt <- vcfR2genind(my_vcf_out_db_invert, strata = strata6, pop = c(rep("BAR", 10),rep("BIS", 10),rep("GB", 10),rep("KIC", 10),rep("MCD", 10), rep("PVD",10)))

rad_out_db_invert.filt
```

```{r}
info_out_db_invert <- as.data.frame(read.table("../strata6_red_12var",header = T,sep = "\t",stringsAsFactors = F))
mystrats_out_db_invert <- as.data.frame(matrix(nrow = length(indNames(rad_out_db_invert.filt)),ncol=16))
just_out_db_invert.strats <- select(info_out_db_invert,c("Population"))
stratted_out_db_invert.filt <- strata(rad_out_db_invert.filt, formula= Population, combine = TRUE,just_out_db_invert.strats)
stratted_out_db_invert.filt@other <- info_out_db_invert[,4:17]

stratted_out_db_invert.filt
```

**Make hierfstat object**

```{r}
hf_out_db_invert.filt <- genind2hierfstat(rad_out_db_invert.filt, pop = c(rep("BAR", 10),rep("BIS", 10),rep("GB", 10),rep("KIC", 10),rep("MCD", 10), rep("PVD",10)))
```


pcadapt outside inversion (db_noinvert) SNP set 

```{r}
# VCF file with db_noinvert SNP dataset
my_vcf_out_db_noinvert <- read.vcfR("outliers6p9g9_db_noinvert.recode.vcf")
```

```{r}
rad_out_db_noinvert.filt <- vcfR2genind(my_vcf_out_db_noinvert, strata = strata6, pop = c(rep("BAR", 10),rep("BIS", 10),rep("GB", 10),rep("KIC", 10),rep("MCD", 10), rep("PVD",10)))

rad_out_db_noinvert.filt
```

```{r}
info_out_db_noinvert <- as.data.frame(read.table("../strata6_red_12var",header = T,sep = "\t",stringsAsFactors = F))
mystrats_out_db_noinvert <- as.data.frame(matrix(nrow = length(indNames(rad_out_db_noinvert.filt)),ncol=16))
just_out_db_noinvert.strats <- select(info_out_db_noinvert,c("Population"))
stratted_out_db_noinvert.filt <- strata(rad_out_db_noinvert.filt, formula= Population, combine = TRUE,just_out_db_noinvert.strats)
stratted_out_db_noinvert.filt@other <- info_out_db_noinvert[,4:17]

stratted_out_db_noinvert.filt
```

**Make hierfstat object**

```{r}
hf_out_db_noinvert.filt <- genind2hierfstat(rad_out_db_noinvert.filt, pop = c(rep("BAR", 10),rep("BIS", 10),rep("GB", 10),rep("KIC", 10),rep("MCD", 10), rep("PVD",10)))
```


LFMM (lfmm) SNP set 

```{r}
# VCF file with lfmm SNP dataset
my_vcf_out_lfmm <- read.vcfR("outlierloci6p9g9_lfmm.recode.vcf")
```

```{r}
rad_out_lfmm.filt <- vcfR2genind(my_vcf_out_lfmm, strata = strata6, pop = c(rep("BAR", 10),rep("BIS", 10),rep("GB", 10),rep("KIC", 10),rep("MCD", 10), rep("PVD",10)))

rad_out_lfmm.filt
```

```{r}
info_out_lfmm <- as.data.frame(read.table("../strata6_red_12var",header = T,sep = "\t",stringsAsFactors = F))
mystrats_out_lfmm <- as.data.frame(matrix(nrow = length(indNames(rad_out_lfmm.filt)),ncol=16))
just_out_lfmm.strats <- select(info_out_lfmm,c("Population"))
stratted_out_lfmm.filt <- strata(rad_out_lfmm.filt, formula= Population, combine = TRUE,just_out_lfmm.strats)
stratted_out_lfmm.filt@other <- info_out_lfmm[,4:17]

stratted_out_lfmm.filt
```

**Make hierfstat object**

```{r}
hf_out_lfmm.filt <- genind2hierfstat(rad_out_lfmm.filt, pop = c(rep("BAR", 10),rep("BIS", 10),rep("GB", 10),rep("KIC", 10),rep("MCD", 10), rep("PVD",10)))
```


LFMM inside inversion (lfmm_invert) SNP set 

```{r}
# VCF file with lfmm_invert SNP dataset
my_vcf_out_lfmm_invert <- read.vcfR("outliers6p9g9_lfmm_invert.recode.vcf")
```

```{r}
rad_out_lfmm_invert.filt <- vcfR2genind(my_vcf_out_lfmm_invert, strata = strata6, pop = c(rep("BAR", 10),rep("BIS", 10),rep("GB", 10),rep("KIC", 10),rep("MCD", 10), rep("PVD",10)))

rad_out_lfmm_invert.filt
```

```{r}
info_out_lfmm_invert <- as.data.frame(read.table("../strata6_red_12var",header = T,sep = "\t",stringsAsFactors = F))
mystrats_out_lfmm_invert <- as.data.frame(matrix(nrow = length(indNames(rad_out_lfmm_invert.filt)),ncol=16))
just_out_lfmm_invert.strats <- select(info_out_lfmm_invert,c("Population"))
stratted_out_lfmm_invert.filt <- strata(rad_out_lfmm_invert.filt, formula= Population, combine = TRUE,just_out_lfmm_invert.strats)
stratted_out_lfmm_invert.filt@other <- info_out_lfmm_invert[,4:17]

stratted_out_lfmm_invert.filt
```

**Make hierfstat object**

```{r}
hf_out_lfmm_invert.filt <- genind2hierfstat(rad_out_lfmm_invert.filt, pop = c(rep("BAR", 10),rep("BIS", 10),rep("GB", 10),rep("KIC", 10),rep("MCD", 10), rep("PVD",10)))
```


LFMM outside inversion (lfmm_noinvert) SNP set 

```{r}
# VCF file with lfmm_noinvert SNP dataset
my_vcf_out_lfmm_noinvert <- read.vcfR("outliers6p9g9_lfmm_noinvert.recode.vcf")
```

```{r}
rad_out_lfmm_noinvert.filt <- vcfR2genind(my_vcf_out_lfmm_noinvert, strata = strata6, pop = c(rep("BAR", 10),rep("BIS", 10),rep("GB", 10),rep("KIC", 10),rep("MCD", 10), rep("PVD",10)))

rad_out_lfmm_noinvert.filt
```

```{r}
info_out_lfmm_noinvert <- as.data.frame(read.table("../strata6_red_12var",header = T,sep = "\t",stringsAsFactors = F))
mystrats_out_lfmm_noinvert <- as.data.frame(matrix(nrow = length(indNames(rad_out_lfmm_noinvert.filt)),ncol=16))
just_out_lfmm_noinvert.strats <- select(info_out_lfmm_noinvert,c("Population"))
stratted_out_lfmm_noinvert.filt <- strata(rad_out_lfmm_noinvert.filt, formula= Population, combine = TRUE,just_out_lfmm_noinvert.strats)
stratted_out_lfmm_noinvert.filt@other <- info_out_lfmm_noinvert[,4:17]

stratted_out_lfmm_noinvert.filt
```

**Make hierfstat object**

```{r}
hf_out_lfmm_noinvert.filt <- genind2hierfstat(rad_out_lfmm_noinvert.filt, pop = c(rep("BAR", 10),rep("BIS", 10),rep("GB", 10),rep("KIC", 10),rep("MCD", 10), rep("PVD",10)))
```

RDA (rda) SNP set 

```{r}
# VCF file with rda SNP dataset
my_vcf_out_rda <- read.vcfR("outliers_rda_final.recode.vcf")
```

```{r}
rad_out_rda.filt <- vcfR2genind(my_vcf_out_rda, strata = strata6, pop = c(rep("BAR", 10),rep("BIS", 10),rep("GB", 10),rep("KIC", 10),rep("MCD", 10), rep("PVD",10)))

rad_out_rda.filt
```

```{r}
info_out_rda <- as.data.frame(read.table("../strata6_red_12var",header = T,sep = "\t",stringsAsFactors = F))
mystrats_out_rda <- as.data.frame(matrix(nrow = length(indNames(rad_out_rda.filt)),ncol=16))
just_out_rda.strats <- dplyr::select(info_out_rda,c("Population"))
stratted_out_rda.filt <- strata(rad_out_rda.filt, formula= Population, combine = TRUE,just_out_rda.strats)
stratted_out_rda.filt@other <- info_out_rda[,4:17]

stratted_out_rda.filt
```

**Make hierfstat object**

```{r}
hf_out_rda.filt <- genind2hierfstat(rad_out_rda.filt, pop = c(rep("BAR", 10),rep("BIS", 10),rep("GB", 10),rep("KIC", 10),rep("MCD", 10), rep("PVD",10)))
```

full_new (full_new) SNP set 

```{r}
# VCF file with full_new SNP dataset
my_vcf_out_full_new <- read.vcfR("outlierloci_new.recode.vcf")
```

```{r}
rad_out_full_new.filt <- vcfR2genind(my_vcf_out_full_new, strata = strata6, pop = c(rep("BAR", 10),rep("BIS", 10),rep("GB", 10),rep("KIC", 10),rep("MCD", 10), rep("PVD",10)))

rad_out_full_new.filt
```

```{r}
info_out_full_new <- as.data.frame(read.table("../strata6_red_12var",header = T,sep = "\t",stringsAsFactors = F))
mystrats_out_full_new <- as.data.frame(matrix(nrow = length(indNames(rad_out_full_new.filt)),ncol=16))
just_out_full_new.strats <- dplyr::select(info_out_full_new,c("Population"))
stratted_out_full_new.filt <- strata(rad_out_full_new.filt, formula= Population, combine = TRUE,just_out_full_new.strats)
stratted_out_full_new.filt@other <- info_out_full_new[,4:17]

stratted_out_full_new.filt
```

**Make hierfstat object**

```{r}
hf_out_full_new.filt <- genind2hierfstat(rad_out_full_new.filt, pop = c(rep("BAR", 10),rep("BIS", 10),rep("GB", 10),rep("KIC", 10),rep("MCD", 10), rep("PVD",10)))
```


full_final (full_final) SNP set 

```{r}
# VCF file with full_final SNP dataset
my_vcf_out_full_final <- read.vcfR("outliers_full_final.recode.vcf")
```

```{r}
rad_out_full_final.filt <- vcfR2genind(my_vcf_out_full_final, strata = strata6, pop = c(rep("BAR", 10),rep("BIS", 10),rep("GB", 10),rep("KIC", 10),rep("MCD", 10), rep("PVD",10)))

rad_out_full_final.filt
```

```{r}
info_out_full_final <- as.data.frame(read.table("../strata6_red_12var",header = T,sep = "\t",stringsAsFactors = F))
mystrats_out_full_final <- as.data.frame(matrix(nrow = length(indNames(rad_out_full_final.filt)),ncol=16))
just_out_full_final.strats <- dplyr::select(info_out_full_final,c("Population"))
stratted_out_full_final.filt <- strata(rad_out_full_final.filt, formula= Population, combine = TRUE,just_out_full_final.strats)
stratted_out_full_final.filt@other <- info_out_full_final[,4:17]

stratted_out_full_final.filt
```

**Make hierfstat object**

```{r}
hf_out_full_final.filt <- genind2hierfstat(rad_out_full_final.filt, pop = c(rep("BAR", 10),rep("BIS", 10),rep("GB", 10),rep("KIC", 10),rep("MCD", 10), rep("PVD",10)))
```


## Pairwise Fst  

Full SNP set 

```{r}
fst_out_full.mat <- pairwise.WCfst(hf_out_full.filt)
```

```{r}
# get bootstrap confidence values for Fst
boot_out_full_fst <- boot.ppfst(hf_out_full.filt,nboot = 1000)
boot5_out_full_fst <- boot.ppfst(hf_out_full.filt,nboot = 1000,quant = 0.5)
```

```{r}
Fst_out_full_ll <- boot5_out_full_fst$ll
```

```{r}
gindF_out_full.fst.mat.triN <- as.matrix(fst_out_full.mat)
colnames(gindF_out_full.fst.mat.triN) <- pop_order_6pops
rownames(gindF_out_full.fst.mat.triN) <- pop_order_6pops
```

```{r}
meltedN_out_full <- reshape2::melt(gindF_out_full.fst.mat.triN, na.rm =TRUE)
round(gindF_out_full.fst.mat.triN,4)
```

```{r}
summary(meltedN_out_full$value)
```

```{r}
#Plotting Pairwise fst
out_full <- ggplot(data = meltedN_out_full, aes(Var2, Var1, fill = value))+ geom_tile(color = "white")+ 
  scale_fill_gradient(low = "white", high = "blue1", name="FST")  +
  ggtitle(expression(atop("Pairwise FST, WC (1984) full outlier SNPs for 6 Populations", atop(italic("N = 60, L = 70"), ""))))+
  labs( x = "Sampling Site", y = "Sampling Site") + 
  theme(axis.text.x = element_text(angle = 45, vjust = 1, size = 12, hjust = 1),axis.text.y = element_text(size = 12)) + 
  theme(axis.title = element_text(size = 12),legend.text = element_text(size =14), legend.title = element_text(size =14)) +
  theme(plot.title = element_text(size = 14)) +
coord_fixed()
out_full
```


pcadapt (db) SNP set 

```{r}
fst_out_db.mat <- pairwise.WCfst(hf_out_db.filt)
```

```{r}
# get bootstrap confidence values for Fst
boot_out_db_fst <- boot.ppfst(hf_out_db.filt,nboot = 1000)
boot5_out_db_fst <- boot.ppfst(hf_out_db.filt,nboot = 1000,quant = 0.5)
```

```{r}
Fst_out_db_ll <- boot5_out_db_fst$ll
```

```{r}
gindF_out_db.fst.mat.triN <- as.matrix(fst_out_db.mat)
colnames(gindF_out_db.fst.mat.triN) <- pop_order_6pops
rownames(gindF_out_db.fst.mat.triN) <- pop_order_6pops
```

```{r}
meltedN_out_db <- melt(gindF_out_db.fst.mat.triN, na.rm =TRUE)
round(gindF_out_db.fst.mat.triN,4)
```

```{r}
summary(meltedN_out_db$value)
```

```{r}
#Plotting Pairwise fst
out_db <- ggplot(data = meltedN_out_db, aes(Var2, Var1, fill = value))+ geom_tile(color = "white")+ 
  scale_fill_gradient(low = "white", high = "blue1", name="FST")  +
  ggtitle(expression(atop("Pairwise FST, WC (1984) db outlier SNPs for 6 Populations", atop(italic("N = 60, L = 58"), ""))))+
  labs( x = "Sampling Site", y = "Sampling Site") + 
  theme(axis.text.x = element_text(angle = 45, vjust = 1, size = 12, hjust = 1),axis.text.y = element_text(size = 12)) + 
  theme(axis.title = element_text(size = 12),legend.text = element_text(size =14), legend.title = element_text(size =14)) +
  theme(plot.title = element_text(size = 14)) +
coord_fixed()
out_db
```


pcadapt inside inversions (db_invert) SNP set 

```{r}
fst_out_db_invert.mat <- pairwise.WCfst(hf_out_db_invert.filt)
```

```{r}
# get bootstrap confidence values for Fst
boot_out_db_invert_fst <- boot.ppfst(hf_out_db_invert.filt,nboot = 1000)
boot5_out_db_invert_fst <- boot.ppfst(hf_out_db_invert.filt,nboot = 1000,quant = 0.5)
```

```{r}
Fst_out_db_invert_ll <- boot5_out_db_invert_fst$ll
```

```{r}
gindF_out_db_invert.fst.mat.triN <- as.matrix(fst_out_db_invert.mat)
colnames(gindF_out_db_invert.fst.mat.triN) <- pop_order_6pops
rownames(gindF_out_db_invert.fst.mat.triN) <- pop_order_6pops
```

```{r}
meltedN_out_db_invert <- melt(gindF_out_db_invert.fst.mat.triN, na.rm =TRUE)
round(gindF_out_db_invert.fst.mat.triN,4)
```

```{r}
summary(meltedN_out_db_invert$value)
```

```{r}
#Plotting Pairwise fst
out_db_invert <- ggplot(data = meltedN_out_db_invert, aes(Var2, Var1, fill = value))+ geom_tile(color = "white")+ 
  scale_fill_gradient(low = "white", high = "blue1", name="FST")  +
  ggtitle(expression(atop("Pairwise FST, WC (1984) db_invert outlier SNPs for 6 Populations", atop(italic("N = 60, L = 36"), ""))))+
  labs( x = "Sampling Site", y = "Sampling Site") + 
  theme(axis.text.x = element_text(angle = 45, vjust = 1, size = 12, hjust = 1),axis.text.y = element_text(size = 12)) + 
  theme(axis.title = element_text(size = 12),legend.text = element_text(size =14), legend.title = element_text(size =14)) +
  theme(plot.title = element_text(size = 14)) +
coord_fixed()
out_db_invert
```


pcadapt outside inversions (db_noinvert) SNP set 

```{r}
fst_out_db_noinvert.mat <- pairwise.WCfst(hf_out_db_noinvert.filt)
```

```{r}
# get bootstrap confidence values for Fst
boot_out_db_noinvert_fst <- boot.ppfst(hf_out_db_noinvert.filt,nboot = 1000)
boot5_out_db_noinvert_fst <- boot.ppfst(hf_out_db_noinvert.filt,nboot = 1000,quant = 0.5)
```

```{r}
Fst_out_db_noinvert_ll <- boot5_out_db_noinvert_fst$ll
```

```{r}
gindF_out_db_noinvert.fst.mat.triN <- as.matrix(fst_out_db_noinvert.mat)
colnames(gindF_out_db_noinvert.fst.mat.triN) <- pop_order_6pops
rownames(gindF_out_db_noinvert.fst.mat.triN) <- pop_order_6pops
```

```{r}
meltedN_out_db_noinvert <- melt(gindF_out_db_noinvert.fst.mat.triN, na.rm =TRUE)
round(gindF_out_db_noinvert.fst.mat.triN,4)
```

```{r}
summary(meltedN_out_db_noinvert$value)
```

```{r}
#Plotting Pairwise fst
out_db_noinvert <- ggplot(data = meltedN_out_db_noinvert, aes(Var2, Var1, fill = value))+ geom_tile(color = "white")+ 
  scale_fill_gradient(low = "white", high = "blue1", name="FST")  +
  ggtitle(expression(atop("Pairwise FST, WC (1984) db_noinvert outlier SNPs for 6 Populations", atop(italic("N = 60, L = 22"), ""))))+
  labs( x = "Sampling Site", y = "Sampling Site") + 
  theme(axis.text.x = element_text(angle = 45, vjust = 1, size = 12, hjust = 1),axis.text.y = element_text(size = 12)) + 
  theme(axis.title = element_text(size = 12),legend.text = element_text(size =14), legend.title = element_text(size =14)) +
  theme(plot.title = element_text(size = 14)) +
coord_fixed()
out_db_noinvert
```


LFMM (lfmm) SNP set 

```{r}
fst_out_lfmm.mat <- pairwise.WCfst(hf_out_lfmm.filt)
```

```{r}
# get bootstrap confidence values for Fst
boot_out_lfmm_fst <- boot.ppfst(hf_out_lfmm.filt,nboot = 1000)
boot5_out_lfmm_fst <- boot.ppfst(hf_out_lfmm.filt,nboot = 1000,quant = 0.5)
```

```{r}
Fst_out_lfmm_ll <- boot5_out_lfmm_fst$ll
```

```{r}
gindF_out_lfmm.fst.mat.triN <- as.matrix(fst_out_lfmm.mat)
colnames(gindF_out_lfmm.fst.mat.triN) <- pop_order_6pops
rownames(gindF_out_lfmm.fst.mat.triN) <- pop_order_6pops
```

```{r}
meltedN_out_lfmm <- melt(gindF_out_lfmm.fst.mat.triN, na.rm =TRUE)
round(gindF_out_lfmm.fst.mat.triN,4)
```

```{r}
summary(meltedN_out_lfmm$value)
```

```{r}
#Plotting Pairwise fst
out_lfmm <- ggplot(data = meltedN_out_lfmm, aes(Var2, Var1, fill = value))+ geom_tile(color = "white")+ 
  scale_fill_gradient(low = "white", high = "blue1", name="FST")  +
  ggtitle(expression(atop("Pairwise FST, WC (1984) lfmm outlier SNPs for 6 Populations", atop(italic("N = 60, L = 14"), ""))))+
  labs( x = "Sampling Site", y = "Sampling Site") + 
  theme(axis.text.x = element_text(angle = 45, vjust = 1, size = 12, hjust = 1),axis.text.y = element_text(size = 12)) + 
  theme(axis.title = element_text(size = 12),legend.text = element_text(size =14), legend.title = element_text(size =14)) +
  theme(plot.title = element_text(size = 14)) +
coord_fixed()
out_lfmm
```


LFMM inside inversions (lfmm_invert) SNP set 

```{r}
fst_out_lfmm_invert.mat <- pairwise.WCfst(hf_out_lfmm_invert.filt)
```

```{r}
# get bootstrap confidence values for Fst
boot_out_lfmm_invert_fst <- boot.ppfst(hf_out_lfmm_invert.filt,nboot = 1000)
boot5_out_lfmm_invert_fst <- boot.ppfst(hf_out_lfmm_invert.filt,nboot = 1000,quant = 0.5)
```

```{r}
Fst_out_lfmm_invert_ll <- boot5_out_lfmm_invert_fst$ll
```

```{r}
gindF_out_lfmm_invert.fst.mat.triN <- as.matrix(fst_out_lfmm_invert.mat)
colnames(gindF_out_lfmm_invert.fst.mat.triN) <- pop_order_6pops
rownames(gindF_out_lfmm_invert.fst.mat.triN) <- pop_order_6pops
```

```{r}
meltedN_out_lfmm_invert <- melt(gindF_out_lfmm_invert.fst.mat.triN, na.rm =TRUE)
round(gindF_out_lfmm_invert.fst.mat.triN,4)
```

```{r}
summary(meltedN_out_lfmm_invert$value)
```

```{r}
#Plotting Pairwise fst
out_lfmm_invert <- ggplot(data = meltedN_out_lfmm_invert, aes(Var2, Var1, fill = value))+ geom_tile(color = "white")+ 
  scale_fill_gradient(low = "white", high = "blue1", name="FST")  +
  ggtitle(expression(atop("Pairwise FST, WC (1984) lfmm_invert outlier SNPs for 6 Populations", atop(italic("N = 60, L = 6"), ""))))+
  labs( x = "Sampling Site", y = "Sampling Site") + 
  theme(axis.text.x = element_text(angle = 45, vjust = 1, size = 12, hjust = 1),axis.text.y = element_text(size = 12)) + 
  theme(axis.title = element_text(size = 12),legend.text = element_text(size =14), legend.title = element_text(size =14)) +
  theme(plot.title = element_text(size = 14)) +
coord_fixed()
out_lfmm_invert
```


LFMM outside inversions (lfmm_noinvert) SNP set 

```{r}
fst_out_lfmm_noinvert.mat <- pairwise.WCfst(hf_out_lfmm_noinvert.filt)
```

```{r}
# get bootstrap confidence values for Fst
boot_out_lfmm_noinvert_fst <- boot.ppfst(hf_out_lfmm_noinvert.filt,nboot = 1000)
boot5_out_lfmm_noinvert_fst <- boot.ppfst(hf_out_lfmm_noinvert.filt,nboot = 1000,quant = 0.5)
```

```{r}
Fst_out_lfmm_noinvert_ll <- boot5_out_lfmm_noinvert_fst$ll
```

```{r}
gindF_out_lfmm_noinvert.fst.mat.triN <- as.matrix(fst_out_lfmm_noinvert.mat)
colnames(gindF_out_lfmm_noinvert.fst.mat.triN) <- pop_order_6pops
rownames(gindF_out_lfmm_noinvert.fst.mat.triN) <- pop_order_6pops
```

```{r}
meltedN_out_lfmm_noinvert <- melt(gindF_out_lfmm_noinvert.fst.mat.triN, na.rm =TRUE)
round(gindF_out_lfmm_noinvert.fst.mat.triN,4)
```

```{r}
summary(meltedN_out_lfmm_noinvert$value)
```

```{r}
#Plotting Pairwise fst
out_lfmm_noinvert <- ggplot(data = meltedN_out_lfmm_noinvert, aes(Var2, Var1, fill = value))+ geom_tile(color = "white")+ 
  scale_fill_gradient(low = "white", high = "blue1", name="FST")  +
  ggtitle(expression(atop("Pairwise FST, WC (1984) lfmm_noinvert outlier SNPs for 6 Populations", atop(italic("N = 60, L = 8"), ""))))+
  labs( x = "Sampling Site", y = "Sampling Site") + 
  theme(axis.text.x = element_text(angle = 45, vjust = 1, size = 12, hjust = 1),axis.text.y = element_text(size = 12)) + 
  theme(axis.title = element_text(size = 12),legend.text = element_text(size =14), legend.title = element_text(size =14)) +
  theme(plot.title = element_text(size = 14)) +
coord_fixed()
out_lfmm_noinvert
```

RDA (rda) SNP set 

```{r}
fst_out_rda.mat <- pairwise.WCfst(hf_out_rda.filt)
```

```{r}
# get bootstrap confidence values for Fst
boot_out_rda_fst <- boot.ppfst(hf_out_rda.filt,nboot = 1000)
boot5_out_rda_fst <- boot.ppfst(hf_out_rda.filt,nboot = 1000,quant = 0.5)
```

```{r}
Fst_out_rda_ll <- boot5_out_rda_fst$ll
```

```{r}
gindF_out_rda.fst.mat.triN <- as.matrix(fst_out_rda.mat)
colnames(gindF_out_rda.fst.mat.triN) <- pop_order_6pops
rownames(gindF_out_rda.fst.mat.triN) <- pop_order_6pops
```

```{r}
meltedN_out_rda <- reshape2::melt(gindF_out_rda.fst.mat.triN, na.rm =TRUE)
round(gindF_out_rda.fst.mat.triN,4)
```

```{r}
summary(meltedN_out_rda$value)
```

```{r}
#Plotting Pairwise fst
out_rda <- ggplot(data = meltedN_out_rda, aes(Var2, Var1, fill = value))+ geom_tile(color = "white")+ 
  scale_fill_gradient(low = "white", high = "blue1", name="FST")  +
  ggtitle(expression(atop("Pairwise FST, WC (1984) rda outlier SNPs for 6 Populations", atop(italic("N = 60, L = 320"), ""))))+
  labs( x = "Sampling Site", y = "Sampling Site") + 
  theme(axis.text.x = element_text(angle = 45, vjust = 1, size = 12, hjust = 1),axis.text.y = element_text(size = 12)) + 
  theme(axis.title = element_text(size = 12),legend.text = element_text(size =14), legend.title = element_text(size =14)) +
  theme(plot.title = element_text(size = 14)) +
coord_fixed()
out_rda
```

full_new (full_new) SNP set 

```{r}
fst_out_full_new.mat <- pairwise.WCfst(hf_out_full_new.filt)
```

```{r}
# get bootstrap confidence values for Fst
boot_out_full_new_fst <- boot.ppfst(hf_out_full_new.filt,nboot = 1000)
boot5_out_full_new_fst <- boot.ppfst(hf_out_full_new.filt,nboot = 1000,quant = 0.5)
```

```{r}
Fst_out_full_new_ll <- boot5_out_full_new_fst$ll
```

```{r}
gindF_out_full_new.fst.mat.triN <- as.matrix(fst_out_full_new.mat)
colnames(gindF_out_full_new.fst.mat.triN) <- pop_order_6pops
rownames(gindF_out_full_new.fst.mat.triN) <- pop_order_6pops
```

```{r}
meltedN_out_full_new <- reshape2::melt(gindF_out_full_new.fst.mat.triN, na.rm =TRUE)
round(gindF_out_full_new.fst.mat.triN,4)
```

```{r}
summary(meltedN_out_full_new$value)
```

```{r}
#Plotting Pairwise fst
out_full_new <- ggplot(data = meltedN_out_full_new, aes(Var2, Var1, fill = value))+ geom_tile(color = "white")+ 
  scale_fill_gradient(low = "white", high = "blue1", name="FST")  +
  ggtitle(expression(atop("Pairwise FST, WC (1984) full_new outlier SNPs for 6 Populations", atop(italic("N = 60, L = 361"), ""))))+
  labs( x = "Sampling Site", y = "Sampling Site") + 
  theme(axis.text.x = element_text(angle = 45, vjust = 1, size = 12, hjust = 1),axis.text.y = element_text(size = 12)) + 
  theme(axis.title = element_text(size = 12),legend.text = element_text(size =14), legend.title = element_text(size =14)) +
  theme(plot.title = element_text(size = 14)) +
coord_fixed()
out_full_new
```

full_final (full_final) SNP set 

```{r}
fst_out_full_final.mat <- pairwise.WCfst(hf_out_full_final.filt)
```

```{r}
# get bootstrap confidence values for Fst
boot_out_full_final_fst <- boot.ppfst(hf_out_full_final.filt,nboot = 1000)
boot5_out_full_final_fst <- boot.ppfst(hf_out_full_final.filt,nboot = 1000,quant = 0.5)
```

```{r}
Fst_out_full_final_ll <- boot5_out_full_final_fst$ll
```

```{r}
gindF_out_full_final.fst.mat.triN <- as.matrix(fst_out_full_final.mat)
colnames(gindF_out_full_final.fst.mat.triN) <- pop_order_6pops
rownames(gindF_out_full_final.fst.mat.triN) <- pop_order_6pops
```

```{r}
meltedN_out_full_final <- reshape2::melt(gindF_out_full_final.fst.mat.triN, na.rm =TRUE)
round(gindF_out_full_final.fst.mat.triN,4)
```

```{r}
summary(meltedN_out_full_final$value)
```

```{r}
#Plotting Pairwise fst
out_full_final <- ggplot(data = meltedN_out_full_final, aes(Var2, Var1, fill = value))+ geom_tile(color = "white")+ 
  scale_fill_gradient(low = "white", high = "blue1", name="FST")  +
  ggtitle(expression(atop("Pairwise FST, WC (1984) full_final outlier SNPs for 6 Populations", atop(italic("N = 60, L = 390"), ""))))+
  labs( x = "Sampling Site", y = "Sampling Site") + 
  theme(axis.text.x = element_text(angle = 45, vjust = 1, size = 12, hjust = 1),axis.text.y = element_text(size = 12)) + 
  theme(axis.title = element_text(size = 12),legend.text = element_text(size =14), legend.title = element_text(size =14)) +
  theme(plot.title = element_text(size = 14)) +
coord_fixed()
out_full_final
```


## PCA 

```{r}
library("PCAviz")  #Visualizing output of PCA
library("cowplot") #Used with PCAviz
```

```{r}
# Create color scheme
# green=#7FC97F, blue = #00008B, orange=#FF7F00, purple=#9A32CD, pink=#FF1493, cyan=#00FFFF
cols_6pops = c("#7FC97F","#00008B","#FF7F00","#9A32CD","#FF1493","#00FFFF")
```

Full SNP set

```{r}
# Plot a PCA for the all outlier loci
x_out_full <- scaleGen(rad_out_full.filt, NA.method = "mean")
out_full_pca <- dudi.pca(x_out_full, center = FALSE, scale = FALSE, scannf = FALSE, nf = 4)
eig_percent_out_full <- round((out_full_pca$eig / (sum(out_full_pca$eig))) * 100, 2)

pca_dat_out_full <- as.data.frame(out_full_pca$li) %>%
                    rownames_to_column(var = "ind") %>%
                    cbind(pop = pop(rad_out_full.filt)) %>%
                    mutate(pop = factor(pop, levels = c("BAR", "BIS", "GB", "KIC", "MCD", "PVD")))


ggplot(data = pca_dat_out_full) +
    geom_point(aes(x = Axis1, y = Axis2, colour = pop), size = 3, alpha = 0.8) +
    scale_colour_manual(values = cols_6pops, name = "Sample site") +
    theme_minimal() +
    labs(x = paste("", "Axis 1: ", eig_percent_out_full[1], "%"), 
         y = paste("", "Axis 2: ", eig_percent_out_full[2], "%"),
         colour = "Population") +
    coord_fixed(ratio = 1) +
    theme(legend.position = "none")

ggsave("outlier_pca_full70.png", width = 7, height = 3, units = "in", bg = "white")
```

```{r}
# Look at the loadings on each PC1
pca_loadings_out_full_pc1 <- out_full_pca$co %>%
  as_tibble(rownames = "allele") %>%
  extract(allele, c("locus","chrom", "pos", "allele"), "(NC_(\\d+)_1_(\\d+)).(\\d)") %>%
  mutate(pos = as.integer(pos)) %>%
  dplyr::select(locus, chrom, pos, allele, Comp1, Comp2) %>%
  mutate(Comp1 = Comp1^2, Comp2 = Comp2^2) %>%
  arrange(dplyr::desc(Comp1))

# Look at the loadings on each PC2
pca_loadings_out_full_pc2 <- out_full_pca$co %>%
  as_tibble(rownames = "allele") %>%
  extract(allele, c("locus","chrom", "pos", "allele"), "(NC_(\\d+)_1_(\\d+)).(\\d)") %>%
  mutate(pos = as.integer(pos)) %>%
  dplyr::select(locus, chrom, pos, allele, Comp1, Comp2) %>%
  mutate(Comp1 = Comp1^2, Comp2 = Comp2^2) %>%
  arrange(dplyr::desc(Comp2))
```

pcadapt (db) SNP set

```{r}
# Plot a PCA for the db outlier loci
x_out_db <- scaleGen(rad_out_db.filt, NA.method = "mean")
out_db_pca <- dudi.pca(x_out_db, center = FALSE, scale = FALSE, scannf = FALSE, nf = 4)
eig_percent_out_db <- round((out_db_pca$eig / (sum(out_db_pca$eig))) * 100, 2)

pca_dat_out_db <- as.data.frame(out_db_pca$li) %>%
                    rownames_to_column(var = "ind") %>%
                    cbind(pop = pop(rad_out_db.filt)) %>%
                    mutate(pop = factor(pop, levels = c("BAR", "BIS", "GB", "KIC", "MCD", "PVD")))


ggplot(data = pca_dat_out_db) +
    geom_point(aes(x = Axis1, y = Axis2, colour = pop), size = 3, alpha = 0.8) +
    scale_colour_manual(values = cols_6pops) +
    theme_minimal() +
    labs(x = paste("", "Axis 1: ", eig_percent_out_db[1], "%"), y = paste("", "Axis 2: ", eig_percent_out_db[2], "%")) +
    coord_fixed(ratio = 1)
```

pcadapt inside inversions (db_invert) SNP set

```{r}
# Plot a PCA for the db_invert outlier loci
x_out_db_invert <- scaleGen(rad_out_db_invert.filt, NA.method = "mean")
out_db_invert_pca <- dudi.pca(x_out_db_invert, center = FALSE, scale = FALSE, scannf = FALSE, nf = 4)
eig_percent_out_db_invert <- round((out_db_invert_pca$eig / (sum(out_db_invert_pca$eig))) * 100, 2)

pca_dat_out_db_invert <- as.data.frame(out_db_invert_pca$li) %>%
                    rownames_to_column(var = "ind") %>%
                    cbind(pop = pop(rad_out_db_invert.filt)) %>%
                    mutate(pop = factor(pop, levels = c("BAR", "BIS", "GB", "KIC", "MCD", "PVD")))


ggplot(data = pca_dat_out_db_invert) +
    geom_point(aes(x = Axis1, y = Axis2, colour = pop), size = 3, alpha = 0.8) +
    scale_colour_manual(values = cols_6pops) +
    theme_minimal() +
    labs(x = paste("", "Axis 1: ", eig_percent_out_db_invert[1], "%"), y = paste("", "Axis 2: ", eig_percent_out_db_invert[2], "%")) +
    coord_fixed(ratio = 1)
```

pcadapt outside inversions (db_noinvert) SNP set

```{r}
# Plot a PCA for the db_invert outlier loci
x_out_db_noinvert <- scaleGen(rad_out_db_noinvert.filt, NA.method = "mean")
out_db_noinvert_pca <- dudi.pca(x_out_db_noinvert, center = FALSE, scale = FALSE, scannf = FALSE, nf = 4)
eig_percent_out_db_noinvert <- round((out_db_noinvert_pca$eig / (sum(out_db_noinvert_pca$eig))) * 100, 2)

pca_dat_out_db_noinvert <- as.data.frame(out_db_noinvert_pca$li) %>%
                    rownames_to_column(var = "ind") %>%
                    cbind(pop = pop(rad_out_db_noinvert.filt)) %>%
                    mutate(pop = factor(pop, levels = c("BAR", "BIS", "GB", "KIC", "MCD", "PVD")))


ggplot(data = pca_dat_out_db_noinvert) +
    geom_point(aes(x = Axis1, y = Axis2, colour = pop), size = 3, alpha = 0.8) +
    scale_colour_manual(values = cols_6pops) +
    theme_minimal() +
    labs(x = paste("", "Axis 1: ", eig_percent_out_db_noinvert[1], "%"), y = paste("", "Axis 2: ", eig_percent_out_db_noinvert[2], "%")) +
    coord_fixed(ratio = 1)
```

LFMM outliers

```{r}
# Plot a PCA for the lfmm outlier loci
x_out_lfmm <- scaleGen(rad_out_lfmm.filt, NA.method = "mean")
out_lfmm_pca <- dudi.pca(x_out_lfmm, center = FALSE, scale = FALSE, scannf = FALSE, nf = 4)
eig_percent_out_lfmm <- round((out_lfmm_pca$eig / (sum(out_lfmm_pca$eig))) * 100, 2)

pca_dat_out_lfmm <- as.data.frame(out_lfmm_pca$li) %>%
                    rownames_to_column(var = "ind") %>%
                    cbind(pop = pop(rad_out_lfmm.filt)) %>%
                    mutate(pop = factor(pop, levels = c("BAR", "BIS", "GB", "KIC", "MCD", "PVD")))


ggplot(data = pca_dat_out_lfmm) +
    geom_point(aes(x = Axis1, y = Axis2, colour = pop), size = 3, alpha = 0.8) +
    scale_colour_manual(values = cols_6pops) +
    theme_minimal() +
    labs(x = paste("", "Axis 1: ", eig_percent_out_lfmm[1], "%"), y = paste("", "Axis 2: ", eig_percent_out_lfmm[2], "%")) +
    coord_fixed(ratio = 1)
```

RDA outliers

```{r}
# Plot a PCA for the rda outlier loci
x_out_rda <- scaleGen(rad_out_rda.filt, NA.method = "mean")
out_rda_pca <- dudi.pca(x_out_rda, center = FALSE, scale = FALSE, scannf = FALSE, nf = 4)
eig_percent_out_rda <- round((out_rda_pca$eig / (sum(out_rda_pca$eig))) * 100, 2)

pca_dat_out_rda <- as.data.frame(out_rda_pca$li) %>%
                    rownames_to_column(var = "ind") %>%
                    cbind(pop = pop(rad_out_rda.filt)) %>%
                    mutate(pop = factor(pop, levels = c("BAR", "BIS", "GB", "KIC", "MCD", "PVD")))


ggplot(data = pca_dat_out_rda) +
    geom_point(aes(x = Axis1, y = Axis2, colour = pop), size = 3, alpha = 0.8) +
    scale_colour_manual(values = cols_6pops) +
    theme_minimal() +
    labs(x = paste("", "Axis 1: ", eig_percent_out_rda[1], "%"), y = paste("", "Axis 2: ", eig_percent_out_rda[2], "%")) +
    coord_fixed(ratio = 1)
```

full_new outliers

```{r}
# Plot a PCA for the full_new outlier loci
x_out_full_new <- scaleGen(rad_out_full_new.filt, NA.method = "mean")
out_full_new_pca <- dudi.pca(x_out_full_new, center = FALSE, scale = FALSE, scannf = FALSE, nf = 4)
eig_percent_out_full_new <- round((out_full_new_pca$eig / (sum(out_full_new_pca$eig))) * 100, 2)

pca_dat_out_full_new <- as.data.frame(out_full_new_pca$li) %>%
                    rownames_to_column(var = "ind") %>%
                    cbind(pop = pop(rad_out_full_new.filt)) %>%
                    mutate(pop = factor(pop, levels = c("BAR", "BIS", "GB", "KIC", "MCD", "PVD")))


ggplot(data = pca_dat_out_full_new) +
    geom_point(aes(x = Axis1, y = Axis2, colour = pop), size = 3, alpha = 0.8) +
    scale_colour_manual(values = cols_6pops) +
    theme_minimal() +
    labs(x = paste("", "Axis 1: ", eig_percent_out_full_new[1], "%"), y = paste("", "Axis 2: ", eig_percent_out_full_new[2], "%")) +
    coord_fixed(ratio = 1)
```

full_final outliers

```{r}
# Plot a PCA for the full_final outlier loci
x_out_full_final <- scaleGen(rad_out_full_final.filt, NA.method = "mean")
out_full_final_pca <- dudi.pca(x_out_full_final, center = FALSE, scale = FALSE, scannf = FALSE, nf = 4)
eig_percent_out_full_final <- round((out_full_final_pca$eig / (sum(out_full_final_pca$eig))) * 100, 2)

pca_dat_out_full_final <- as.data.frame(out_full_final_pca$li) %>%
                    rownames_to_column(var = "ind") %>%
                    cbind(pop = pop(rad_out_full_final.filt)) %>%
                    mutate(pop = factor(pop, levels = c("BAR", "BIS", "GB", "KIC", "MCD", "PVD")))

ggplot(data = pca_dat_out_full_final) +
    geom_point(aes(x = Axis1, y = Axis2, colour = pop), size = 3, alpha = 0.8) +
    scale_colour_manual(values = cols_6pops, name = "Sample site") +
    theme_minimal() +
    labs(x = paste("", "Axis 1: ", eig_percent_out_full_final[1], "%"), 
         y = paste("", "Axis 2: ", eig_percent_out_full_final[2], "%"),
         colour = "Population") +
    coord_fixed(ratio = 1)
ggsave("outlier_pca_full390.png", width = 6, height = 4.6, units = "in", dpi = 300, bg = "white")
```

```{r}
# Look at the loadings on each PC1
pca_loadings_out_full_final_pc1 <- out_full_final_pca$co %>%
  as_tibble(rownames = "allele") %>%
  extract(allele, c("locus","chrom", "pos", "allele"), "(NC_(\\d+)_1_(\\d+)).(\\d)") %>%
  mutate(pos = as.integer(pos)) %>%
  dplyr::select(locus, chrom, pos, allele, Comp1, Comp2) %>%
  mutate(Comp1 = Comp1^2, Comp2 = Comp2^2) %>%
  arrange(dplyr::desc(Comp1))
print(pca_loadings_out_full_final_pc1)

# Look at the loadings on each PC2
pca_loadings_out_full_final_pc2 <- out_full_final_pca$co %>%
  as_tibble(rownames = "allele") %>%
  extract(allele, c("locus","chrom", "pos", "allele"), "(NC_(\\d+)_1_(\\d+)).(\\d)") %>%
  mutate(pos = as.integer(pos)) %>%
  dplyr::select(locus, chrom, pos, allele, Comp1, Comp2) %>%
  mutate(Comp1 = Comp1^2, Comp2 = Comp2^2) %>%
  arrange(dplyr::desc(Comp2))
print(pca_loadings_out_full_final_pc2)
```

## Genetic diversity (observed and expected heterozygosity)

Full SNP set

```{r}
comb_out_full <- summary(stratted_out_full.filt)
names(stratted_out_full.filt)
```

```{r}
plot(comb_out_full$Hobs, xlab="Loci number", ylab="Observed Heterozygosity", 
     main="Observed heterozygosity per locus")
```

```{r}
plot(comb_out_full$Hobs,comb_out_full$Hexp, xlab="Hobs", ylab="Hexp", 
     main="Expected heterozygosity as a function of observed heterozygosity per locus")
```

```{r}
bartlett.test(list(comb_out_full$Hexp, comb_out_full$Hobs)) # a test : H0: Hexp = Hobs
```

**_No Significant difference between Observed and expected heterzygosity_**.  

```{r}
basicstat_out_full <- basic.stats(hf_out_full.filt, diploid = TRUE, digits = 3)
```

```{r}
as.data.frame(basicstat_out_full$overall)
```

```{r}
# get bootstrap confidence values for Fis
boot_out_full <- boot.ppfis(hf_out_full.filt,nboot = 1000)
boot5_out_full <- boot.ppfis(hf_out_full.filt,nboot = 1000,quant = 0.5)
```

```{r}
# combine all pop statistics
colnames(basicstat_out_full$Ho) <- pop_order_6pops
Ho_out_full <- colMeans(basicstat_out_full$Ho,na.rm = T)
He_out_full <- colMeans(basicstat_out_full$Hs,na.rm = T)
Fis_out_full <- boot5_out_full$fis.ci$ll
y_out_full <- cbind(pop_order_6pops,Ho_out_full,He_out_full, Fis_out_full, strata_6pops[,3:17])
```

```{r}
summary(He_out_full)
```

```{r}
summary(Fis_out_full)
```

Plotting F statistics versus each environmental variables (including lat and long)

This can be modified for each F statistic (He, Ho, Fis) - just make sure to change the title in the function. 

```{r}
# Function to generate and display plots for all environmental variables
generate_plots <- function(data, response_var, env_vars, pop_order_var) {
  # Define color scheme
  color_scheme <- c("#7FC97F","#00008B","#FF7F00","#9A32CD","#FF1493","#00FFFF")
  
  for (env_var in env_vars) {
    # Ensure variable names are valid
    if (!all(c(response_var, env_var, pop_order_var) %in% colnames(data))) {
      cat("Skipping variable", env_var, "as it is not found in the data\n")
      next
    }
    
    # Remove rows with missing values in the relevant columns
    subset_data <- data[complete.cases(data[, c(response_var, env_var, pop_order_var)]), ]
    
    # Calculate R² value
    formula <- as.formula(paste(response_var, "~", env_var))
    R2_value <- round(summary(lm(formula, data = subset_data))$r.squared, 8)
    
    # Generate plot title
    plot_title <- paste("Observed Heterozygosity vs", env_var, ", outlier SNPs 6 pops")
    
    # Generate plot
    p <- ggplot(subset_data, aes_string(x = env_var, y = response_var)) + 
      geom_point(aes_string(color = pop_order_var), shape = 16, size = 3) + 
      scale_color_manual(values = color_scheme) +
      geom_smooth(method = "lm", color = "black") + 
      ggtitle(plot_title) +
      annotate(geom = "text", label = paste("italic(R^2)==", R2_value), x = Inf, y = Inf, hjust = 1.1, vjust = 2, parse = TRUE) +
      scale_x_continuous() +
      theme_bw() +
      theme(legend.title = element_blank(),
            plot.title = element_text(hjust = 0.5))
    
    # Print the plot
    print(p)
    
    cat("Generated plot for", env_var, "\n")
  }
}
```

```{r}
generate_plots(y_out_full, "Ho_out_full", colnames(y_out_full)[5:19], "pop_order_6pops")
```

```{r}
generate_plots(y_out_full, "He_out_full", colnames(y_out_full)[5:19], "pop_order_6pops")
```


pcadapt (db) SNP set

```{r}
comb_out_db <- summary(stratted_out_db.filt)
names(stratted_out_db.filt)
```

```{r}
plot(comb_out_db$Hobs, xlab="Loci number", ylab="Observed Heterozygosity", 
     main="Observed heterozygosity per locus")
```

```{r}
plot(comb_out_db$Hobs,comb_out_db$Hexp, xlab="Hobs", ylab="Hexp", 
     main="Expected heterozygosity as a function of observed heterozygosity per locus")
```

```{r}
bartlett.test(list(comb_out_db$Hexp, comb_out_db$Hobs)) # a test : H0: Hexp = Hobs
```

**_No Significant difference between Observed and expected heterozygosity_**.  

```{r}
basicstat_out_db <- basic.stats(hf_out_db.filt, diploid = TRUE, digits = 3)
```

```{r}
as.data.frame(basicstat_out_db$overall)
```

```{r}
# get bootstrap confidence values for Fis
boot_out_db <- boot.ppfis(hf_out_db.filt,nboot = 1000)
boot5_out_db <- boot.ppfis(hf_out_db.filt,nboot = 1000,quant = 0.5)
```

```{r}
# combine all pop statistics
colnames(basicstat_out_db$Ho) <- pop_order_6pops
Ho_out_db <- colMeans(basicstat_out_db$Ho,na.rm = T)
He_out_db <- colMeans(basicstat_out_db$Hs,na.rm = T)
Fis_out_db <- boot5_out_db$fis.ci$ll
y_out_db <- cbind(pop_order_6pops,Ho_out_db,He_out_db, Fis_out_db, strata_6pops[,3:17])
```

```{r}
summary(He_out_db)
```

```{r}
summary(Fis_out_db)
```

Plotting F statistics versus each environmental variables (including lat and long)

This can be modified for each F statistic (He, Ho, Fis) - just make sure to change the title in the function. 

```{r}
generate_plots(y_out_db, "Ho_out_db", colnames(y_out_db)[5:19], "pop_order_6pops")
```

```{r}
generate_plots(y_out_db, "He_out_db", colnames(y_out_db)[5:19], "pop_order_6pops")
```

 pcadapt inside inversion (db_invert) SNP set

```{r}
comb_out_db_invert <- summary(stratted_out_db_invert.filt)
names(stratted_out_db_invert.filt)
```

```{r}
plot(comb_out_db_invert$Hobs, xlab="Loci number", ylab="Observed Heterozygosity", 
     main="Observed heterozygosity per locus")
```

```{r}
plot(comb_out_db_invert$Hobs,comb_out_db_invert$Hexp, xlab="Hobs", ylab="Hexp", 
     main="Expected heterozygosity as a function of observed heterozygosity per locus")
```

```{r}
bartlett.test(list(comb_out_db_invert$Hexp, comb_out_db_invert$Hobs)) # a test : H0: Hexp = Hobs
```

**_No Significant difference between Observed and expected heterzygosity_**.  

```{r}
basicstat_out_db_invert <- basic.stats(hf_out_db_invert.filt, diploid = TRUE, digits = 3)
```

```{r}
as.data.frame(basicstat_out_db_invert$overall)
```

```{r}
# get bootstrap confidence values for Fis
boot_out_db_invert <- boot.ppfis(hf_out_db_invert.filt,nboot = 1000)
boot5_out_db_invert <- boot.ppfis(hf_out_db_invert.filt,nboot = 1000,quant = 0.5)
```

```{r}
# combine all pop statistics
colnames(basicstat_out_db_invert$Ho) <- pop_order_6pops
Ho_out_db_invert <- colMeans(basicstat_out_db_invert$Ho,na.rm = T)
He_out_db_invert <- colMeans(basicstat_out_db_invert$Hs,na.rm = T)
Fis_out_db_invert <- boot5_out_db_invert$fis.ci$ll
y_out_db_invert <- cbind(pop_order_6pops,Ho_out_db_invert,He_out_db_invert, Fis_out_db_invert, strata_6pops[,3:17])
```

```{r}
summary(He_out_db_invert)
```

```{r}
summary(Fis_out_db_invert)
```

Plotting F statistics versus each environmental variables (including lat and long)

This can be modified for each F statistic (He, Ho, Fis) - just make sure to change the title in the function. 

```{r}
generate_plots(y_out_db_invert, "Ho_out_db_invert", colnames(y_out_db_invert)[5:19], "pop_order_6pops")
```

```{r}
generate_plots(y_out_db_invert, "He_out_db_invert", colnames(y_out_db_invert)[5:19], "pop_order_6pops")
```

pcadapt outside inversions (db_noinvert) SNP set

```{r}
comb_out_db_noinvert <- summary(stratted_out_db_noinvert.filt)
names(stratted_out_db_noinvert.filt)
```

```{r}
plot(comb_out_db_noinvert$Hobs, xlab="Loci number", ylab="Observed Heterozygosity", 
     main="Observed heterozygosity per locus")
```

```{r}
plot(comb_out_db_noinvert$Hobs,comb_out_db_noinvert$Hexp, xlab="Hobs", ylab="Hexp", 
     main="Expected heterozygosity as a function of observed heterozygosity per locus")
```

```{r}
bartlett.test(list(comb_out_db_noinvert$Hexp, comb_out_db_noinvert$Hobs)) # a test : H0: Hexp = Hobs
```

**_No Significant difference between Observed and expected heterzygosity_**.  

```{r}
basicstat_out_db_noinvert <- basic.stats(hf_out_db_noinvert.filt, diploid = TRUE, digits = 3)
```

```{r}
as.data.frame(basicstat_out_db_noinvert$overall)
```

```{r}
# get bootstrap confidence values for Fis
boot_out_db_noinvert <- boot.ppfis(hf_out_db_noinvert.filt,nboot = 1000)
boot5_out_db_noinvert <- boot.ppfis(hf_out_db_noinvert.filt,nboot = 1000,quant = 0.5)
```

```{r}
# combine all pop statistics
colnames(basicstat_out_db_noinvert$Ho) <- pop_order_6pops
Ho_out_db_noinvert <- colMeans(basicstat_out_db_noinvert$Ho,na.rm = T)
He_out_db_noinvert <- colMeans(basicstat_out_db_noinvert$Hs,na.rm = T)
Fis_out_db_noinvert <- boot5_out_db_noinvert$fis.ci$ll
y_out_db_noinvert <- cbind(pop_order_6pops,Ho_out_db_noinvert,He_out_db_noinvert, Fis_out_db_noinvert, strata_6pops[,3:17])
```

```{r}
summary(He_out_db_noinvert)
```

```{r}
summary(Fis_out_db_noinvert)
```

Plotting F statistics versus each environmental variables (including lat and long)

This can be modified for each F statistic (He, Ho, Fis) - just make sure to change the title in the function. 

```{r}
generate_plots(y_out_db_noinvert, "Ho_out_db_noinvert", colnames(y_out_db_noinvert)[5:19], "pop_order_6pops")
```

```{r}
generate_plots(y_out_db_noinvert, "He_out_db_noinvert", colnames(y_out_db_noinvert)[5:19], "pop_order_6pops")
```

LFMM (lfmm) SNP set

```{r}
comb_out_lfmm <- summary(stratted_out_lfmm.filt)
names(stratted_out_lfmm.filt)
```

```{r}
plot(comb_out_lfmm$Hobs, xlab="Loci number", ylab="Observed Heterozygosity", 
     main="Observed heterozygosity per locus")
```

```{r}
plot(comb_out_lfmm$Hobs,comb_out_lfmm$Hexp, xlab="Hobs", ylab="Hexp", 
     main="Expected heterozygosity as a function of observed heterozygosity per locus")
```

```{r}
bartlett.test(list(comb_out_lfmm$Hexp, comb_out_lfmm$Hobs)) # a test : H0: Hexp = Hobs
```

**_No Significant difference between Observed and expected heterozygosity_**.  

```{r}
basicstat_out_lfmm <- basic.stats(hf_out_lfmm.filt, diploid = TRUE, digits = 3)
```

```{r}
as.data.frame(basicstat_out_lfmm$overall)
```

```{r}
# get bootstrap confidence values for Fis
boot_out_lfmm <- boot.ppfis(hf_out_lfmm.filt,nboot = 1000)
boot5_out_lfmm <- boot.ppfis(hf_out_lfmm.filt,nboot = 1000,quant = 0.5)
```

```{r}
# combine all pop statistics
colnames(basicstat_out_lfmm$Ho) <- pop_order_6pops
Ho_out_lfmm <- colMeans(basicstat_out_lfmm$Ho,na.rm = T)
He_out_lfmm <- colMeans(basicstat_out_lfmm$Hs,na.rm = T)
Fis_out_lfmm <- boot5_out_lfmm$fis.ci$ll
y_out_lfmm <- cbind(pop_order_6pops,Ho_out_lfmm,He_out_lfmm, Fis_out_lfmm, strata_6pops[,3:17])
```

```{r}
summary(He_out_lfmm)
```

```{r}
summary(Fis_out_lfmm)
```

Plotting F statistics versus each environmental variables (including lat and long)

This can be modified for each F statistic (He, Ho, Fis) - just make sure to change the title in the function. 

```{r}
# generate_plots(y_out_lfmm, "Ho_out_lfmm", colnames(y_out_lfmm)[5:18], "pop_order_6pops")
```


 LFMM inside inversion (lfmm_invert) SNP set

```{r}
comb_out_lfmm_invert <- summary(stratted_out_lfmm_invert.filt)
names(stratted_out_lfmm_invert.filt)
```

```{r}
plot(comb_out_lfmm_invert$Hobs, xlab="Loci number", ylab="Observed Heterozygosity", 
     main="Observed heterozygosity per locus")
```

```{r}
plot(comb_out_lfmm_invert$Hobs,comb_out_lfmm_invert$Hexp, xlab="Hobs", ylab="Hexp", 
     main="Expected heterozygosity as a function of observed heterozygosity per locus")
```

```{r}
bartlett.test(list(comb_out_lfmm_invert$Hexp, comb_out_lfmm_invert$Hobs)) # a test : H0: Hexp = Hobs
```

**_No Significant difference between Observed and expected heterozygosity_**.  

```{r}
basicstat_out_lfmm_invert <- basic.stats(hf_out_lfmm_invert.filt, diploid = TRUE, digits = 3)
```

```{r}
as.data.frame(basicstat_out_lfmm_invert$overall)
```

```{r}
# get bootstrap confidence values for Fis
boot_out_lfmm_invert <- boot.ppfis(hf_out_lfmm_invert.filt,nboot = 1000)
boot5_out_lfmm_invert <- boot.ppfis(hf_out_lfmm_invert.filt,nboot = 1000,quant = 0.5)
```

```{r}
# combine all pop statistics
colnames(basicstat_out_lfmm_invert$Ho) <- pop_order_6pops
Ho_out_lfmm_invert <- colMeans(basicstat_out_lfmm_invert$Ho,na.rm = T)
He_out_lfmm_invert <- colMeans(basicstat_out_lfmm_invert$Hs,na.rm = T)
Fis_out_lfmm_invert <- boot5_out_lfmm_invert$fis.ci$ll
y_out_lfmm_invert <- cbind(pop_order_6pops,Ho_out_lfmm_invert,He_out_lfmm_invert, Fis_out_lfmm_invert, strata_6pops[,3:17])
```

```{r}
summary(He_out_lfmm_invert)
```

```{r}
summary(Fis_out_lfmm_invert)
```

Plotting F statistics versus each environmental variables (including lat and long)

This can be modified for each F statistic (He, Ho, Fis) - just make sure to change the title in the function. 

```{r}
# generate_plots(y_out_lfmm_invert, "Ho_out_lfmm_invert", colnames(y_out_lfmm_invert)[5:18], "pop_order_6pops")
```


LFMM outside inversions (lfmm_noinvert) SNP set

```{r}
comb_out_lfmm_noinvert <- summary(stratted_out_lfmm_noinvert.filt)
names(stratted_out_lfmm_noinvert.filt)
```

```{r}
plot(comb_out_lfmm_noinvert$Hobs, xlab="Loci number", ylab="Observed Heterozygosity", 
     main="Observed heterozygosity per locus")
```

```{r}
plot(comb_out_lfmm_noinvert$Hobs,comb_out_lfmm_noinvert$Hexp, xlab="Hobs", ylab="Hexp", 
     main="Expected heterozygosity as a function of observed heterozygosity per locus")
```

```{r}
bartlett.test(list(comb_out_lfmm_noinvert$Hexp, comb_out_lfmm_noinvert$Hobs)) # a test : H0: Hexp = Hobs
```

**_No Significant difference between Observed and expected heterozygosity_**.  

```{r}
basicstat_out_lfmm_noinvert <- basic.stats(hf_out_lfmm_noinvert.filt, diploid = TRUE, digits = 3)
```

```{r}
as.data.frame(basicstat_out_lfmm_noinvert$overall)
```

```{r}
# get bootstrap confidence values for Fis
boot_out_lfmm_noinvert <- boot.ppfis(hf_out_lfmm_noinvert.filt,nboot = 1000)
boot5_out_lfmm_noinvert <- boot.ppfis(hf_out_lfmm_noinvert.filt,nboot = 1000,quant = 0.5)
```

```{r}
# combine all pop statistics
colnames(basicstat_out_lfmm_noinvert$Ho) <- pop_order_6pops
Ho_out_lfmm_noinvert <- colMeans(basicstat_out_lfmm_noinvert$Ho,na.rm = T)
He_out_lfmm_noinvert <- colMeans(basicstat_out_lfmm_noinvert$Hs,na.rm = T)
Fis_out_lfmm_noinvert <- boot5_out_lfmm_noinvert$fis.ci$ll
y_out_lfmm_noinvert <- cbind(pop_order_6pops,Ho_out_lfmm_noinvert,He_out_lfmm_noinvert, Fis_out_lfmm_noinvert, strata_6pops[,3:17])
```

```{r}
summary(He_out_lfmm_noinvert)
```

```{r}
summary(Fis_out_lfmm_noinvert)
```

Plotting F statistics versus each environmental variables (including lat and long)

This can be modified for each F statistic (He, Ho, Fis) - just make sure to change the title in the function. 

```{r}
# generate_plots(y_out_lfmm_noinvert, "Ho_out_lfmm_noinvert", colnames(y_out_lfmm_noinvert)[5:18], "pop_order_6pops")
```


RDA (rda) SNP set

```{r}
comb_out_rda <- summary(stratted_out_rda.filt)
names(stratted_out_rda.filt)
```

```{r}
plot(comb_out_rda$Hobs, xlab="Loci number", ylab="Observed Heterozygosity", 
     main="Observed heterozygosity per locus")
```

```{r}
plot(comb_out_rda$Hobs,comb_out_rda$Hexp, xlab="Hobs", ylab="Hexp", 
     main="Expected heterozygosity as a function of observed heterozygosity per locus")
```

```{r}
bartlett.test(list(comb_out_rda$Hexp, comb_out_rda$Hobs)) # a test : H0: Hexp = Hobs
```

**Significant difference between Observed and expected heterzygosity_**.  

```{r}
basicstat_out_rda <- basic.stats(hf_out_rda.filt, diploid = TRUE, digits = 3)
```

```{r}
as.data.frame(basicstat_out_rda$overall)
```

```{r}
# get bootstrap confidence values for Fis
boot_out_rda <- boot.ppfis(hf_out_rda.filt,nboot = 1000)
boot5_out_rda <- boot.ppfis(hf_out_rda.filt,nboot = 1000,quant = 0.5)
```

```{r}
# combine all pop statistics
colnames(basicstat_out_rda$Ho) <- pop_order_6pops
Ho_out_rda <- colMeans(basicstat_out_rda$Ho,na.rm = T)
He_out_rda <- colMeans(basicstat_out_rda$Hs,na.rm = T)
Fis_out_rda <- boot5_out_rda$fis.ci$ll
y_out_rda <- cbind(pop_order_6pops,Ho_out_rda,He_out_rda, Fis_out_rda, strata_6pops[,3:17])
```

```{r}
summary(He_out_rda)
```

```{r}
summary(Fis_out_rda)
```

Plotting F statistics versus each environmental variables (including lat and long)

This can be modified for each F statistic (He, Ho, Fis) - just make sure to change the title in the function. 

```{r}
# generate_plots(y_out_rda, "Ho_out_rda", colnames(y_out_rda)[5:18], "pop_order_6pops")
```

full_new (full_new) SNP set

```{r}
comb_out_full_new <- summary(stratted_out_full_new.filt)
names(stratted_out_full_new.filt)
```

```{r}
plot(comb_out_full_new$Hobs, xlab="Loci number", ylab="Observed Heterozygosity", 
     main="Observed heterozygosity per locus")
```

```{r}
plot(comb_out_full_new$Hobs,comb_out_full_new$Hexp, xlab="Hobs", ylab="Hexp", 
     main="Expected heterozygosity as a function of observed heterozygosity per locus")
```

```{r}
bartlett.test(list(comb_out_full_new$Hexp, comb_out_full_new$Hobs)) # a test : H0: Hexp = Hobs
```

**Significant difference between Observed and expected heterzygosity_**.  

```{r}
basicstat_out_full_new <- basic.stats(hf_out_full_new.filt, diploid = TRUE, digits = 3)
```

```{r}
as.data.frame(basicstat_out_full_new$overall)
```

```{r}
# get bootstrap confidence values for Fis
boot_out_full_new <- boot.ppfis(hf_out_full_new.filt,nboot = 1000)
boot5_out_full_new <- boot.ppfis(hf_out_full_new.filt,nboot = 1000,quant = 0.5)
```

```{r}
# combine all pop statistics
colnames(basicstat_out_full_new$Ho) <- pop_order_6pops
Ho_out_full_new <- colMeans(basicstat_out_full_new$Ho,na.rm = T)
He_out_full_new <- colMeans(basicstat_out_full_new$Hs,na.rm = T)
Fis_out_full_new <- boot5_out_full_new$fis.ci$ll
y_out_full_new <- cbind(pop_order_6pops,Ho_out_full_new,He_out_full_new, Fis_out_full_new, strata_6pops[,3:17])
```

```{r}
summary(He_out_full_new)
```

```{r}
summary(Fis_out_full_new)
```

Plotting F statistics versus each environmental variables (including lat and long)

This can be modified for each F statistic (He, Ho, Fis) - just make sure to change the title in the function. 

```{r}
# generate_plots(y_out_full_new, "Ho_out_full_new", colnames(y_out_full_new)[5:18], "pop_order_6pops")
```

full_final (full_final) SNP set

```{r}
comb_out_full_final <- summary(stratted_out_full_final.filt)
names(stratted_out_full_final.filt)
```

```{r}
plot(comb_out_full_final$Hobs, xlab="Loci number", ylab="Observed Heterozygosity", 
     main="Observed heterozygosity per locus")
```

```{r}
plot(comb_out_full_final$Hobs,comb_out_full_final$Hexp, xlab="Hobs", ylab="Hexp", 
     main="Expected heterozygosity as a function of observed heterozygosity per locus")
```

```{r}
bartlett.test(list(comb_out_full_final$Hexp, comb_out_full_final$Hobs)) # a test : H0: Hexp = Hobs
```

**Significant difference between Observed and expected heterzygosity_**.  

```{r}
basicstat_out_full_final <- basic.stats(hf_out_full_final.filt, diploid = TRUE, digits = 3)
```

```{r}
as.data.frame(basicstat_out_full_final$overall)
```

```{r}
# get bootstrap confidence values for Fis
boot_out_full_final <- boot.ppfis(hf_out_full_final.filt,nboot = 1000)
boot5_out_full_final <- boot.ppfis(hf_out_full_final.filt,nboot = 1000,quant = 0.5)
```

```{r}
# combine all pop statistics
colnames(basicstat_out_full_final$Ho) <- pop_order_6pops
Ho_out_full_final <- colMeans(basicstat_out_full_final$Ho,na.rm = T)
He_out_full_final <- colMeans(basicstat_out_full_final$Hs,na.rm = T)
Fis_out_full_final <- boot5_out_full_final$fis.ci$ll
y_out_full_final <- cbind(pop_order_6pops,Ho_out_full_final,He_out_full_final, Fis_out_full_final, strata_6pops[,3:17])
```

```{r}
summary(He_out_full_final)
```

```{r}
summary(Fis_out_full_final)
```


## Seascape Redundancy Analysis 

### Loading necessary packages 

```{r}
# Load packages
library(adegenet)
library(poppr)
library(dplyr)
library(reshape2)
library(ggplot2)
library(vcfR)
library(psych)
library(adespatial)
library(vegan)
```

### Prep Environmental Data

```{r}
# Import environmental data
env.raw = strata_6pops[,3:17]
```

```{r}
# Import spatial data
dbmem.raw = read.csv("../dbmems_6pops.csv")
```

```{r}
# Removing latitude and longitude
env.data = subset(env.raw, select = -c(Latitude,Longitude,SewageEffluent))
# pairs.panels(env.full_out6, scale = TRUE)
```

```{r}
#standardize the environmental data 
# Scale and center variables
env.z <- decostand(env.data, method = "standardize")
# Variables are now centered around a mean of 0
round(apply(env.z, 2, mean), 1)
# and scaled to have a standard deviation of 1
apply(env.z, 2, sd)
```

```{r}
#combine env and dbmem datasets 
env.full = cbind(env.z, dbmem.raw)
str(env.full)
```

### Prep allele frequency data

I'm only going to run RDA on pcadapt outliers, as the environmental associations with the LFMM outliers is already known.

pcadapt (db) SNP set 

```{r}
# Explore data
rad_out_db.filt
nLoc(rad_out_db.filt) # number of loci
nPop(rad_out_db.filt) # number of sites
nInd(rad_out_db.filt) # number of individuals
summary(rad_out_db.filt$pop) # sample size

# Calculate allele frequencies for each site
allele_freqs_out_db = data.frame(rraf(rad_out_db.filt, by_pop = TRUE, correction = FALSE), check.names = FALSE)

# Keep only the first of the two alleles for each SNP (since p=1-q).
allele_freqs_out_db = allele_freqs_out_db[, seq(1, dim(allele_freqs_out_db)[2], 2)]

# Export allele frequencies
write.csv(allele_freqs_out_db, file = "all_allele_freqs_out_db.csv", row.names = TRUE)
```

**Minor allele frequencies**

```{r}
# Separate genind object by site
site_list_out_db = seppop(rad_out_db.filt)
names(site_list_out_db)

# Calculate the minor allele frequency for each site
maf_list_out_db = lapply(site_list_out_db, FUN = minorAllele)

# Convert list to dataframe
maf_out_db = as.data.frame(maf_list_out_db) %>% t() %>% as.data.frame()

# Export minor allele frequencies
write.csv(maf_out_db, file = "minor_allele_freqs_out_db.csv", row.names = TRUE)
```

 pcadapt inside inversions (db_invert) SNP set 

```{r}
# Explore data
rad_out_db_invert.filt
nLoc(rad_out_db_invert.filt) # number of loci
nPop(rad_out_db_invert.filt) # number of sites
nInd(rad_out_db_invert.filt) # number of individuals
summary(rad_out_db_invert.filt$pop) # sample size

# Calculate allele frequencies for each site
allele_freqs_out_db_invert = data.frame(rraf(rad_out_db_invert.filt, by_pop = TRUE, correction = FALSE), check.names = FALSE)

# Keep only the first of the two alleles for each SNP (since p=1-q).
allele_freqs_out_db_invert = allele_freqs_out_db_invert[, seq(1, dim(allele_freqs_out_db_invert)[2], 2)]

# Export allele frequencies
write.csv(allele_freqs_out_db_invert, file = "all_allele_freqs_out_db_invert.csv", row.names = TRUE)
```

**Minor allele frequencies**

```{r}
# Separate genind object by site
site_list_out_db_invert = seppop(rad_out_db_invert.filt)
names(site_list_out_db_invert)

# Calculate the minor allele frequency for each site
maf_list_out_db_invert = lapply(site_list_out_db_invert, FUN = minorAllele)

# Convert list to dataframe
maf_out_db_invert = as.data.frame(maf_list_out_db_invert) %>% t() %>% as.data.frame()

# Export minor allele frequencies
write.csv(maf_out_db_invert, file = "minor_allele_freqs_out_db_invert.csv", row.names = TRUE)
```

 pcadapt outside inversions (db_noinvert) SNP set 

```{r}
# Explore data
rad_out_db_noinvert.filt
nLoc(rad_out_db_noinvert.filt) # number of loci
nPop(rad_out_db_noinvert.filt) # number of sites
nInd(rad_out_db_noinvert.filt) # number of individuals
summary(rad_out_db_noinvert.filt$pop) # sample size

# Calculate allele frequencies for each site
allele_freqs_out_db_noinvert = data.frame(rraf(rad_out_db_noinvert.filt, by_pop = TRUE, correction = FALSE), check.names = FALSE)

# Keep only the first of the two alleles for each SNP (since p=1-q).
allele_freqs_out_db_noinvert = allele_freqs_out_db_noinvert[, seq(1, dim(allele_freqs_out_db_noinvert)[2], 2)]

# Export allele frequencies
write.csv(allele_freqs_out_db_noinvert, file = "all_allele_freqs_out_db_noinvert.csv", row.names = TRUE)
```

**Minor allele frequencies**

```{r}
# Separate genind object by site
site_list_out_db_noinvert = seppop(rad_out_db_noinvert.filt)
names(site_list_out_db_noinvert)

# Calculate the minor allele frequency for each site
maf_list_out_db_noinvert = lapply(site_list_out_db_noinvert, FUN = minorAllele)

# Convert list to dataframe
maf_out_db_noinvert = as.data.frame(maf_list_out_db_noinvert) %>% t() %>% as.data.frame()

# Export minor allele frequencies
write.csv(maf_out_db_noinvert, file = "minor_allele_freqs_out_db_noinvert.csv", row.names = TRUE)
```


### Identify significant environmental predictors in RDA model

```{r}
# Set seed
set.seed(123)
```

pcadapt (db) SNP set

```{r}
# Use forward selection to identify significant environmental variables with ordiR2step
# first we need to create a null model and then a full model 
## Null model
RDA0_out_db <- rda(allele_freqs_out_db ~ 1,  env.full) 

## Full model
RDA_out_db <- rda(allele_freqs_out_db ~ ., env.full)

## adjusted R squared of full. model
adjR2.RDA_out_db <- RsquareAdj(RDA_out_db)$adj.r.squared

# Running ordiR2step to identify significant environmental variables in the model 
mod_out_db <- ordiR2step(RDA0_out_db, scope = formula(RDA_out_db), Pin = 0.1, permutations = 1000, R2scope = FALSE, direction = "both")
```

```{r}
# Use forward selection to identify significant environmental variables
Y_out_db = allele_freqs_out_db
X_out_db = env.z
env.for_out_db = forward.sel(Y_out_db, X_out_db, alpha = TRUE)
env.for_out_db

# Use forward selection to identify significant dbmems
dbmem.for_out_db = forward.sel(Y = allele_freqs_out_db, X = dbmem.raw, alpha = TRUE)
dbmem.for_out_db
```

pcadapt inside inversions (db_invert) SNP set

```{r}
# Use forward selection to identify significant environmental variables with ordiR2step
# first we need to create a null model and then a full model 
## Null model
RDA0_out_db_invert <- rda(allele_freqs_out_db_invert ~ 1,  env.full) 

## Full model
RDA_out_db_invert <- rda(allele_freqs_out_db_invert ~ ., env.full)

## adjusted R squared of full. model
adjR2.RDA_out_db_invert <- RsquareAdj(RDA_out_db_invert)$adj.r.squared

# Running ordiR2step to identify significant environmental variables in the model 
mod_out_db_invert <- ordiR2step(RDA0_out_db_invert, scope = formula(RDA_out_db_invert), Pin = 0.1, permutations = 1000, R2scope = FALSE, direction = "both")
```

```{r}
# Use forward selection to identify significant environmental variables
Y_out_db_invert = allele_freqs_out_db_invert
X_out_db_invert = env.z
env.for_out_db_invert = forward.sel(Y_out_db_invert, X_out_db_invert, alpha = TRUE)
env.for_out_db_invert

# Use forward selection to identify significant dbmems
dbmem.for_out_db_invert = forward.sel(Y = allele_freqs_out_db_invert, X = dbmem.raw, alpha = TRUE)
dbmem.for_out_db_invert
```

pcadapt outside inversions (db_noinvert) SNP set

```{r}
# Use forward selection to identify significant environmental variables with ordiR2step
# first we need to create a null model and then a full model 
## Null model
RDA0_out_db_noinvert <- rda(allele_freqs_out_db_noinvert ~ 1,  env.full) 

## Full model
RDA_out_db_noinvert <- rda(allele_freqs_out_db_noinvert ~ ., env.full)

## adjusted R squared of full. model
adjR2.RDA_out_db_noinvert <- RsquareAdj(RDA_out_db_noinvert)$adj.r.squared

# Running ordiR2step to identify significant environmental variables in the model 
mod_out_db_noinvert <- ordiR2step(RDA0_out_db_noinvert, scope = formula(RDA_out_db_noinvert), Pin = 0.1, permutations = 1000, R2scope = FALSE, direction = "both")
```

```{r}
# Use forward selection to identify significant environmental variables
Y_out_db_noinvert = allele_freqs_out_db_noinvert
X_out_db_noinvert = env.z
env.for_out_db_noinvert = forward.sel(Y_out_db_noinvert, X_out_db_noinvert, alpha = TRUE)
env.for_out_db_noinvert

# Use forward selection to identify significant dbmems
dbmem.for_out_db_noinvert = forward.sel(Y = allele_freqs_out_db_noinvert, X = dbmem.raw, alpha = TRUE)
dbmem.for_out_db_noinvert
```


### Running the Redundancy analysis

**Prepping for plotting** 

```{r}
# Create a data frame to correctly color regions
# Row names are the same across all allele frequency data sets  
col_dframe = data.frame("site" = rownames(allele_freqs_out_db))

# Function to add regional labels to dataframe
addregion_6pops = function(x){
  # If pop label is present function will output the region
  if(x=="BAR") y = " Barrington "
  if(x=="BIS") y = " Bissel Cove "
  if(x=="GB") y = " Greenwich Bay "
  if(x=="KIC") y = " Kickemuit "
  if(x=="MCD") y = " Donovan Marsh "
  if(x=="PVD") y = " Providence "
  return(y)
}

# Add regional labels
col_dframe$region = sapply(col_dframe$site, addregion_6pops)

# Add factor levels
region_order_6pops = c(" Barrington "," Bissel Cove "," Greenwich Bay "," Kickemuit "," Donovan Marsh ", " Providence ")
col_dframe$region = factor(col_dframe$region, levels = region_order_6pops)
```

```{r}
# Create colour scheme
# blue=#000088, green=#7FC97F, orange=#FF7F00, red=#E31A1C, purple=#9A32CD, pink=#FF1493, yellow=#FFD700, cyan=#00FFFF
cols_6pops = c("#7FC97F","#00008B","#FF7F00","#9A32CD","#FF1493","#00FFFF")
```


**Running RDA and plotting** 

pcadapt (db) SNP set

```{r}
# Perform RDA with all significant variables
rda_out_db = rda(allele_freqs_out_db ~ PC2 + PC3, data = env.full, scale = TRUE)
rda_out_db

# Model summaries
RsquareAdj(rda_out_db) # adjusted Rsquared 
vif.cca(rda_out_db) # variance inflation factor (<10 OK)
anova.cca(rda_out_db, permutations = 1000) # db model
anova.cca(rda_out_db, permutations = 1000, by="margin") # per variable 

# Variance explained by each canonical axis
summary(eigenvals(rda_out_db, model = "constrained"))
screeplot(rda_out_db)
```

```{r}
# Visualize results of RDA
#png("rda_out_db.png", width = 8, height = 7, units = "in", res = 600)
plot(rda_out_db, type="n", scaling = 3)
# SITES
points(rda_out_db, display="sites", pch=21, scaling=3, cex=1.5, col="black",
       bg=cols_6pops[col_dframe$region]) # sites
text(rda_out_db, display="sites", scaling = 3, col="black", font=2, pos=4)
# PREDICTORS
text(rda_out_db, display="bp", scaling=3, col="red1", cex=1, lwd=2)
# SNPS
# text(rda_out_db, display="species", scaling = 3, col="blue", cex=0.7, pos=4) # SNPs
# LEGEND
#legend("bottomright", legend=levels(col_dframe$region), bty="n", col="black",
      #pch=21, cex=1.2, pt.bg=cols_6pops)
# OTHER LABELS
adj.R2 = round(RsquareAdj(rda_out_db)$adj.r.squared, 3)
mtext(bquote(italic("Adj.R")^"2"~"= "~.(adj.R2)), side = 3, adj = 0.5)
# dev.off()
```

```{r}
# First, run the ANOVA to get the p-value
anova_result <- anova.cca(rda_out_db, permutations = 1000)
p_value <- anova_result$`Pr(>F)`[1]  # Extract the overall p-value

# Set up the plot with transparent background
png("rda_out_db.png", width = 9, height = 7, units = "in", res = 600, bg = "white")

# Increase the base font size
par(cex = 1.2)  # This will increase all text sizes by 20%

# Create the plot without axis labels
plot(rda_out_db, type="n", scaling = 3, 
     col.axis = "black", col.lab = "black", col.main = "black", col.sub = "black",
     xlab = "", ylab = "")  # Suppress axis labels here

# SITES
points(rda_out_db, display="sites", pch=21, scaling=3, cex=1.8, col="black",
       bg=adjustcolor(cols_6pops[col_dframe$region], alpha.f = 0.7))  # sites with semi-transparent fill
text(rda_out_db, display="sites", scaling = 3, col="black", font=2, pos=4, cex = 1.2)

# PREDICTORS
text(rda_out_db, display="bp", scaling=3, col="red1", cex=1.0, lwd=2)

# OTHER LABELS
#adj.R2 = round(RsquareAdj(rda_out_db)$adj.r.squared, 3)
#p_value_formatted = format.pval(p_value, digits = 3)
#mtext(bquote(italic("Adj.R")^"2"~"= "~.(adj.R2)~";"~"p"~"="~.(p_value_formatted)), 
      #side = 3, adj = 0.5, col = "black", cex = 1.2)

# Add axis labels in black with larger font (only once)
title(xlab = "RDA1 (82.5%)", ylab = "RDA2 (17.5%)", col.lab = "black", cex.lab = 1.4)

# Increase axis text size
par(cex.axis = 1.2)

# Add a white box around the plot
box(col = "black")

dev.off()
```

### Partial redundancy analysis - controlling for geographic location

```{r}
# Perform RDA while controlling for geographical location
pRDA_out_db = rda(allele_freqs_out_db ~ PC2 + PC3 + Condition(MEM1 + MEM2),
           data = env.full, scale = TRUE)
pRDA_out_db
RsquareAdj(pRDA_out_db) # adjusted Rsquared 
vif.cca(pRDA_out_db) # variance inflation factor (<10 OK)
anova.cca(pRDA_out_db, permutations = 1000) # db model
anova.cca(pRDA_out_db, permutations = 1000, by = "margin") # per variable

# Variance explained by each canonical axis
summary(eigenvals(pRDA_out_db, model = "constrained"))
screeplot(pRDA_out_db)
```

```{r}
# Visualise results of pRDA
#png("partial_rda_out_db.png", width = 9, height = 7, units = "in", res = 600)
plot(pRDA_out_db, type="n", scaling = 3)
#title("Seascape partial redundancy analysis")
# SITES
points(pRDA_out_db, display="sites", pch=21, scaling=3, cex=1.5, col="black",
       bg=cols_6pops[col_dframe$region]) # sites
text(pRDA_out_db, display="sites", scaling = 3, col="black", font=2, pos=4)
# PREDICTORS
text(pRDA_out_db, display="bp", scaling=3, col="red1", cex=1, lwd=2)
# SNPS
# text(pRDA, display="species", scaling = 3, col="blue", cex=0.7, pos=4) # SNPs
# LEGEND
#legend("topleft", legend=levels(col_dframe_out6$region), bty="n", col="black",
       #pch=21, cex=1.2, pt.bg=cols_6pops)
# OTHER LABELS
adj.R2 = round(RsquareAdj(pRDA_out_db)$adj.r.squared, 3)
mtext(bquote(italic("Adj.R")^"2"~"= "~.(adj.R2)), side = 3, adj = 0.5)
#dev.off()
```


pcadapt inside inversion (db_invert) SNP set

```{r}
# Perform RDA with all significant variables
rda_out_db_invert = rda(allele_freqs_out_db_invert ~ PC2 + PC3, data = env.full, scale = TRUE)
rda_out_db_invert

# Model summaries
RsquareAdj(rda_out_db_invert) # adjusted Rsquared 
vif.cca(rda_out_db_invert) # variance inflation factor (<10 OK)
anova.cca(rda_out_db_invert, permutations = 1000) # db_invert model
anova.cca(rda_out_db_invert, permutations = 1000, by="margin") # per variable 

# Variance explained by each canonical axis
summary(eigenvals(rda_out_db_invert, model = "constrained"))
screeplot(rda_out_db_invert)
```

```{r}
# Visualize results of RDA
#png("rda_out_db_invert.png", width = 8, height = 7, units = "in", res = 600)
plot(rda_out_db_invert, type="n", scaling = 3)
# SITES
points(rda_out_db_invert, display="sites", pch=21, scaling=3, cex=1.5, col="black",
       bg=cols_6pops[col_dframe$region]) # sites
text(rda_out_db_invert, display="sites", scaling = 3, col="black", font=2, pos=4)
# PREDICTORS
text(rda_out_db_invert, display="bp", scaling=3, col="red1", cex=1, lwd=2)
# SNPS
# text(rda_out_db_invert, display="species", scaling = 3, col="blue", cex=0.7, pos=4) # SNPs
# LEGEND
#legend("bottomright", legend=levels(col_dframe$region), bty="n", col="black",
      #pch=21, cex=1.2, pt.bg=cols_6pops)
# OTHER LABELS
adj.R2 = round(RsquareAdj(rda_out_db_invert)$adj.r.squared, 3)
mtext(bquote(italic("Adj.R")^"2"~"= "~.(adj.R2)), side = 3, adj = 0.5)
# dev.off()
```

```{r}
# First, run the ANOVA to get the p-value
anova_result <- anova.cca(rda_out_db_invert, permutations = 1000)
p_value <- anova_result$`Pr(>F)`[1]  # Extract the overall p-value

# Set up the plot with transparent background
png("rda_out_db_invert.png", width = 9, height = 7, units = "in", res = 600, bg = "white")

# Increase the base font size
par(cex = 1.2)  # This will increase all text sizes by 20%

# Create the plot without axis labels
plot(rda_out_db_invert, type="n", scaling = 3, 
     col.axis = "black", col.lab = "black", col.main = "black", col.sub = "black",
     xlab = "", ylab = "")  # Suppress axis labels here

# SITES
points(rda_out_db_invert, display="sites", pch=21, scaling=3, cex=1.8, col="black",
       bg=adjustcolor(cols_6pops[col_dframe$region], alpha.f = 0.7))  # sites with semi-transparent fill
text(rda_out_db_invert, display="sites", scaling = 3, col="black", font=2, pos=4, cex = 1.2)

# PREDICTORS
text(rda_out_db_invert, display="bp", scaling=3, col="red1", cex=1.0, lwd=2)

# OTHER LABELS
#adj.R2 = round(RsquareAdj(rda_out_db_invert)$adj.r.squared, 3)
#p_value_formatted = format.pval(p_value, digits = 3)
#mtext(bquote(italic("Adj.R")^"2"~"= "~.(adj.R2)~";"~"p"~"="~.(p_value_formatted)), 
      #side = 3, adj = 0.5, col = "black", cex = 1.2)

# Add axis labels in black with larger font (only once)
title(xlab = "RDA1 (92.4%)", ylab = "RDA2 (7.6%)", col.lab = "black", cex.lab = 1.4)

# Increase axis text size
par(cex.axis = 1.2)

# Add a white box around the plot
box(col = "black")

dev.off()
```

### Partial redundancy analysis - controlling for geographic location

```{r}
# Perform RDA while controlling for geographical location
pRDA_out_db_invert = rda(allele_freqs_out_db_invert ~ PC2 + PC3 + Condition(MEM1 + MEM2),
           data = env.full, scale = TRUE)
pRDA_out_db_invert
RsquareAdj(pRDA_out_db_invert) # adjusted Rsquared 
vif.cca(pRDA_out_db_invert) # variance inflation factor (<10 OK)
anova.cca(pRDA_out_db_invert, permutations = 1000) # db_invert model
anova.cca(pRDA_out_db_invert, permutations = 1000, by = "margin") # per variable

# Variance explained by each canonical axis
summary(eigenvals(pRDA_out_db_invert, model = "constrained"))
screeplot(pRDA_out_db_invert)
```

```{r}
# Visualise results of pRDA
#png("partial_rda_out_db_invert.png", width = 9, height = 7, units = "in", res = 600)
plot(pRDA_out_db_invert, type="n", scaling = 3)
#title("Seascape partial redundancy analysis")
# SITES
points(pRDA_out_db_invert, display="sites", pch=21, scaling=3, cex=1.5, col="black",
       bg=cols_6pops[col_dframe$region]) # sites
text(pRDA_out_db_invert, display="sites", scaling = 3, col="black", font=2, pos=4)
# PREDICTORS
text(pRDA_out_db_invert, display="bp", scaling=3, col="red1", cex=1, lwd=2)
# SNPS
# text(pRDA, display="species", scaling = 3, col="blue", cex=0.7, pos=4) # SNPs
# LEGEND
#legend("topleft", legend=levels(col_dframe_out6$region), bty="n", col="black",
       #pch=21, cex=1.2, pt.bg=cols_6pops)
# OTHER LABELS
adj.R2 = round(RsquareAdj(pRDA_out_db_invert)$adj.r.squared, 3)
mtext(bquote(italic("Adj.R")^"2"~"= "~.(adj.R2)), side = 3, adj = 0.5)
#dev.off()
```


pcadapt outside inversions (db_noinvert) SNP set

```{r}
# Perform RDA with all significant variables
rda_out_db_noinvert = rda(allele_freqs_out_db_noinvert ~ PC2 + PC3, data = env.full, scale = TRUE)
rda_out_db_noinvert

# Model summaries
RsquareAdj(rda_out_db_noinvert) # adjusted Rsquared 
vif.cca(rda_out_db_noinvert) # variance inflation factor (<10 OK)
anova.cca(rda_out_db_noinvert, permutations = 1000) # db_noinvert model
anova.cca(rda_out_db_noinvert, permutations = 1000, by="margin") # per variable 

# Variance explained by each canonical axis
summary(eigenvals(rda_out_db_noinvert, model = "constrained"))
screeplot(rda_out_db_noinvert)
```

```{r}
# Visualize results of RDA
#png("rda_out_db_noinvert.png", width = 8, height = 7, units = "in", res = 600)
plot(rda_out_db_noinvert, type="n", scaling = 3)
# SITES
points(rda_out_db_noinvert, display="sites", pch=21, scaling=3, cex=1.5, col="black",
       bg=cols_6pops[col_dframe$region]) # sites
text(rda_out_db_noinvert, display="sites", scaling = 3, col="black", font=2, pos=4)
# PREDICTORS
text(rda_out_db_noinvert, display="bp", scaling=3, col="red1", cex=1, lwd=2)
# SNPS
# text(rda_out_db_noinvert, display="species", scaling = 3, col="blue", cex=0.7, pos=4) # SNPs
# LEGEND
#legend("bottomright", legend=levels(col_dframe$region), bty="n", col="black",
      #pch=21, cex=1.2, pt.bg=cols_6pops)
# OTHER LABELS
adj.R2 = round(RsquareAdj(rda_out_db_noinvert)$adj.r.squared, 3)
mtext(bquote(italic("Adj.R")^"2"~"= "~.(adj.R2)), side = 3, adj = 0.5)
# dev.off()
```

```{r}
# First, run the ANOVA to get the p-value
anova_result <- anova.cca(rda_out_db_noinvert, permutations = 1000)
p_value <- anova_result$`Pr(>F)`[1]  # Extract the overall p-value

# Set up the plot with transparent background
png("rda_out_db_noinvert.png", width = 9, height = 7, units = "in", res = 600, bg = "white")

# Increase the base font size
par(cex = 1.2)  # This will increase all text sizes by 20%

# Create the plot without axis labels
plot(rda_out_db_noinvert, type="n", scaling = 3, 
     col.axis = "black", col.lab = "black", col.main = "black", col.sub = "black",
     xlab = "", ylab = "")  # Suppress axis labels here

# SITES
points(rda_out_db_noinvert, display="sites", pch=21, scaling=3, cex=1.8, col="black",
       bg=adjustcolor(cols_6pops[col_dframe$region], alpha.f = 0.7))  # sites with semi-transparent fill
text(rda_out_db_noinvert, display="sites", scaling = 3, col="black", font=2, pos=4, cex = 1.2)

# PREDICTORS
text(rda_out_db_noinvert, display="bp", scaling=3, col="red1", cex=1.0, lwd=2)

# OTHER LABELS
#adj.R2 = round(RsquareAdj(rda_out_db_noinvert)$adj.r.squared, 3)
#p_value_formatted = format.pval(p_value, digits = 3)
#mtext(bquote(italic("Adj.R")^"2"~"= "~.(adj.R2)~";"~"p"~"="~.(p_value_formatted)), 
      #side = 3, adj = 0.5, col = "black", cex = 1.2)

# Add axis labels in black with larger font (only once)
title(xlab = "RDA1 (64.7%)", ylab = "RDA2 (35.3%)", col.lab = "black", cex.lab = 1.4)

# Increase axis text size
par(cex.axis = 1.2)

# Add a white box around the plot
box(col = "black")

dev.off()
```

### Partial redundancy analysis - controlling for geographic location

```{r}
# Perform RDA while controlling for geographical location
pRDA_out_db_noinvert = rda(allele_freqs_out_db_noinvert ~ PC2 + PC3 + Condition(MEM1 + MEM2),
           data = env.full, scale = TRUE)
pRDA_out_db_noinvert
RsquareAdj(pRDA_out_db_noinvert) # adjusted Rsquared 
vif.cca(pRDA_out_db_noinvert) # variance inflation factor (<10 OK)
anova.cca(pRDA_out_db_noinvert, permutations = 1000) # db_noinvert model
anova.cca(pRDA_out_db_noinvert, permutations = 1000, by = "margin") # per variable

# Variance explained by each canonical axis
summary(eigenvals(pRDA_out_db_noinvert, model = "constrained"))
screeplot(pRDA_out_db_noinvert)
```

```{r}
# Visualise results of pRDA
#png("partial_rda_out_db_noinvert.png", width = 9, height = 7, units = "in", res = 600)
plot(pRDA_out_db_noinvert, type="n", scaling = 3)
#title("Seascape partial redundancy analysis")
# SITES
points(pRDA_out_db_noinvert, display="sites", pch=21, scaling=3, cex=1.5, col="black",
       bg=cols_6pops[col_dframe$region]) # sites
text(pRDA_out_db_noinvert, display="sites", scaling = 3, col="black", font=2, pos=4)
# PREDICTORS
text(pRDA_out_db_noinvert, display="bp", scaling=3, col="red1", cex=1, lwd=2)
# SNPS
# text(pRDA, display="species", scaling = 3, col="blue", cex=0.7, pos=4) # SNPs
# LEGEND
#legend("topleft", legend=levels(col_dframe_out6$region), bty="n", col="black",
       #pch=21, cex=1.2, pt.bg=cols_6pops)
# OTHER LABELS
adj.R2 = round(RsquareAdj(pRDA_out_db_noinvert)$adj.r.squared, 3)
mtext(bquote(italic("Adj.R")^"2"~"= "~.(adj.R2)), side = 3, adj = 0.5)
#dev.off()
```


## Allele frequency heatmaps 

Here, I want to identify SNPs that show distinct patterns of allele frequencies that may be driving the patterns in the RDA. I'm going to calculate allele frequencies for the db outliers inside the inversions and outside the inversions that had the highest loadings on the RDA axes. I'll check for RDA1 and RDA2. My hope is that this will reveal significant loci which I can then link to specific genes and protein functions the SNPs fall within. 

```{r}
#install.packages("remotes")
#remotes::install_github("chollenbeck/pantomime")
```

```{r}
library(pantomime)
```

RDA1 

```{r}
contributions_db_invert <- (scores(rda_out_db_invert, display = "species", scaling = 2)^2)

# Calculate percentage contributions
perc_contributions_db_invert <- (contributions_db_invert / rowSums(contributions_db_invert) * 100)

# Ranking them by their percent contributions 
perc_contributions_db_invert_df <- as.data.frame(perc_contributions_db_invert) %>%
  arrange(desc(RDA1))

# Option 1: Top 10%
top10_cutoff_db_invert <- quantile(perc_contributions_db_invert[,1], 0.9)

# Option 2: Above average
avg_cutoff_db_invert <- mean(perc_contributions_db_invert[,1])

high_contrib_vars_top10_db_invert <- which(perc_contributions_db_invert[,1] > top10_cutoff_db_invert)

high_contrib_vars_avg_db_invert <- which(perc_contributions_db_invert[,1] > avg_cutoff_db_invert)

# Plot contributions for the first axis
barplot(sort(perc_contributions_db_invert[,1], decreasing = TRUE), 
        main = "Variable Contributions to RDA1",
        ylab = "Percentage Contribution")
```

```{r}
# Identify SNPs with significant loadings for RDA1
rda_loadings_out_db_invert_rda1 <- loadings_out_db_invert %>%
  as_tibble(rownames = "allele") %>%
  extract(allele, c("locus","chrom", "pos", "allele"), "(NC_(\\d+)_1_(\\d+)).(\\d)") %>%
  mutate(pos = as.integer(pos)) %>%
  dplyr::select(locus,chrom, pos, allele, RDA1, RDA2) %>%
  mutate(RDA1 = RDA1^2, RDA2 = RDA2^2) %>%
  arrange(desc(RDA1))

db_invert_pos_values_rda <- c("63925159","65249871","65249956","65249962","65267956","65410348","69663441","69663980","73286223","73479264","78251922","31392783","31899147","32464100","32511570","34965793","34965894","36419831","41726211","41726273","42741412","42884268","42885299")

top_loadings_out_db_invert_rda1 <- rda_loadings_out_db_invert_rda1[rda_loadings_out_db_invert_rda1$pos %in% db_invert_pos_values_rda, ]

top_loadings_out_db_invert_rda1 %>%
  arrange(desc(RDA1))

rad_out_db_invert.filt %>%
  pantomime::get_allele_freqs()

rad_out_db_invert.filt %>%
  pantomime::get_allele_freqs() %>%
  filter(allele == 0, locus %in% top_loadings_out_db_invert_rda1$locus) %>%
  extract(locus, c("chrom", "pos"), "NC_(\\d+)_1_(\\d+)", remove = FALSE) %>%
  mutate(pos = as.integer(pos)) %>%
  mutate(pop = factor(pop, levels = c("BAR", "BIS", "GB", "KIC", "MCD", "PVD"))) %>%
  mutate(chrom_pos = paste("NC", chrom, "_", pos)) %>%  # Create a new column for facet labels
  ggplot(aes(x = pop, y = freq, fill = pop)) +
    geom_col() +
    scale_fill_manual(values = cols_6pops) +
    facet_wrap(~chrom_pos, scales = "free_y") +  # Use chrom_pos for faceting
    theme_minimal() +
    theme(
      strip.text = element_text(size = 8),  # Adjust the size of the facet labels
      axis.text.x = element_text(angle = 45, hjust = 1),  # Rotate x-axis labels
      legend.position = "none"  # Remove legend if not needed
    ) +
    labs(x = "Population", y = "Allele Frequency")
```

RDA2

```{r}
# Identify SNPs with significant loadings for RDA2
rda_loadings_out_db_invert_rda2 <- loadings_out_db_invert %>%
  as_tibble(rownames = "allele") %>%
  extract(allele, c("locus","chrom", "pos", "allele"), "(NC_(\\d+)_1_(\\d+)).(\\d)") %>%
  mutate(pos = as.integer(pos)) %>%
  select(locus,chrom, pos, allele, RDA1, RDA2) %>%
  mutate(RDA1 = RDA1^2, RDA2 = RDA2^2) %>%
  arrange(desc(RDA2))

top_loadings_out_db_invert_rda2 <- rda_loadings_out_db_invert_rda2 %>%
  top_n(10, RDA2)

rad_out_db_invert.filt %>%
  pantomime::get_allele_freqs()

rad_out_db_invert.filt %>%
  pantomime::get_allele_freqs() %>%
  filter(allele == 0, locus %in% top_loadings_out_db_invert_rda2$locus) %>%
  extract(locus, c("chrom", "pos"), "NC_(\\d+)_1_(\\d+)", remove = FALSE) %>%
  mutate(pos = as.integer(pos)) %>%
  mutate(pop = factor(pop, levels = c("BAR", "BIS", "GB", "KIC", "MCD", "PVD"))) %>%
  ggplot(aes(x = pop, y = freq, fill = pop)) +
    geom_col() +
    scale_fill_manual(values = cols_6pops) +
    facet_wrap(~locus)
```

```{r}
ggplot(rda_loadings_out_db_invert, aes(x = pos, y = RDA1)) +
  geom_point() +
  facet_wrap(~chrom) +
  theme_minimal() +
  scale_color_manual()
```

```{r}
ggplot(rda_loadings_out_db_invert, aes(x = pos, y = RDA2)) +
  geom_point() +
  facet_wrap(~chrom) +
  theme_minimal() +
  scale_color_manual()
```

pcadapt outliers outside inversions set

```{r}
# Save loadings from RDA
loadings_out_db_noinvert <- scores(rda_out_db_noinvert, display = "species", scaling = 0)
```


SNPs outside inversions

pcadapt outliers inside inversions set

RDA1

```{r}
contributions_db_noinvert <- (scores(rda_out_db_noinvert, display = "species", scaling = 2)^2)

# Calculate percentage contributions
perc_contributions_db_noinvert <- (contributions_db_noinvert / rowSums(contributions_db_noinvert) * 100)

# Ranking them by their percent contributions 
perc_contributions_db_noinvert_df <- as.data.frame(perc_contributions_db_noinvert) %>%
  arrange(desc(RDA1))

# Option 1: Top 10%
top10_cutoff_db_noinvert <- quantile(perc_contributions_db_noinvert[,1], 0.9)

# Option 2: Above average
avg_cutoff_db_noinvert <- mean(perc_contributions_db_noinvert[,1])

high_contrib_vars_top10_db_noinvert <- which(perc_contributions_db_noinvert[,1] > top10_cutoff_db_noinvert)

high_contrib_vars_avg_db_noinvert <- which(perc_contributions_db_noinvert[,1] > avg_cutoff_db_noinvert)

# Plot contributions for the first axis
barplot(sort(perc_contributions_db_noinvert[,1], decreasing = TRUE), 
        main = "Variable Contributions to RDA1",
        ylab = "Percentage Contribution")
```

```{r}
# Identify SNPs with significant loadings for RDA1
rda_loadings_out_db_noinvert_rda1 <- loadings_out_db_noinvert %>%
  as_tibble(rownames = "allele") %>%
  extract(allele, c("locus","chrom", "pos", "allele"), "(NC_(\\d+)_1_(\\d+)).(\\d)") %>%
  mutate(pos = as.integer(pos)) %>%
  dplyr::select(locus,chrom, pos, allele, RDA1, RDA2) %>%
  mutate(RDA1 = RDA1^2, RDA2 = RDA2^2) %>%
  arrange(desc(RDA1))

db_noinvert_pos_values_rda <- c("8465456","8465500","8465519","8466007","8748233","13250687","29901078","40734412","42946753","23152223","39585945")

top_loadings_out_db_noinvert_rda1 <- rda_loadings_out_db_noinvert_rda1[rda_loadings_out_db_noinvert_rda1$pos %in% db_noinvert_pos_values_rda, ]

top_loadings_out_db_noinvert_rda1 %>%
  arrange(desc(RDA1))

rad_out_db_noinvert.filt %>%
  pantomime::get_allele_freqs()

rad_out_db_noinvert.filt %>%
  pantomime::get_allele_freqs() %>%
  filter(allele == 0, locus %in% top_loadings_out_db_noinvert_rda1$locus) %>%
  extract(locus, c("chrom", "pos"), "NC_(\\d+)_1_(\\d+)", remove = FALSE) %>%
  mutate(pos = as.integer(pos)) %>%
  mutate(pop = factor(pop, levels = c("BAR", "BIS", "GB", "KIC", "MCD", "PVD"))) %>%
  mutate(chrom_pos = paste("NC", chrom, "_", pos)) %>%  # Create a new column for facet labels
  ggplot(aes(x = pop, y = freq, fill = pop)) +
    geom_col() +
    scale_fill_manual(values = cols_6pops) +
    facet_wrap(~chrom_pos, scales = "free_y") +  # Use chrom_pos for faceting
    theme_minimal() +
    theme(
      strip.text = element_text(size = 8),  # Adjust the size of the facet labels
      axis.text.x = element_text(angle = 45, hjust = 1),  # Rotate x-axis labels
      legend.position = "none"  # Remove legend if not needed
    ) +
    labs(x = "Population", y = "Allele Frequency")
```

RDA2

```{r}
# Identify SNPs with significant loadings for RDA2
rda_loadings_out_db_noinvert_rda2 <- loadings_out_db_noinvert %>%
  as_tibble(rownames = "allele") %>%
  extract(allele, c("locus","chrom", "pos", "allele"), "(NC_(\\d+)_1_(\\d+)).(\\d)") %>%
  mutate(pos = as.integer(pos)) %>%
  select(locus,chrom, pos, allele, RDA1, RDA2) %>%
  mutate(RDA1 = RDA1^2, RDA2 = RDA2^2) %>%
  arrange(desc(RDA2))

top_loadings_out_db_noinvert_rda2 <- rda_loadings_out_db_noinvert_rda2 %>%
  top_n(9, RDA2)

rad_out_db_noinvert.filt %>%
  pantomime::get_allele_freqs()

rad_out_db_noinvert.filt %>%
  pantomime::get_allele_freqs() %>%
  filter(allele == 0, locus %in% top_loadings_out_db_noinvert_rda2$locus) %>%
  extract(locus, c("chrom", "pos"), "NC_(\\d+)_1_(\\d+)", remove = FALSE) %>%
  mutate(pos = as.integer(pos)) %>%
  mutate(pop = factor(pop, levels = c("BAR", "BIS", "GB", "KIC", "MCD", "PVD"))) %>%
  ggplot(aes(x = pop, y = freq, fill = pop)) +
    geom_col() +
    scale_fill_manual(values = cols_6pops) +
    facet_wrap(~locus)
```

```{r}
ggplot(rda_loadings_out_db_noinvert, aes(x = pos, y = RDA1)) +
  geom_point() +
  facet_wrap(~chrom) +
  theme_minimal() +
  scale_color_manual()
```

```{r}
ggplot(rda_loadings_out_db_noinvert, aes(x = pos, y = RDA2)) +
  geom_point() +
  facet_wrap(~chrom) +
  theme_minimal() +
  scale_color_manual()
```

## Correspondence analysis-based RDA

Outlier SNPs inside inversions

```{r}
# Run a CA on the genind object
obj_out_db_invert <- genind2genpop(rad_out_db_invert.filt)
ca1_out_db_invert <- dudi.coa(tab(obj_out_db_invert), scannf=FALSE, nf=10)
eig_ca1_out_db_invert <- round((ca1_out_db_invert$eig / (sum(ca1_out_db_invert$eig))) * 100, 2)
out_ca_out_db_invert <- ca1_out_db_invert$li[,1:2]
rownames(out_ca_out_db_invert) <- popNames(rad_out_db_invert.filt)
```

```{r}
out_ca_out_db_invert <- out_ca_out_db_invert %>%
  rownames_to_column(var = "Population") %>%
  mutate(Population = factor(Population, levels = c("BAR", "BIS", "GB", "KIC", "MCD", "PVD"))) %>%
  select(Population, Axis1, Axis2)
```

```{r}
strata_ca_out_db_invert <- strata_6pops %>%
  left_join(out_ca_out_db_invert, by = "Population")
```

```{r}
# Plot the CA
ggplot(strata_ca_out_db_invert, aes(x = Axis1, y = Axis2, colour = Population)) +
  geom_point(size = 6, alpha = 0.8) +
  scale_colour_manual(values = cols_6pops,
                      name = "") +
  theme_minimal() +
  theme(plot.title = element_text(size = 14)) +
  labs(x = paste("", "Axis 1: ", eig_ca1_out_db_invert[1], "%"), y = paste("", "Axis 2: ", eig_ca1_out_db_invert[2], "%"))
```

```{r}
# Plot CA1 versus PC2 
strata_ca_out_db_invert %>%
  ggplot(aes(x = Axis1, y = PC2)) +
  geom_smooth(method = "lm", col = "darkgrey", lty = 1) +
  geom_point(aes(color = Population), size = 4) +
  scale_color_manual(values = cols_6pops) +
  theme_minimal() +
  labs(
       x = "CA1",
       y = "PC2")
```

```{r}
# Linear model for CA1 versus PC2
pc2_mod_out_db_invert <- lm(Axis1 ~ PC3, data = strata_ca_out_db_invert)
summary(pc2_mod_out_db_invert)
```

```{r}
ca_loadings_out_db_invert_ca1 <- ca1_out_db_invert$co %>%
  as_tibble(rownames = "allele") %>%
  extract(allele, c("locus","chrom", "pos", "allele"), "(NC_(\\d+)_1_(\\d+)).(\\d)") %>%
  mutate(pos = as.integer(pos)) %>%
  dplyr::select(locus, chrom, pos, allele, Comp1, Comp2) %>%
  mutate(Comp1 = Comp1^2, Comp2 = Comp2^2) %>%
  arrange(desc(Comp1))

db_invert_pos_values <- c("63925159","69663441","69663980","32673372","34965793","41726211","41726273", "65694744","65267956","34965894","31899147","42741412","65267874","66635829","73286223", "36420436","78393626","78393650","65267667")

top_loadings_out_db_invert_ca1 <- ca_loadings_out_db_invert_ca1[ca_loadings_out_db_invert_ca1$pos %in% db_invert_pos_values, ]

rad_out_db_invert.filt %>%
  pantomime::get_allele_freqs() %>%
  filter(allele == 0, locus %in% top_loadings_out_db_invert_ca1$locus) %>%
  extract(locus, c("chrom", "pos"), "NC_(\\d+)_1_(\\d+)", remove = FALSE) %>%
  mutate(pos = as.integer(pos)) %>%
  mutate(pop = factor(pop, levels = c("BAR", "BIS", "GB", "KIC", "MCD", "PVD"))) %>%
  mutate(chrom_pos = paste("NC", chrom, "_", pos)) %>%  # Create a new column for facet labels
  ggplot(aes(x = pop, y = freq, fill = pop)) +
    geom_col() +
    scale_fill_manual(values = cols_6pops) +
    facet_wrap(~chrom_pos, scales = "free_y") +  # Use chrom_pos for faceting
    theme_minimal() +
    theme(
      strip.text = element_text(size = 8),  # Adjust the size of the facet labels
      axis.text.x = element_text(angle = 45, hjust = 1),  # Rotate x-axis labels
      legend.position = "none"  # Remove legend if not needed
    ) +
    labs(x = "Population", y = "Allele Frequency")
```


Outlier SNPs outside inversions

```{r}
# Run a CA on the genind object
obj_out_db_noinvert <- genind2genpop(rad_out_db_noinvert.filt)
ca1_out_db_noinvert <- dudi.coa(tab(obj_out_db_noinvert), scannf=FALSE, nf=10)
eig_ca1_out_db_noinvert <- round((ca1_out_db_noinvert$eig / (sum(ca1_out_db_noinvert$eig))) * 100, 2)
out_ca_out_db_noinvert <- ca1_out_db_noinvert$li[,1:2]
rownames(out_ca_out_db_noinvert) <- popNames(rad_out_db_noinvert.filt)
```

```{r}
out_ca_out_db_noinvert <- out_ca_out_db_noinvert %>%
  rownames_to_column(var = "Population") %>%
  mutate(Population = factor(Population, levels = c("BAR", "BIS", "GB", "KIC", "MCD", "PVD"))) %>%
  select(Population, Axis1, Axis2)
```

```{r}
strata_ca_out_db_noinvert <- strata_6pops %>%
  left_join(out_ca_out_db_noinvert, by = "Population")
```

```{r}
# Plot the CA
ggplot(strata_ca_out_db_noinvert, aes(x = Axis1, y = Axis2, colour = Population)) +
  geom_point(size = 6, alpha = 0.8) +
  scale_colour_manual(values = cols_6pops,
                      name = "") +
  theme_minimal() +
  theme(plot.title = element_text(size = 14)) +
  labs(x = paste("", "Axis 1: ", eig_ca1_out_db_noinvert[1], "%"), y = paste("", "Axis 2: ", eig_ca1_out_db_noinvert[2], "%"))
```

```{r}
# Plot CA1 versus PC3 
strata_ca_out_db_noinvert %>%
  ggplot(aes(x = Axis1, y = PC2)) +
  geom_smooth(method = "lm", col = "darkgrey", lty = 1) +
  geom_point(aes(color = Population), size = 4) +
  scale_color_manual(values = cols_6pops) +
  theme_minimal() +
  labs(
       x = "CA1",
       y = "PC2")
```

```{r}
# Linear model for CA1 versus PC2
pc3_mod_out_db_noinvert <- lm(Axis1 ~ PC3, data = strata_ca_out_db_noinvert)
summary(pc3_mod_out_db_noinvert)
```

```{r}
# Identify SNPs with significant loadings for CA1
ca_loadings_out_db_noinvert_ca1 <- ca1_out_db_noinvert$co %>%
  as_tibble(rownames = "allele") %>%
  extract(allele, c("locus","chrom", "pos", "allele"), "(NC_(\\d+)_1_(\\d+)).(\\d)") %>%
  mutate(pos = as.integer(pos)) %>%
  dplyr::select(locus, chrom, pos, allele, Comp1, Comp2) %>%
  mutate(Comp1 = Comp1^2, Comp2 = Comp2^2) %>%
  arrange(desc(Comp1))

db_noinvert_pos_values <- c("29902142","8748233","1884313","42946753","47484403","49770765","39585945","8466007","40734412")

top_loadings_out_db_noinvert_ca1 <- ca_loadings_out_db_noinvert_ca1[ca_loadings_out_db_noinvert_ca1$pos %in% db_noinvert_pos_values, ]

rad_out_db_noinvert.filt %>%
  pantomime::get_allele_freqs()

rad_out_db_noinvert.filt %>%
  pantomime::get_allele_freqs() %>%
  filter(allele == 0, locus %in% top_loadings_out_db_noinvert_ca1$locus) %>%
  extract(locus, c("chrom", "pos"), "NC_(\\d+)_1_(\\d+)", remove = FALSE) %>%
  mutate(pos = as.integer(pos)) %>%
  mutate(pop = factor(pop, levels = c("BAR", "BIS", "GB", "KIC", "MCD", "PVD"))) %>%
  mutate(chrom_pos = paste("NC", chrom, "_", pos)) %>%  # Create a new column for facet labels
  ggplot(aes(x = pop, y = freq, fill = pop)) +
    geom_col() +
    scale_fill_manual(values = cols_6pops) +
    facet_wrap(~chrom_pos, scales = "free_y") +  # Use chrom_pos for faceting
    theme_minimal() +
    theme(
      strip.text = element_text(size = 8),  # Adjust the size of the facet labels
      axis.text.x = element_text(angle = 45, hjust = 1),  # Rotate x-axis labels
      legend.position = "none"  # Remove legend if not needed
    ) +
    labs(x = "Population", y = "Allele Frequency")
```

Correspondence analysis and allele frequencies plot for LFMM outliers 

```{r}
# Run a CA on the genind object
obj_out_lfmm <- genind2genpop(rad_out_lfmm.filt)
ca1_out_lfmm <- dudi.coa(tab(obj_out_lfmm), scannf=FALSE, nf=10)
eig_ca1_out_lfmm <- round((ca1_out_lfmm$eig / (sum(ca1_out_lfmm$eig))) * 100, 2)
out_ca_out_lfmm <- ca1_out_lfmm$li[,1:2]
rownames(out_ca_out_lfmm) <- popNames(rad_out_lfmm.filt)
```

```{r}
out_ca_out_lfmm <- out_ca_out_lfmm %>%
  rownames_to_column(var = "Population") %>%
  mutate(Population = factor(Population, levels = c("BAR", "BIS", "GB", "KIC", "MCD", "PVD"))) %>%
  dplyr::select(Population, Axis1, Axis2)
```

```{r}
strata_ca_out_lfmm <- strata_6pops %>%
  left_join(out_ca_out_lfmm, by = "Population")
```

```{r}
# Identify SNPs with significant loadings for CA1
ca_loadings_out_lfmm_ca1 <- ca1_out_lfmm$co %>%
  as_tibble(rownames = "allele") %>%
  extract(allele, c("locus","chrom", "pos", "allele"), "(NC_(\\d+)_1_(\\d+)).(\\d)") %>%
  mutate(pos = as.integer(pos)) %>%
  dplyr::select(locus, chrom, pos, allele, Comp1, Comp2) %>%
  mutate(Comp1 = Comp1^2, Comp2 = Comp2^2) %>%
  arrange(desc(Comp1))

lfmm_pos_values <- c("34532963","8396484","8396493","24512990","49023293","42294461","15925374","79541695")

top_loadings_out_lfmm_ca1 <- ca_loadings_out_lfmm_ca1[ca_loadings_out_lfmm_ca1$pos %in% lfmm_pos_values, ]

rad_out_lfmm.filt %>%
  pantomime::get_allele_freqs() %>%
  filter(allele == 0, locus %in% top_loadings_out_lfmm_ca1$locus) %>%
  extract(locus, c("chrom", "pos"), "NC_(\\d+)_1_(\\d+)", remove = FALSE) %>%
  mutate(pos = as.integer(pos)) %>%
  mutate(pop = factor(pop, levels = c("BAR", "BIS", "GB", "KIC", "MCD", "PVD"))) %>%
  mutate(chrom_pos = paste("NC", chrom, "_", pos)) %>%  # Create a new column for facet labels
  ggplot(aes(x = pop, y = freq, fill = pop)) +
    geom_col() +
    scale_fill_manual(values = cols_6pops) +
    facet_wrap(~chrom_pos, scales = "free_y") +  # Use chrom_pos for faceting
    theme_minimal() +
    theme(
      strip.text = element_text(size = 8),  # Adjust the size of the facet labels
      axis.text.x = element_text(angle = 45, hjust = 1),  # Rotate x-axis labels
      legend.position = "none"  # Remove legend if not needed
    ) +
    labs(x = "Population", y = "Allele Frequency")
```


```{r}
rad_out_lfmm.filt %>%
  pantomime::get_allele_freqs() %>%
  filter(allele == 0, locus %in% rad_out_lfmm.filt$loc.fac) %>%
  extract(locus, c("chrom", "pos"), "NC_(\\d+)_1_(\\d+)", remove = FALSE) %>%
  mutate(pos = as.integer(pos)) %>%
  mutate(pop = factor(pop, levels = c("BAR", "BIS", "GB", "KIC", "MCD", "PVD"))) %>%
  mutate(chrom_pos = paste("NC", chrom, "_", pos)) %>%  # Create a new column for facet labels
  ggplot(aes(x = pop, y = freq, fill = pop)) +
    geom_col() +
    scale_fill_manual(values = cols_6pops) +
    facet_wrap(~chrom_pos, scales = "free_y") +  # Use chrom_pos for faceting
    theme_minimal() +
    theme(
      strip.text = element_text(size = 8),  # Adjust the size of the facet labels
      axis.text.x = element_text(angle = 45, hjust = 1),  # Rotate x-axis labels
      legend.position = "none"  # Remove legend if not needed
    ) +
    labs(x = "Population", y = "Allele Frequency")
```


Full final set of outliers (390)

```{r}
# Run a CA on the genind object
obj_out_full_final <- genind2genpop(rad_out_full_final.filt)
ca1_out_full_final <- dudi.coa(tab(obj_out_full_final), scannf=FALSE, nf=10)
eig_ca1_out_full_final <- round((ca1_out_full_final$eig / (sum(ca1_out_full_final$eig))) * 100, 2)
out_ca_out_full_final <- ca1_out_full_final$li[,1:2]
rownames(out_ca_out_full_final) <- popNames(rad_out_full_final.filt)
```

```{r}
out_ca_out_full_final <- out_ca_out_full_final %>%
  rownames_to_column(var = "Population") %>%
  mutate(Population = factor(Population, levels = c("BAR", "BIS", "GB", "KIC", "MCD", "PVD"))) %>%
  dplyr::select(Population, Axis1, Axis2)
```

```{r}
strata_ca_out_full_final <- strata_6pops %>%
  left_join(out_ca_out_full_final, by = "Population")
```

```{r}
# Plot the CA
ggplot(strata_ca_out_full_final, aes(x = Axis1, y = Axis2, colour = Population)) +
  geom_point(size = 6, alpha = 0.8) +
  scale_colour_manual(values = cols_6pops,
                      name = "") +
  theme_minimal() +
  theme(plot.title = element_text(size = 14)) +
  labs(x = paste("", "Axis 1: ", eig_ca1_out_full_final[1], "%"), y = paste("", "Axis 2: ", eig_ca1_out_full_final[2], "%"))
```

```{r}
# Plot CA1 versus June Salinity Min
strata_ca_out_full_final %>%
  ggplot(aes(x = Axis1, y = June_Salinity_Min)) +
  geom_smooth(method = "lm", col = "darkgrey", lty = 1) +
  geom_point(aes(color = Population), size = 4) +
  scale_color_manual(values = cols_6pops) +
  theme_minimal() +
  labs(
       x = "Axis1",
       y = "Minimum Salinity")
```

```{r}
# Linear model for CA1 versus PC2
sal_min_mod_out_full_final <- lm(Axis1 ~ June_Salinity_Min, data = strata_ca_out_full_final)
summary(sal_min_mod_out_full_final)
```


## RDAForest

```{r}
#install.packages("gradientForest", repos="http://R-Forge.R-project.org")
#install.packages("clv")
#install.packages("sp")
#install.packages("/home/azyck/NB_capture_both/NB_ddhaplo/NB_ddhaplo_working/NB_PopSeaGen_Analysis_working/RDA-forest/RDAforest_2.5.1.tar.gz")
#install.packages("rnaturalearth")
```

```{r}
library(RDAforest)
library(gradientForest)
library(mapdata)
library(poppr)
```

### Data exploration

Check for missing data.

```{r}
sum(is.na(rad_out_db.filt$tab))
```

RDAforest doesn't handle missing data, so I'm going to impute missing data using mean allele frequencies

```{r}
gen_out_db.imp <- apply(rad_out_db.filt$tab, 2, function(x) replace(x, is.na(x), as.numeric(names(which.max(table(x))))))
sum(is.na(gen_out_db.imp)) # No NAs
```

Let’s compute genetic distances based on genotypic correlation, and make an unconstrained ordination:
```{r}
# distances:
cordist_out_db=1-cor(t(gen_out_db.imp))
# ordination:
ord_out_db=capscale(cordist_out_db~1)
```

```{r}
# Creating a population vector that I can add back in 
pop = c(rep("BAR", 10),rep("BIS", 10),rep("GB", 10),rep("KIC", 10),rep("MCD", 10), rep("PVD",10))
```


Proportion of variance explained by PCs (“MDS” is the mathematically identical to principal components so we will call them PCs):

```{r}
plot(ord_out_db$CA$eig/sum(ord_out_db$CA$eig),xlab="PC",ylab="proportion of variance explained")
```

Probably 2 leading principal components.
Extracting scores, plotting ordination:

```{r}
so_out_db=data.frame(scores(ord_out_db,scaling=1,display="sites"))
ggplot(so_out_db,aes(MDS1,MDS2,color=pop))+geom_point()+coord_equal()+theme_bw()
```

### Exploring Isolation By Distance (IBD)

```{r}
# converting lat, lon to great circle distances
GCD=gcd.dist(strata6[,3:4])
latlon.gcd=GCD[[1]]
distGCD=GCD[[2]]

plot(as.dist(cordist_out_db)~distGCD,pch=16,cex=0.6,col=rgb(0,0,0,alpha=0.2))
```

```{r}
protest(capscale(distGCD~1),capscale(cordist_out_db~1))
```

Not a strong correlation, so I don't need to worry about regressing the geographic information out of the genetic data.

### Exploratory RDA-forest analysis

We will use all predictors and more leading PCs (25, specified by the option pcs2keep=c(1:25)) that we will likely need, just to see how it looks. 

```{r}
env <- strata6_full[,5:44]
env <- as.matrix(env)
rownames(env) <- strata6_full$Individuals
```

```{r}
gf_out_db=makeGF(ord_out_db,env,pcs2keep=NULL)
```




## Functional annotation of outlier SNPs 

First checking for outlier SNPs that are exons and inside genes. Bed files containing locations of exons, genes, and coding sequences were created from a .gaf file for the eastern oyster. 

I'm going to use vcftools to parse out the outlier SNPs that are in exon, gene, and coding sequence regions. 

Full set of outlier SNPs 

58 SNPs in exon regions, 68 SNPs in gene regions, 30 SNPs in coding sequences 

```
# VCF file containing all exon regions
$ vcftools --vcf outlierloci6p9g9_red.recode.vcf --recode --recode-INFO-all --bed sorted.ref3.0.exon.sc.bed --out outlier_full_exon

output:
After filtering, kept 60 out of 60 Individuals
Outputting VCF file...
	Read 337295 BED file entries.
After filtering, kept 58 out of a possible 70 Sites
Run Time = 0.00 seconds
```

```
# VCF file containing all gene regions
$ vcftools --vcf outlierloci6p9g9_red.recode.vcf --recode --recode-INFO-all --bed sorted.ref3.0.gene.sc.bed --out outlier_full_gene

output:
After filtering, kept 60 out of 60 Individuals
Outputting VCF file...
	Read 35016 BED file entries.
After filtering, kept 68 out of a possible 70 Sites
Run Time = 0.00 seconds
```

```
# VCF file containing all coding sequences 
$ vcftools --vcf outlierloci6p9g9_red.recode.vcf --recode --recode-INFO-all --bed sorted.ref3.0.CDS.sc.bed --out outlier_full_CDS

output:
After filtering, kept 60 out of 60 Individuals
Outputting VCF file...
	Read 295525 BED file entries.
After filtering, kept 30 out of a possible 70 Sites
Run Time = 0.00 seconds
```

Identified chromosome and position info for the SNPs in genes.

```
# full outliers in genes
# Convert SNP indices to locus position and chromosome info
$ mawk '!/#/' outlier_full_gene.recode.vcf | cut -f1,2 > outlier_full_gene_loci
$ NUM=(`cat outlier_full_gene_loci | wc -l`)
$ paste <(seq 1 $NUM) outlier_full_gene_loci > loci_outlier_full_gene.plus.index 
```

```
$ head loci_outlier_full_gene.plus.index

output:
1	NC_035789.1	8465456
2	NC_035789.1	8465500
3	NC_035789.1	8465519
4	NC_035789.1	8466007
5	NC_035780.1	8748233
6	NC_035780.1	13250687
7	NC_035780.1	24512990
8	NC_035780.1	29901078
9	NC_035780.1	29902142
10	NC_035780.1	63719373
```

From this list, I made a full dataset of the gene IDs and proteins. I also marked which SNPs were in exons and coding sequences. 

```
# full outliers in CDS
# Convert SNP indices to locus position and chromosome info
$ mawk '!/#/' outlier_full_CDS.recode.vcf | cut -f1,2 > outlier_full_CDS_loci
$ NUM=(`cat outlier_full_CDS_loci | wc -l`)
$ paste <(seq 1 $NUM) outlier_full_CDS_loci > loci_outlier_full_CDS.plus.index 
```

Updated full set of outliers (361)

282 SNPs in exon regions, 342 SNPs in gene regions, 186 SNPs in coding sequences 

```
# VCF file containing all exon regions
$ vcftools --vcf outlierloci_new.recode.vcf --recode --recode-INFO-all --bed sorted.ref3.0.exon.sc.bed --out outlier_full_new_exon

output:
After filtering, kept 60 out of 60 Individuals
Outputting VCF file...
	Read 337295 BED file entries.
After filtering, kept 282 out of a possible 361 Sites
Run Time = 0.00 seconds
```

```
# VCF file containing all gene regions
$ vcftools --vcf outlierloci_new.recode.vcf --recode --recode-INFO-all --bed sorted.ref3.0.gene.sc.bed --out outlier_full_new_gene

output:
After filtering, kept 60 out of 60 Individuals
Outputting VCF file...
	Read 35016 BED file entries.
After filtering, kept 342 out of a possible 361 Sites
Run Time = 0.00 seconds
```

```
# VCF file containing all coding sequences 
$ vcftools --vcf outlierloci_new.recode.vcf --recode --recode-INFO-all --bed sorted.ref3.0.CDS.sc.bed --out outlier_full_new_CDS

output:
After filtering, kept 60 out of 60 Individuals
Outputting VCF file...
	Read 295525 BED file entries.
After filtering, kept 186 out of a possible 361 Sites
Run Time = 1.00 seconds
```

pcadapt outlier SNPs 

50 SNPs in exon regions, 56 SNPs in gene regions, 25 SNPs in coding sequences 

```
# VCF file containing all exon regions
$ vcftools --vcf outlierloci6p9g9_db.recode.vcf --recode --recode-INFO-all --bed sorted.ref3.0.exon.sc.bed --out outlier_db_exon

output:
After filtering, kept 60 out of 60 Individuals
Outputting VCF file...
	Read 337295 BED file entries.
After filtering, kept 50 out of a possible 58 Sites
Run Time = 0.00 seconds
```

```
# VCF file containing all gene regions
$ vcftools --vcf outlierloci6p9g9_db.recode.vcf --recode --recode-INFO-all --bed sorted.ref3.0.gene.sc.bed --out outlier_db_gene

output:
After filtering, kept 60 out of 60 Individuals
Outputting VCF file...
	Read 35016 BED file entries.
After filtering, kept 56 out of a possible 58 Sites
Run Time = 0.00 seconds
```

```
# VCF file containing all coding sequences 
$ vcftools --vcf outlierloci6p9g9_db.recode.vcf --recode --recode-INFO-all --bed sorted.ref3.0.CDS.sc.bed --out outlier_db_CDS

output:
After filtering, kept 60 out of 60 Individuals
Outputting VCF file...
	Read 295525 BED file entries.
After filtering, kept 25 out of a possible 58 Sites
Run Time = 0.00 seconds
```

pcadapt outlier SNPs inside inversions 

31 SNPs in exon regions, 35 SNPs in gene regions, 8 SNPs in coding sequences 

```
# VCF file containing all exon regions
$ vcftools --vcf outliers6p9g9_db_invert.recode.vcf --recode --recode-INFO-all --bed sorted.ref3.0.exon.sc.bed --out outlier_db_invert_exon

output:
After filtering, kept 60 out of 60 Individuals
Outputting VCF file...
	Read 337295 BED file entries.
After filtering, kept 31 out of a possible 36 Sites
Run Time = 0.00 seconds
```

```
# VCF file containing all gene regions
$ vcftools --vcf outliers6p9g9_db_invert.recode.vcf --recode --recode-INFO-all --bed sorted.ref3.0.gene.sc.bed --out outlier_db_invert_gene

output:
After filtering, kept 60 out of 60 Individuals
Outputting VCF file...
	Read 35016 BED file entries.
After filtering, kept 35 out of a possible 36 Sites
Run Time = 0.00 seconds
```

```
# VCF file containing all coding sequences 
$ vcftools --vcf outliers6p9g9_db_invert.recode.vcf --recode --recode-INFO-all --bed sorted.ref3.0.CDS.sc.bed --out outlier_db_invert_CDS

output:
After filtering, kept 60 out of 60 Individuals
Outputting VCF file...
	Read 295525 BED file entries.
After filtering, kept 8 out of a possible 36 Sites
Run Time = 0.00 seconds
```

pcadapt outlier SNPs outside inversions 

19 SNPs in exon regions, 21 SNPs in gene regions, 17 SNPs in coding sequences 

```
# VCF file containing all exon regions
$ vcftools --vcf outliers6p9g9_db_noinvert.recode.vcf --recode --recode-INFO-all --bed sorted.ref3.0.exon.sc.bed --out outlier_db_noinvert_exon

output:
After filtering, kept 60 out of 60 Individuals
Outputting VCF file...
	Read 337295 BED file entries.
After filtering, kept 19 out of a possible 22 Sites
Run Time = 0.00 seconds
```

```
# VCF file containing all gene regions
$ vcftools --vcf outliers6p9g9_db_noinvert.recode.vcf --recode --recode-INFO-all --bed sorted.ref3.0.gene.sc.bed --out outlier_db_noinvert_gene

output:
After filtering, kept 60 out of 60 Individuals
Outputting VCF file...
	Read 35016 BED file entries.
After filtering, kept 21 out of a possible 22 Sites
Run Time = 0.00 seconds
```

```
# VCF file containing all coding sequences 
$ vcftools --vcf outliers6p9g9_db_noinvert.recode.vcf --recode --recode-INFO-all --bed sorted.ref3.0.CDS.sc.bed --out outlier_db_noinvert_CDS

output:
After filtering, kept 60 out of 60 Individuals
Outputting VCF file...
	Read 295525 BED file entries.
After filtering, kept 17 out of a possible 22 Sites
Run Time = 0.00 seconds
```


LFMM outlier SNPs 

10 SNPs in exon regions, 14 SNPs in gene regions, 5 SNPs in coding sequences 

```
# VCF file containing all exon regions
$ vcftools --vcf outlierloci6p9g9_lfmm.recode.vcf --recode --recode-INFO-all --bed sorted.ref3.0.exon.sc.bed --out outlier_lfmm_exon

output:
After filtering, kept 60 out of 60 Individuals
Outputting VCF file...
	Read 337295 BED file entries.
After filtering, kept 10 out of a possible 14 Sites
Run Time = 0.00 seconds
```

```
# VCF file containing all gene regions
$ vcftools --vcf outlierloci6p9g9_lfmm.recode.vcf --recode --recode-INFO-all --bed sorted.ref3.0.gene.sc.bed --out outlier_lfmm_gene

output:
After filtering, kept 60 out of 60 Individuals
Outputting VCF file...
	Read 35016 BED file entries.
After filtering, kept 14 out of a possible 14 Sites
Run Time = 0.00 seconds
```

```
# VCF file containing all coding sequences 
$ vcftools --vcf outlierloci6p9g9_lfmm.recode.vcf --recode --recode-INFO-all --bed sorted.ref3.0.CDS.sc.bed --out outlier_lfmm_CDS

output:
After filtering, kept 60 out of 60 Individuals
Outputting VCF file...
	Read 295525 BED file entries.
After filtering, kept 5 out of a possible 14 Sites
Run Time = 0.00 seconds
```


LFMM outlier SNPs inside inversions 

6 SNPs in exon regions, 6 SNPs in gene regions, 2 SNPs in coding sequences 

```
# VCF file containing all exon regions
$ vcftools --vcf outliers6p9g9_lfmm_invert.recode.vcf --recode --recode-INFO-all --bed sorted.ref3.0.exon.sc.bed --out outlier_lfmm_invert_exon

output:
After filtering, kept 60 out of 60 Individuals
Outputting VCF file...
	Read 337295 BED file entries.
After filtering, kept 6 out of a possible 6 Sites
Run Time = 0.00 seconds
```

```
# VCF file containing all gene regions
$ vcftools --vcf outliers6p9g9_lfmm_invert.recode.vcf --recode --recode-INFO-all --bed sorted.ref3.0.gene.sc.bed --out outlier_lfmm_invert_gene

output:
After filtering, kept 60 out of 60 Individuals
Outputting VCF file...
	Read 35016 BED file entries.
After filtering, kept 6 out of a possible 6 Sites
Run Time = 0.00 seconds
```

```
# VCF file containing all coding sequences 
$ vcftools --vcf outliers6p9g9_lfmm_invert.recode.vcf --recode --recode-INFO-all --bed sorted.ref3.0.CDS.sc.bed --out outlier_lfmm_invert_CDS

output:
After filtering, kept 60 out of 60 Individuals
Outputting VCF file...
	Read 295525 BED file entries.
After filtering, kept 2 out of a possible 6 Sites
Run Time = 1.00 seconds
```


LFMM outlier SNPs outside inversions 

4 SNPs in exon regions, 8 SNPs in gene regions, 3 SNPs in coding sequences 

```
# VCF file containing all exon regions
$ vcftools --vcf outliers6p9g9_lfmm_noinvert.recode.vcf --recode --recode-INFO-all --bed sorted.ref3.0.exon.sc.bed --out outlier_lfmm_noinvert_exon

output:
After filtering, kept 60 out of 60 Individuals
Outputting VCF file...
	Read 337295 BED file entries.
After filtering, kept 4 out of a possible 8 Sites
Run Time = 0.00 seconds
```

```
# VCF file containing all gene regions
$ vcftools --vcf outliers6p9g9_lfmm_noinvert.recode.vcf --recode --recode-INFO-all --bed sorted.ref3.0.gene.sc.bed --out outlier_lfmm_noinvert_gene

output:
After filtering, kept 60 out of 60 Individuals
Outputting VCF file...
	Read 35016 BED file entries.
After filtering, kept 8 out of a possible 8 Sites
Run Time = 0.00 seconds
```

```
# VCF file containing all coding sequences 
$ vcftools --vcf outliers6p9g9_lfmm_noinvert.recode.vcf --recode --recode-INFO-all --bed sorted.ref3.0.CDS.sc.bed --out outlier_lfmm_noinvert_CDS

output:
After filtering, kept 60 out of 60 Individuals
Outputting VCF file...
	Read 295525 BED file entries.
After filtering, kept 3 out of a possible 8 Sites
Run Time = 0.00 seconds
```

RDA outliers (291)

224 SNPs in exon regions, 274 SNPs in gene regions, 156 SNPs in coding sequences 

```
# VCF file containing all exon regions
$ vcftools --vcf outliers_rda.recode.vcf --recode --recode-INFO-all --bed sorted.ref3.0.exon.sc.bed --out outlier_rda_exon

output:
After filtering, kept 60 out of 60 Individuals
Outputting VCF file...
	Read 337295 BED file entries.
After filtering, kept 224 out of a possible 291 Sites
Run Time = 0.00 seconds
```

```
# VCF file containing all gene regions
$ vcftools --vcf outliers_rda.recode.vcf --recode --recode-INFO-all --bed sorted.ref3.0.gene.sc.bed --out outlier_rda_gene

output:
After filtering, kept 60 out of 60 Individuals
Outputting VCF file...
	Read 35016 BED file entries.
After filtering, kept 274 out of a possible 291 Sites
Run Time = 0.00 seconds
```

```
# VCF file containing all coding sequences 
$ vcftools --vcf outliers_rda.recode.vcf --recode --recode-INFO-all --bed sorted.ref3.0.CDS.sc.bed --out outlier_rda_CDS

output:
After filtering, kept 60 out of 60 Individuals
Outputting VCF file...
	Read 295525 BED file entries.
After filtering, kept 156 out of a possible 291 Sites
Run Time = 0.00 seconds
```

Full final outlier set (390 SNPs):

299 SNPs in exon regions, 368 SNPs in gene regions, 195 SNPs in coding sequences 

```
# VCF file containing all exon regions
$ vcftools --vcf outliers_full_final.recode.vcf --recode --recode-INFO-all --bed sorted.ref3.0.exon.sc.bed --out outlier_full_final_exon

output:
After filtering, kept 60 out of 60 Individuals
Outputting VCF file...
	Read 337295 BED file entries.
After filtering, kept 299 out of a possible 390 Sites
Run Time = 0.00 seconds
```

```
# VCF file containing all gene regions
$ vcftools --vcf outliers_full_final.recode.vcf --recode --recode-INFO-all --bed sorted.ref3.0.gene.sc.bed --out outlier_full_final_gene

output:
After filtering, kept 60 out of 60 Individuals
Outputting VCF file...
	Read 35016 BED file entries.
After filtering, kept 368 out of a possible 390 Sites
Run Time = 1.00 seconds
```

```
# VCF file containing all coding sequences 
$ vcftools --vcf outliers_full_final.recode.vcf --recode --recode-INFO-all --bed sorted.ref3.0.CDS.sc.bed --out outlier_full_final_CDS

output:
After filtering, kept 60 out of 60 Individuals
Outputting VCF file...
	Read 295525 BED file entries.
After filtering, kept 195 out of a possible 390 Sites
Run Time = 0.00 seconds
```

I'm going to save the genes that correspond to the outlier loci. I want to see how many unique genes there are are and how many overlap across the three outlier detection methods. The list outlier SNP chromosomes and positions for each program are saved in `outlier_pcadapt_thinned500.loci.txt`, `outlier_lfmm2_6p9g9_red.loci.txt`, `outliers_rda_fullneut.loci.txt`. I need to first convert these .txt files to .bed files

```
$ cut -f1,2 outlier_pcadapt_thinned500.loci.txt | sort -k1,2 | uniq | mawk '{print $1 "\t" $2-1 "\t" $2}' | bedtools sort -i - > Pcadapt.Cand.Loci.bed
$ cut -f1,2 outlier_lfmm2_6p9g9_red.loci.txt | sort -k1,2 | uniq | mawk '{print $1 "\t" $2-1 "\t" $2}' | bedtools sort -i - > LFMM.Cand.Loci.bed
$ cut -f1,2 outliers_rda_fullneut.loci.txt | sort -k1,2 | uniq | mawk '{print $1 "\t" $2-1 "\t" $2}' | bedtools sort -i - > RDA.Cand.Loci.bed
```

I'll then intersect these .bed files with the bed file containing the gene information for the whole genome:

```
$ bedtools intersect -wb -a Pcadapt.Cand.Loci.bed -b sorted.ref3.0.gene.bed | grep -oh "gene=LOC.*;g" | sed 's/gene=//g' | sed 's/;g//g' | sort | uniq > Pcadapt.Genes
$ bedtools intersect -wb -a LFMM.Cand.Loci.bed -b sorted.ref3.0.gene.bed | grep -oh "gene=LOC.*;g" | sed 's/gene=//g' | sed 's/;g//g' | sort | uniq > LFMM.Genes
$ bedtools intersect -wb -a RDA.Cand.Loci.bed -b sorted.ref3.0.gene.bed | grep -oh "gene=LOC.*;g" | sed 's/gene=//g' | sed 's/;g//g' | sort | uniq > RDA.Genes
```

This gives me a separate count of unique genes for each program. I'm not going to combine these three files to make a file with all unique outlier genes. 

```
$ cat Pcadapt.Genes LFMM.Genes RDA.Genes | sort -u > All.Outlier.Genes
$ cat All.Outlier.Genes | wc -l

272
```

When I added the number of genes together for each individual program, it summed to 282 which means there is some overlap in genes for each program. I want to see which and how many overlap. First, I need to sort tthe gene lists for each program:

```
$ sort Pcadapt.Genes > sorted.pcadapt
$ sort LFMM.Genes > sorted.lfmm
$ sort RDA.Genes > sorted.rda
```

Now looking for entries that overlap across each file

```
# pcadapt and lfmm
$ comm -12 sorted.pcadapt sorted.lfmm
LOC111099610
LOC111135070

# pcadapt and rda
$ comm -12 sorted.pcadapt sorted.rda
LOC111101025
LOC111101168
LOC111135070
LOC111138046

# lfmm and rda
$ comm -12 sorted.lfmm sorted.rda
LOC111110485
LOC111130378
LOC111131017
LOC111132962
LOC111135070
```

One gene, LOC111135070, is shared across all three programs. 


Now, I'm going to split the outliers up into a set of inversion SNPs and non-inversion SNPs. Then I'm going to repeat the same steps above. This is bit repetitive, but I'm going to remake the bed files and then intersect with the bed file that has the information for the large inversions. Then I will intersect the invert loci with the genes 

```
$ awk '{print $1 "\t" $2 "\t" ($2+1)}' outlier_pcadapt_thinned500.loci.txt > pcadapt_outliers.bed
$ bedtools intersect -a pcadapt_outliers.bed -b Detected_large_inversions.bed -wa > pcadapt_invert
$ bedtools intersect -wb -a pcadapt_invert -b sorted.ref3.0.gene.bed | grep -oh "gene=LOC.*;g" | sed 's/gene=//g' | sed 's/;g//g' | sort | uniq > Pcadapt.Invert.Genes

$ awk '{print $1 "\t" $2 "\t" ($2+1)}' outlier_lfmm2_6p9g9_red.loci.txt > lfmm_outliers.bed
$ bedtools intersect -a lfmm_outliers.bed -b Detected_large_inversions.bed -wa > lfmm_invert
$ bedtools intersect -wb -a lfmm_invert -b sorted.ref3.0.gene.bed | grep -oh "gene=LOC.*;g" | sed 's/gene=//g' | sed 's/;g//g' | sort | uniq > LFMM.Invert.Genes

$ awk '{print $1 "\t" $2 "\t" ($2+1)}' outliers_rda_fullneut.loci.txt > rda_outliers.bed
$ bedtools intersect -a rda_outliers.bed -b Detected_large_inversions.bed -wa > rda_invert
$ bedtools intersect -wb -a rda_invert -b sorted.ref3.0.gene.bed | grep -oh "gene=LOC.*;g" | sed 's/gene=//g' | sed 's/;g//g' | sort | uniq > RDA.Invert.Genes

$ cat Pcadapt.Invert.Genes LFMM.Invert.Genes RDA.Invert.Genes | sort -u > All.Invert.Outlier.Genes
$ $ cat All.Invert.Outlier.Genes | wc -l
35
```

Sort and look for overlapped genes across the three programs 

```
$ sort Pcadapt.Invert.Genes > sorted.pcadapt.invert
$ sort LFMM.Invert.Genes > sorted.lfmm.invert
$ sort RDA.Invert.Genes > sorted.rda.invert


$ comm -12 sorted.pcadapt.invert sorted.lfmm.invert
LOC111099610
LOC111135070

$ comm -12 sorted.pcadapt.invert sorted.rda.invert
LOC111101025
LOC111101168
LOC111135070
LOC111138046

$ comm -12 sorted.lfmm.invert sorted.rda.invert
LOC111132962
LOC111135070
```

One gene, LOC111135070, is shared across all three programs. 


Repeating for the non-inversion SNPs

```
$ bedtools subtract -a pcadapt_outliers.bed -b Detected_large_inversions.bed > pcadapt_noinvert 
$ bedtools intersect -wb -a pcadapt_noinvert -b sorted.ref3.0.gene.bed | grep -oh "gene=LOC.*;g" | sed 's/gene=//g' | sed 's/;g//g' | sort | uniq > Pcadapt.NoInvert.Genes

$ bedtools subtract -a lfmm_outliers.bed -b Detected_large_inversions.bed > lfmm_noinvert
$ bedtools intersect -wb -a lfmm_noinvert -b sorted.ref3.0.gene.bed | grep -oh "gene=LOC.*;g" | sed 's/gene=//g' | sed 's/;g//g' | sort | uniq > LFMM.NoInvert.Genes

$ bedtools subtract -a rda_outliers.bed -b Detected_large_inversions.bed > rda_noinvert
$ bedtools intersect -wb -a rda_noinvert -b sorted.ref3.0.gene.bed | grep -oh "gene=LOC.*;g" | sed 's/gene=//g' | sed 's/;g//g' | sort | uniq > RDA.NoInvert.Genes

$ cat Pcadapt.NoInvert.Genes LFMM.NoInvert.Genes RDA.NoInvert.Genes | sort -u > All.NoInvert.Outlier.Genes
$ cat All.NoInvert.Outlier.Genes | wc -l
237
```

```
$ sort LFMM.NoInvert.Genes > sorted.lfmm.noinvert
$ sort RDA.NoInvert.Genes > sorted.rda.noinvert
$ sort Pcadapt.NoInvert.Genes > sorted.pcadapt.noinvert

$ comm -12 sorted.pcadapt.noinvert sorted.lfmm.noinvert 
$ comm -12 sorted.pcadapt.noinvert sorted.rda.noinvert 
$ comm -12 sorted.lfmm.noinvert sorted.rda.noinvert 
LOC111110485
LOC111130378
LOC111131017
```

Only three genes overlap between RDA and LFMM. No genes overlap for all three programs. 



## Gene Ontology with TopGO

I have a gaf from NCBI with the GO terms for genes in the eastern oyster genome. I'm going to take this file and extract the Gene IDs and associated GO terms.

In terminal:

```
$ wget https://ftp.ncbi.nlm.nih.gov/genomes/all/GCF/002/022/765/GCF_002022765.2_C_virginica-3.0/GCF_002022765.2_C_virginica-3.0_gene_ontology.gaf.gz

$ gunzip GCF_002022765.2_C_virginica-3.0_gene_ontology.gaf.gz

$ mv GCF_002022765.2_C_virginica-3.0_gene_ontology.gaf gene_ontology.gaf
```

Making a new .tab file with just the gene ID and GO IDs (3rd and 5th columns) renamed "gene_name" and "GeneOntologyIDs". The gene_name column will have one unique observation per row and all associated GO terms will be saved in the same row, with values separaetd by ;

```
$ awk 'NR>9 {print $3 "\t" $5}' gene_ontology.gaf | 
> sed '1i gene_name\tGeneOntologyIDs' | 
> sort | awk '
>     NR==1 {print; next}
>     $1!=prev {
>         if (NR>2) print prev "\t" goids;
>         prev=$1;
>         goids=$2;
>         next
>     }
>     {
>         goids=goids "; " $2
>     }
>     END {print prev "\t" goids}
> ' > gene2go_full.tab
```

Gene IDs for the candidate loci were extracted using a bed file containing all genes. The candidate gene IDs are saved to `Sig.loci.1.LOC`

```
$ cut -f1,2 all.new.outliers | sort -k1,2 | uniq | mawk '{print $1 "\t" $2-1 "\t" $2}' | bedtools sort -i - > Full.Cand.Loci.bed

$ bedtools intersect -wb -a Full.Cand.Loci.bed -b sorted.ref3.0.gene.bed | grep -oh "gene=LOC.*;g" | sed 's/gene=//g' | sed 's/;g//g' | sort | uniq > Sig.loci.1.LOC
```

I'm making another file with the gene IDs and GO IDs for the candidate loci:

```
$ grep -F -f Sig.loci.1.LOC gene2go_full.tab > cand_genes_GO.tab

# Add header
$ awk 'BEGIN {print "gene_name\tGeneOntologyIDs"} {print}' cand_genes_GO.tab > cand_genes_GO_head.tab
```

```
$ head cand_genes_GO_head.tab 
gene_name	GeneOntologyIDs
LOC111099053	GO:0003743; GO:0006413
LOC111099197	GO:0030246
LOC111100382	GO:0008168; GO:0010468; GO:0016021; GO:0032259; GO:0046872; GO:0071557; GO:0071558
LOC111100601	GO:0016020; GO:0016125; GO:0016705; GO:0046872
LOC111100630	GO:0005890; GO:0006813; GO:0006814; GO:0006875; GO:0030004; GO:0098662
LOC111100706	GO:0004497; GO:0005506; GO:0016705; GO:0020037; GO:0055114
LOC111101025	GO:0003743; GO:0004004; GO:0005524; GO:0006413
LOC111101041	GO:0005524; GO:0015991; GO:0016787; GO:0033180; GO:0046034
LOC111101168	GO:0003341; GO:0005930; GO:0036159; GO:0060285
```


In R:

```{r}
#if (!require("BiocManager", quietly = TRUE))
    #install.packages("BiocManager")

#BiocManager::install("topGO")
```

```{r setup, include=FALSE}
library(topGO)
```

Read in gene2go information 

Old file names: all_genes (gene2go_full.tab); candidate genes (gene2go.tab)

```{r}
all_gene2go<-read.delim("gene2go_full.tab", sep="\t")
out_gene2go<-read.delim("cand_genes_GO_rda_head.tab", sep="\t")
```

Make list of genes for input to TopGO

```{r}
cand_genes <- as.character(unique(out_gene2go$gene_name))
all_genes <- as.character(all_gene2go$gene_name)
out_GeneList <- factor(as.integer(all_genes %in% cand_genes))
names(out_GeneList) <- all_genes
```

The following code will perform GO enrichment using the weighted Fisher's exact test  

Read in gene-to-go-mappings
```{r}
gene2go_topgo<-readMappings("gene2go_full.tab", IDsep=";", sep="\t")
```

Set function to select genes of interest (ie those that have pvalue < 0.05)
```{r}
topDiffGenes <- function(allScore) {
return(allScore < 0.05)}
```

### Biological Processes

Create `topGOdata` object, which is required for topGO analysis
```{r}
GO_OUT_BP_pi <-new("topGOdata", ontology="BP", gene2GO=gene2go_topgo, allGenes=out_GeneList, annot = annFUN.gene2GO, geneSel=topDiffGenes)
```

Run GO enrichment test 
```{r}
GO_OUT_BP_FE_pi <- runTest(GO_OUT_BP_pi, algorithm="weight01", statistic="fisher")
```

Generate results table 
```{r}
GO_OUT_BP_En_pi <- GenTable(GO_OUT_BP_pi, Fisher = GO_OUT_BP_FE_pi, orderBy = "Fisher",  topNodes = 100, numChar = 100)
```
Only taking the top 100 GO terms 

Filter by significant results
```{r}
GO_OUT_BP_En_pi$Fisher<-as.numeric(GO_OUT_BP_En_pi$Fisher)
GO_OUT_BP_En_sig_pi<-GO_OUT_BP_En_pi[GO_OUT_BP_En_pi$Fisher<0.05,]
```

Merge `GO_OUT_BP_En_sig_pi` with GO and gene info. 
```{r}
# Separate GO terms 
out_gene2go <- out_gene2go %>%
  separate_rows(GeneOntologyIDs, sep = ";")

# Ensure GO terms in both datasets are formatted similarly (trim whitespaces)
out_gene2go$GeneOntologyIDs <- trimws(out_gene2go$GeneOntologyIDs)
GO_OUT_BP_En_sig_pi$GO.ID <- trimws(GO_OUT_BP_En_sig_pi$GO.ID)

# Join the datasets based on GO term
GO_OUT_BP_En_sig_gene_pi <- out_gene2go %>%
  left_join(GO_OUT_BP_En_sig_pi, by = c("GeneOntologyIDs" = "GO.ID")) %>%
  na.omit()

# Add ontology column 
GO_OUT_BP_En_sig_gene_pi$ontology <- "Biological Processes"
```

### Cellular Components

Create `topGOdata` object, which is required for topGO analysis
```{r}
GO_OUT_CC_pi <-new("topGOdata", ontology="CC", gene2GO=gene2go_topgo, allGenes=out_GeneList, annot = annFUN.gene2GO, geneSel=topDiffGenes)
```

Run GO enrichment test 
```{r}
GO_OUT_CC_FE_pi <- runTest(GO_OUT_CC_pi, algorithm="weight01", statistic="fisher")
```

Generate results table 
```{r}
GO_OUT_CC_En_pi <- GenTable(GO_OUT_CC_pi, Fisher = GO_OUT_CC_FE_pi, orderBy = "Fisher",  topNodes = 100, numChar = 100)
```
Only taking the top 100 GO terms 

Filter by significant results
```{r}
GO_OUT_CC_En_pi$Fisher<-as.numeric(GO_OUT_CC_En_pi$Fisher)
GO_OUT_CC_En_sig_pi<-GO_OUT_CC_En_pi[GO_OUT_CC_En_pi$Fisher<0.05,]
```

Merge `GO_OUT_CC_En_sig_pi` with GO and gene info. 
```{r}
# Separate GO terms 
out_gene2go <- out_gene2go %>%
  separate_rows(GeneOntologyIDs, sep = ";")

# Ensure GO terms in both datasets are formatted similarly (trim whitespaces)
out_gene2go$GeneOntologyIDs <- trimws(out_gene2go$GeneOntologyIDs)
GO_OUT_CC_En_sig_pi$GO.ID <- trimws(GO_OUT_CC_En_sig_pi$GO.ID)

# Join the datasets based on GO term
GO_OUT_CC_En_sig_gene_pi <- out_gene2go %>%
  left_join(GO_OUT_CC_En_sig_pi, by = c("GeneOntologyIDs" = "GO.ID")) %>%
  na.omit()

# Add ontology column 
GO_OUT_CC_En_sig_gene_pi$ontology <- "Cellular Components"
```

### Molecular Functions

Create `topGOdata` object, which is required for topGO analysis
```{r}
GO_OUT_MF_pi <-new("topGOdata", ontology="MF", gene2GO=gene2go_topgo, allGenes=out_GeneList, annot = annFUN.gene2GO, geneSel=topDiffGenes)
```

Run GO enrichment test 
```{r}
GO_OUT_MF_FE_pi <- runTest(GO_OUT_MF_pi, algorithm="weight01", statistic="fisher")
```

Generate results table 
```{r}
GO_OUT_MF_En_pi <- GenTable(GO_OUT_MF_pi, Fisher = GO_OUT_MF_FE_pi, orderBy = "Fisher",  topNodes = 100, numChar = 100)
```
Only taking the top 100 GO terms 

Filter by significant results
```{r}
GO_OUT_MF_En_pi$Fisher<-as.numeric(GO_OUT_MF_En_pi$Fisher)
GO_OUT_MF_En_sig_pi<-GO_OUT_MF_En_pi[GO_OUT_MF_En_pi$Fisher<0.05,]
```

Merge `GO_OUT_MF_En_sig_pi` with GO and gene info. 
```{r}
# Separate GO terms 
out_gene2go <- out_gene2go %>%
  separate_rows(GeneOntologyIDs, sep = ";")

# Ensure GO terms in both datasets are formatted similarly (trim whitespaces)
out_gene2go$GeneOntologyIDs <- trimws(out_gene2go$GeneOntologyIDs)
GO_OUT_MF_En_sig_pi$GO.ID <- trimws(GO_OUT_MF_En_sig_pi$GO.ID)

# Join the datasets based on GO term
GO_OUT_MF_En_sig_gene_pi <- out_gene2go %>%
  left_join(GO_OUT_MF_En_sig_pi, by = c("GeneOntologyIDs" = "GO.ID")) %>%
  na.omit()

# Add ontology column 
GO_OUT_MF_En_sig_gene_pi$ontology <- "Molecular Functions"
```

### Join ontologies

Bind `GO_OUT_BP_En_sig_gene_pi`, `GO_OUT_CC_En_sig_gene_pi`, and `GO_OUT_MF_En_sig_gene_pi` to make a df that has significantly enriched GO terms for all ontologies 
```{r}
GO_OUT_En_sig_gene_pi <- rbind(GO_OUT_BP_En_sig_gene_pi, GO_OUT_CC_En_sig_gene_pi, GO_OUT_MF_En_sig_gene_pi)
```

Merge `pi_bed_filt` with `GO_PEVE_En_sig_gene` and calculate proportion of significant genes
```{r}
merged_out_genes <- GO_OUT_En_sig_gene_pi %>%
  mutate(prop.sig.genes = Significant/Annotated)

# Write to csv 
write.csv(merged_out_genes, "OUT_GO_en_sig_gene_rda.csv")
```

Plot - need to work out 
```{r}
GO_plot_pi<-ggplot(merged_out_genes, aes(x = Term, y = -log(Fisher), size = prop.sig.genes, fill = -log(Fisher))) +
  expand_limits(y = 1.5) +
  ylim(1, 7.25) +
  #Add horizontal lines with a single aesthetic value
  geom_hline(yintercept = -log10(0.01), linetype = "longdash", colour = "black", linewidth = .6) +
  geom_hline(yintercept = -log10(0.001), linetype = "solid", colour = "black", linewidth = .6) +
  geom_point(shape = 21) + 
  scale_size(range = c(2, 12)) + 
  scale_fill_continuous(low = "#1AD3D1FF", high = "#4686FBFF") +
  xlab('') + 
  ylab('Enrichment score') +
  #labs(caption = 'Cut-off lines at p=0.01 and p=0.001') +
  theme_bw() +
  facet_grid(vars(ontology), scales = "free", space = "free_y") +
  coord_flip(); GO_plot_pi
```

I wanted to look at GO terms that overlapped across programs split up into inversion and non-inversion datasets. In Excel extracted the GO terms for each program, saved the unique terms and sorted them. I then saved each list as a txt file in terminal and compared overlapping GO terms across programs. 

For the inversion dataset:

```
$ cat Pcadapt.Invert.GO.sorted | wc -l
28

$ cat LFMM.Invert.GO.sorted | wc -l
9

$ cat RDA.Invert.GO.sorted | wc -l
11

$ comm -12 Pcadapt.Invert.GO.sorted LFMM.Invert.GO.sorted 
GO:0010971
GO:0035329
GO:0110032
$ comm -12 Pcadapt.Invert.GO.sorted RDA.Invert.GO.sorted 
GO:0031213
$ comm -12 LFMM.Invert.GO.sorted RDA.Invert.GO.sorted 

$ cat Pcadapt.Invert.GO.sorted LFMM.Invert.GO.sorted RDA.Invert.GO.sorted | sort -u > All.Invert.Outlier.GO
$ cat All.Invert.Outlier.GO | wc -l
44
```


For the non-inversion dataset:

```
$ cat Pcadapt.NoInvert.GO.sorted | wc -l
12

$ cat LFMM.NoInvert.GO.sorted | wc -l
6

$ cat RDA.NoInvert.GO.sorted | wc -l
104
```

```
$ comm -12 Pcadapt.NoInvert.GO.sorted LFMM.NoInvert.GO.sorted 
$ comm -12 Pcadapt.NoInvert.GO.sorted RDA.NoInvert.GO.sorted 
$ comm -12 LFMM.NoInvert.GO.sorted RDA.NoInvert.GO.sorted 
GO:0004565

$ cat Pcadapt.NoInvert.GO.sorted LFMM.NoInvert.GO.sorted RDA.NoInvert.GO.sorted | sort -u > All.NoInvert.Outlier.GO
$ cat All.NoInvert.Outlier.GO | wc -l
121
```

Only one GO term overlaps between the LFMM and RDA dataset. 

I also looked at the number of unique genes corresponding to the GO terms for the inversion and non-inversion datasets. There were 19 unique genes that were annotated with GO terms for the inversion datast. 
