---
title: "NB Population Genomic Analysis on Non-Haplotig Masked Neutral SNPs"
author: "Amy Zyck"
date: "5/27/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Population analyses on Neutral SNP dataset 

I first created a new directory `NB_PopGen_Neutral` within `PATH: /home/azyck/NB_capture/NB_ddocent`. I linked `neutralloci.recode.vcf` to this working directory. The RMarkdown file should be saved in this same working directory. 

```{r}
#Loading all the necessary packages
library(adegenet)
library(vcfR)
library("radiator") # Conversion from vcf to a lot of other formats
library("dplyr")
library("hierfstat")
library("ggplot2") #For plotting
library("reshape2") #For plotting
library("plyr")
library("cowplot") #For plotting manuscript figs
library(PCAviz)  #Visualizing output of PCA
library("stringr")
library("bigsnpr") # package for Linkage Disequilibrium pruning
```

# All Neutral SNPs

## Making files

**Make genind object**

`neutralloci.recode.vcf` contains neutral SNPs for populations PVD, GB, BIS, and NAR. Steps for generating this VCF file are located in [EecSeq_Cvirginica_dDocent.md](https://github.com/amyzyck/EecSeq_NB_EasternOyster/blob/master/Analysis/EecSeq_Cvirginica_dDocent.md) and [EecSeq_Cvirginica_Filtering.md](https://github.com/amyzyck/EecSeq_NB_EasternOyster/blob/master/Analysis/EecSeq_Cvirginica_Filtering.md). Population NIN was removed from the VCF file following steps in [EecSeq_Cvirginica_OutlierDetection.md](https://github.com/amyzyck/EecSeq_NB_EasternOyster/blob/master/Analysis/EecSeq_Cvirginica_OutlierDetection.md).

`strata` contains population, environmental, and library information for each sample - can be accessed [here](https://github.com/amyzyck/EecSeq_NB_EasternOyster/blob/master/Analysis/PopGen_SeaGen_Analyses/strata).

```{r}
my_vcf <- read.vcfR("neutralloci.recode.vcf")
strata <- read.table("strata", header=TRUE)
```

```{r}
rad.filt <- vcfR2genind(my_vcf, strata = strata, pop = c(rep("BIS", 10),rep("GB", 10),rep("NAR", 10), rep("PVD",10)))
```

```{r}
rad.filt
```

```{r}
#Providing population names for plotting
pop_order <- c("BIS","GB","NAR","PVD")
```

**Read in the other info from .strata file and extract information such as locality, latitude, and longitude.**
```{r}
info <- as.data.frame(read.table("strata",header = T,sep = "\t",stringsAsFactors = F))
mystrats <- as.data.frame(matrix(nrow = length(indNames(rad.filt)),ncol=10))
colnames(mystrats) <- c("Population","Latitude","Longitude","Distance","SE","Temperature","Salinity","pH","Chlorophylla","DO")
just.strats <- select(info,c("Population"))
stratted.filt <- strata(rad.filt, formula= Population, combine = TRUE,just.strats)
stratted.filt@other <- select(info, Latitude,Longitude,Distance,SE,Temperature,Salinity,pH,Chlorophylla,DO)
```

```{r}
stratted.filt
```

**Repeat for quasi-independent set of SNPs**

Steps for filtering are documented for OutFLANK in [EecSeq_Cvirginica_OutlierDetection.md](https://github.com/amyzyck/EecSeq_NB_EasternOyster/blob/master/Analysis/EecSeq_Cvirginica_OutlierDetection.md). However, I am repeating the steps here. 

```{r}
my_vcf <- read.vcfR("neutralloci.recode.vcf")

geno <- extract.gt(my_vcf) # Character matrix containing the genotypes
position <- getPOS(my_vcf) # Positions in bp
chromosome <- getCHROM(my_vcf) # Chromosome information

G <- matrix(NA, nrow = nrow(geno), ncol = ncol(geno))

G[geno %in% c("0/0", "0|0")] <- 0
G[geno  %in% c("0/1", "1/0", "1|0", "0|1")] <- 1
G[geno %in% c("1/1", "1|1")] <- 2

# NA should be replaced with “9” to work with the functions in the OutFLANK package
G[is.na(G)] <- 9
```

Chromosomes need to be of class integer for this to work.
```{r}
# Visualizing chromosomes
chrom_unique <- unique(chromosome)
print(chrom_unique)
```

```{r}
# Removing "NC_" from the chromosome name so it can be converted to an integer
chrom_new <- chromosome%>%str_replace("NC_","")

# Converting the character string into an integer
chrom1 <- as.integer(chrom_new)
```

`chromosome` and `position` need to be sorted for imputation to work.

```{r}
chrom_sort <- sort(chrom1)
pos_sort <- sort(position)
```

_Note: This filtering program does not allow for missing genotype values._

I can either remove the missing genotype values or impute missing genotypes. I'm going to try both and see how the two methods differ during the trimming step. 

### Remove missing genotypes

```{r}
# removing missing data 
G_miss <- matrix(NA, nrow = nrow(geno), ncol = ncol(geno))

G_miss[geno %in% c("0/0", "0|0")] <- 0
G_miss[geno  %in% c("0/1", "1/0", "1|0", "0|1")] <- 1
G_miss[geno %in% c("1/1", "1|1")] <- 2

# removing missing data
G_miss[na.omit(G_miss)]
```

```{r}
# Converting the genotype matrix to class FBM.code256 using the matrix where missing data was removed
G1_miss <- add_code256(big_copy(t(G_miss),type="raw"),code=bigsnpr:::CODE_012)
```

```{r}
newpc_miss <- snp_autoSVD(G1_miss,infos.chr =chrom_sort,infos.pos = pos_sort)
which_pruned_miss <- attr(newpc_miss, which="subset") # Indexes of remaining SNPS after pruning
length(which_pruned_miss)
```

### Using `snp_fastImputeSimple` to impute missing genotypes

This requires the genotpe matrix to be converted to class FBM.code256.

> A reference class for storing and accessing up to 256 arbitrary different values using a Filebacked Big Matrix of type unsigned char.

```{r}
# Converting the genotype matrix to class FBM.code256
G1 <- add_code256(big_copy(t(G),type="raw"),code=bigsnpr:::CODE_012)
```

The chunk below imputs missing genotypes in order for the following code to run. Fast imputation via mode, mean, sampling according to allele frequencies, or 0.

Method argument: Either "random" (sampling according to allele frequencies), "mean0" (rounded mean), "mean2" (rounded mean to 2 decimal places), "mode" (most frequent call).

Depending on the method I choose, the # of SNPs kept after running the following chunk changes, so I'm not sure which method to use. 

```{r}
# Imputing missing genotypes
G_missSimple <- snp_fastImputeSimple(G1,method = "mode") # Using mode or no method call results in the same number of SNPs in the following code 
```

Truncated SVD while limiting LD.

```{r}
newpc<-snp_autoSVD(G_missSimple,infos.chr =chrom_sort,infos.pos = pos_sort)
which_pruned <- attr(newpc, which="subset") # Indexes of remaining SNPS after pruning
length(which_pruned)
```

Either removing missing data or imputation results in a similar number of trimmed SNPs, so I will continue forward with missing data removed.

```{r}
invisible(lapply(which_pruned_miss, write, "pruned_data.txt", append=TRUE))
```

**In terminal**

```
$ mawk '!/#/' neutralloci.recode.vcf | cut -f1,2 > totalloci
$ NUM=(`cat totalloci | wc -l`)
$ paste <(seq 1 $NUM) totalloci > loci.plus.index
$ cat pruned_data.txt | parallel "grep -w ^{} loci.plus.index" | cut -f2,3> pruned_data_neutral.loci.txt

$ head pruned_data_neutral.loci.txt

output:
NC_035780.1	573785
NC_035780.1	573788
NC_035780.1	574229
NC_035780.1	574307
NC_035780.1	574313
NC_035780.1	574321
NC_035780.1	574444
NC_035780.1	574580
NC_035780.1	574616
NC_035780.1	574657
```

Create VCF file with just the pruned_data loci 
```
$ vcftools --vcf neutralloci.recode.vcf --recode --recode-INFO-all --positions pruned_data_neutral.loci.txt --out pruned_data_neutral

output:
fter filtering, kept 40 out of 40 Individuals
Outputting VCF file...
After filtering, kept 14880 out of a possible 74520 Sites
Run Time = 3.00 seconds
```

```{r}
my_vcf_u <- read.vcfR("pruned_data_neutral.recode.vcf")
strata<- read.table("strata", header=TRUE)
```

```{r}
rad.u <- vcfR2genind(my_vcf_u, strata = strata, pop = c(rep("BIS", 10),rep("GB", 10),rep("NAR", 10), rep("PVD",10)))
```

```{r}
rad.u
```


```{r}
stratted.u <- strata(rad.u, formula= Population, combine = TRUE,just.strats)
stratted.u@other <- select(info, Latitude,Longitude, Distance,SE, Temperature,Salinity,pH,Chlorophylla,DO)
```

```{r}
stratted.u
```

**Make hierfstat object**
```{r}
hf.filt <- genind2hierfstat(rad.filt, pop = c(rep("BIS", 10),rep("GB", 10),rep("NAR", 10), rep("PVD",10))) 
```

```{r}
hf.u <- genind2hierfstat(rad.u, pop = c(rep("BIS", 10),rep("GB", 10),rep("NAR", 10), rep("PVD",10)))
hf.u <- hf.u$hierfstat.no.imputation
```


## Estimating effective migration surfaces [(EEMS)](https://github.com/dipetkov/eems)

The program `runeems_snps` implements the EEMS method for analyzing spatial population structure. This version uses the pairwise genetic dissimilarity matrix computed from SNP data.

Here is the code used to run `runeems_snps` in both RStudio and command-line. Detailed steps for running this program can be accessed [here](https://github.com/dipetkov/eems/tree/master/runeems_snps). Input files are created in RStudio or manually (see below) and saved to the directory where `runeems_snps` will be run. The program is then run in the command-line. The outputs are then plotted in RStudio. 

`runeems_snps` requires three data input files that have the same file name but different extension. The description below assumes that datapath is the full path + the file name (but without the extension).

__1. datapath.diffs__

`datapath.diffs` is the matrix of average pairwise genetic dissimilarities. This can be computed with bed2diffs from genetic data in plink binary format.

The dissimilarity matrix is nonnegative, symmetric, with 0s on the main diagonal. These conditions are necessary but not sufficient for diffs to be a valid dissimilarity matrix. Mathematically, diffs should be conditionally negative definite.

**Steps for generating `datapath.diffs` are completed in RStudio.**
```{r}
# V1 methtod to get diffs matrix, preferred
bed2diffs_v1 <- function(Geno) {
  nIndiv <- nrow(Geno)
  nSites <- ncol(Geno)
  Diffs <- matrix(0, nIndiv, nIndiv)
  
  for (i in seq(nIndiv - 1)) {
    for (j in seq(i + 1, nIndiv)) {
      x <- Geno[i, ]
      y <- Geno[j, ]
      Diffs[i, j] <- mean((x - y)^2, na.rm = TRUE)
      Diffs[j, i] <- Diffs[i, j]
    }
  }
  Diffs
}
```

```{r}
# V2 method to get .diffs matrix, only if V1 doesn't work
bed2diffs_v2 <- function(Geno) {
  nIndiv <- nrow(Geno)
  nSites <- ncol(Geno)
  Miss <- is.na(Geno)
  ## Impute NAs with the column means (= twice the allele frequencies)
  Mean <- matrix(colMeans(Geno, na.rm = TRUE), ## a row of means
                 nrow = nIndiv, ncol = nSites, byrow = TRUE) ## a matrix with nIndiv identical rows of means
  Mean[Miss == 0] <- 0 ## Set the means that correspond to observed genotypes to 0
  Geno[Miss == 1] <- 0 ## Set the missing genotypes to 0 (used to be NA) 
  Geno <- Geno + Mean
  ## Compute similarities
  Sim <- Geno %*% t(Geno) / nSites
  SelfSim <- diag(Sim) ## self-similarities
  vector1s <- rep(1, nIndiv) ## vector of 1s
  ## This chunk generates a `diffs` matrix
  Diffs <- SelfSim %*% t(vector1s) + vector1s %*% t(SelfSim) - 2 * Sim
  Diffs
}
```

```{r}
geno <- stratted.filt@tab
```

```{r}
# Get rid of non-biallelic loci
multi.loci <- names(which(stratted.filt@loc.n.all != 2))
multi.cols <- which(grepl(paste0("^", multi.loci, "\\.\\d+$", collapse = "|"), colnames(geno)))
if (length(multi.cols)) geno <- geno[, - multi.cols]
nloci <- dim(geno)[2] / 2
dim(geno)
```

```{r}
stopifnot(identical(stratted.filt@type, 'codom'))
```

```{r}
# bed2diffs functions  
diffs.v1 <- bed2diffs_v1(geno)
diffs.v2 <- bed2diffs_v2(geno)
# Round to 6 digits
diffs.v1 <- round(diffs.v1, digits = 6)
diffs.v2 <- round(diffs.v2, digits = 6)
```

**Check that the dissimilarity matrix has one positive eigenvalue and nIndiv-1 negative eigenvalues, as required by a full-rank Euclidean distance matrix. If the V1 method does not make a Euclidean matrix, you must use V2.**

```{r}
tail(sort(round(eigen(diffs.v1)$values, digits = 2)))
```

```{r}
tail(sort(round(eigen(diffs.v2)$values, digits = 2)))
```

```{r}
# Set suffix for EEMS input files
suf <- "neutraldata-filt"
```

```{r}
# This saves the file to directory 
write.table(diffs.v1, paste(suf,".v1.diffs",sep=""), 
            col.names = FALSE, row.names = FALSE, quote = FALSE)
```


__2. datapath.coord__

`datapath.coord` are the sample coordinates, two coordinates per sample, one sample per line. The sampling locations should be given in the same order as the rows and columns of the dissimilarity matrix.

**Steps for generating `datapath.coord` are completed in RStudio.**
```{r}
## Get gps coordinates from previously created info matrix
xOR.info <- dplyr::filter(info)
gps_matrix <- select(xOR.info,c("Longitude","Latitude"))
```

```{r}
#write .coord file
write.table(gps_matrix, paste(suf,".v1.coord",sep=""),col.names = FALSE, row.names = FALSE,quote = FALSE)
```

__3. datapath.outer__

`datapath.outer` are the habitat coordinates, as a sequence of vertices that form a closed polygon. The habitat vertices should be listed counterclockwise and the first vertex should also be the last vertex, so that the outline is a closed ring. Otherwise, EEMS attempts to "correct" the polygon and prints a warning message.

**`datapath.outer` is created manually in Excel,based on site coordinates gathered from Google Maps, copied into terminal using `nano`, and saved as the file with the appropriate extension.** 


**`runeems_snps` is then run in command-line following the steps documented in [NB_EEMS_OutlierHap.md](https://github.com/amyzyck/EecSeq_NB_EasternOyster/blob/master/Analysis/PopGen_SeaGen_Analyses/EEMS/NB_EEMS_OutlierHap.md)  

**Back in RStudio to plot `runeems_snps` outputs**

```{r}
# Install rEEMSplots
library(rEEMSplots)
```

```{r} 
# Plotting EEMS after running runeems_snps
path = "./NB_EEMS_Neutral/"
dirs = c(paste0(path,"neutraldata-D200-chain1"), paste0(path,"neutraldata-D300-chain1"), paste0(path,"neutraldata-D600-chain1"))
```

```{r}
eems.plots(mcmcpath = c(paste0(path,"./neutraldata-D200-chain1"), paste0(path,"neutraldata-D300-chain1"), paste0(path,"neutraldata-D600-chain1")), plotpath = paste0(path,"neutraldata-All-plots"),
           longlat = T,add.grid=F,add.outline = T,add.demes = T,
           projection.in = "+proj=longlat +datum=WGS84",projection.out = "+proj=merc +datum=WGS84",
           add.map = T,add.abline = T, add.r.squared = T)
```

The plots are saved to the same directory where `runeems_snps` was run. 

## Pairwise Fst

```{r}
fst.mat <- pairwise.WCfst(hf.filt)
```

```{r}
gindF.fst.mat.triN <- as.matrix(fst.mat)
colnames(gindF.fst.mat.triN) <- pop_order
rownames(gindF.fst.mat.triN) <- pop_order
```

```{r}
meltedN <- melt(gindF.fst.mat.triN, na.rm =TRUE)
round(gindF.fst.mat.triN,4)
```

```{r}
summary(meltedN$value)
```

```{r}
#Plotting Pairwise fst
neutral <- ggplot(data = meltedN, aes(Var2, Var1, fill = value))+ geom_tile(color = "white")+ 
  scale_fill_gradient(low = "white", high = "chocolate1", name="FST")  +
  ggtitle(expression(atop("Pairwise FST, WC (1984) Neutral SNPs", atop(italic("N = 40, L = 74,520"), ""))))+
  labs( x = "Sampling Site", y = "Sampling Site") + 
  theme(axis.text.x = element_text(angle = 45, vjust = 1, size = 12, hjust = 1),axis.text.y = element_text(size = 12)) + 
  theme(axis.title = element_text(size = 12),legend.text = element_text(size =14), legend.title = element_text(size =14)) +
  theme(plot.title = element_text(size = 14)) +
coord_fixed()
neutral
```

## Genetic diversity (observed and expected heterozygosity)

```{r}
comb <- summary(stratted.filt)
names(comb)
```

```{r}
plot(comb$Hobs, xlab="Loci number", ylab="Observed Heterozygosity", 
     main="Observed heterozygosity per locus")
```


```{r}
plot(comb$Hobs,comb$Hexp, xlab="Hobs", ylab="Hexp", 
     main="Expected heterozygosity as a function of observed heterozygosity per locus")
```


```{r}
bartlett.test(list(comb$Hexp, comb$Hobs)) # a test : H0: Hexp = Hobs
```

**_Significant difference between Observed and expected heterzygosity_**.  

```{r}
basicstat <- basic.stats(hf.filt, diploid = TRUE, digits = 3)
```

```{r}
as.data.frame(basicstat$overall)
```

```{r}
# get bootstrap confidence values for Fis
boot <- boot.ppfis(hf.filt,nboot = 1000)
boot5 <- boot.ppfis(hf.filt,nboot = 1000,quant = 0.5)
```

```{r}
# add latitude for each population
latitude = c(41.545, 41.654, 41.505, 41.816)
```

```{r}
# add longitude for each population
longitude = c(-71.431, -71.445, -71.453, -71.391)
```

```{r}
# add distance for each population
distance = c(4.76, 0.47, 15.41, 1.49)
```

```{r}
# add sewage effluent for each population
sewage = c(8.82, 14.60, 2.03, 59.86)
```

```{r}
# add temperature for each population
temperature = c(23, 24, 25, 23)
```

```{r}
# add salinity for each population
salinity = c(30, 28, 18, 25)
```

```{r}
# add pH for each population
pH = c(7.9, 7.4, 7.6, 7.4)
```

```{r}
# add Chlorophylla for each population
Chlor_a = c(4.9, 18.8, 4.6, 8.1)
```

```{r}
# add DO for each population
DO = c(8.2, 5.7, 7, 4.9)
```

```{r}
# combine all pop statistics
colnames(basicstat$Ho) <- pop_order
Ho <- colMeans(basicstat$Ho,na.rm = T)
He <- colMeans(basicstat$Hs,na.rm = T)
Fis<- boot5$fis.ci$ll
y <- cbind(pop_order,Ho,He,Fis,boot$fis.ci,latitude, longitude, distance,sewage, temperature, salinity, pH, Chlor_a, DO)
y
```

```{r}
summary(He)
```

```{r}
summary(Fis)
```

```{r}
# Plot He vs Latitude
R2 = round(summary(lm(y$He ~ y$latitude))$r.squared, 4)             
ggplot(y, aes(x = latitude, y = He)) + geom_point(aes(shape=pop_order, color=pop_order), size = 3) + 
  scale_shape_manual(values = c(15,18,17,16)) +
geom_smooth(method=lm) + 
ggtitle("Expected heterozygosity vs Latitude, Neutral SNPs") +
annotate(geom = "text", label=paste("italic(R^2)==",R2), x=41.55, y=0.2635, parse=T) +
scale_x_continuous()
```

```{r}
# Plot Ho vs Latitude
R2 = round(summary(lm(y$Ho ~ y$latitude))$r.squared, 4)             
ggplot(y, aes(x = latitude, y = Ho)) + geom_point(aes(shape=pop_order, color=pop_order), size = 3) + 
  scale_shape_manual(values = c(15,18,17,16)) + 
geom_smooth(method=lm) + 
ggtitle("Observed heterozygosity vs Latitude, Neutral SNPs") +
annotate(geom = "text", label=paste("italic(R^2)==",R2), x=41.55, y=0.2635, parse=T) +
scale_x_continuous()
```

```{r}
#Plot He vs Longitude
R3 = round(summary(lm(y$He ~ y$longitude))$r.squared, 4)             
ggplot(y, aes(x = longitude, y = He)) + geom_point(aes(shape=pop_order, color=pop_order), size = 3) + 
  scale_shape_manual(values = c(15,18,17,16)) +
geom_smooth(method=lm) + 
ggtitle("Expected heterozygosity vs Longitude, Neutral SNPs") +
annotate(geom = "text", label=paste("italic(R^2)==",R3), x=-71.45, y=0.2635, parse=T) +
scale_x_reverse()
```

```{r}
#Plot Ho vs Longitude
R3 = round(summary(lm(y$Ho ~ y$longitude))$r.squared, 4)             
ggplot(y, aes(x = longitude, y = Ho)) + geom_point(aes(shape=pop_order, color=pop_order), size = 3) + 
  scale_shape_manual(values = c(15,18,17,16)) +
geom_smooth(method=lm) + 
ggtitle("Observed heterozygosity vs Longitude, Neutral SNPs") +
annotate(geom = "text", label=paste("italic(R^2)==",R3), x=-71.45, y=0.2635, parse=T) +
scale_x_reverse()
```

```{r}
#Plot He vs Distance from sewage outflow 
R4 = round(summary(lm(y$He ~ y$distance))$r.squared, 4)             
ggplot(y, aes(x = distance, y = He)) + geom_point(aes(shape=pop_order, color=pop_order), size = 3) + 
  scale_shape_manual(values = c(15,18,17,16)) + 
geom_smooth(method=lm) + 
ggtitle("Expected heterozygosity vs Distance, Neutral SNPs") +
annotate(geom = "text", label=paste("italic(R^2)==",R4), x=2, y=0.2685, parse=T) +
scale_x_reverse()
```

```{r}
#Plot Ho vs Distance from sewage outflow 
R4 = round(summary(lm(y$Ho ~ y$distance))$r.squared, 4)             
ggplot(y, aes(x = distance, y = Ho)) + geom_point(aes(shape=pop_order, color=pop_order), size = 3) + 
  scale_shape_manual(values = c(15,18,17,16)) + 
geom_smooth(method=lm) + 
ggtitle("Observed heterozygosity vs Distance, Neutral SNPs") +
annotate(geom = "text", label=paste("italic(R^2)==",R4), x=2, y=0.2635, parse=T) +
scale_x_reverse()
```

```{r}
#Plot He vs Sewage Effluent 
R4 = round(summary(lm(y$He ~ y$sewage))$r.squared, 4)             
ggplot(y, aes(x = sewage, y = He)) + geom_point(aes(shape=pop_order, color=pop_order), size = 3) + 
  scale_shape_manual(values = c(15,18,17,16)) + 
geom_smooth(method=lm) + 
ggtitle("Expected heterozygosity vs Sewage Effluent, Neutral SNPs") +
annotate(geom = "text", label=paste("italic(R^2)==",R4), x=20, y=0.265, parse=T) +
scale_x_reverse()
```

```{r}
#Plot Ho vs Sewage Effluent 
R4 = round(summary(lm(y$Ho ~ y$sewage))$r.squared, 4)             
ggplot(y, aes(x = sewage, y = Ho)) + geom_point(aes(shape=pop_order, color=pop_order), size = 3) + 
  scale_shape_manual(values = c(15,18,17,16)) + 
geom_smooth(method=lm) + 
ggtitle("Observed heterozygosity vs Sewage Effluent, Neutral SNPs") +
annotate(geom = "text", label=paste("italic(R^2)==",R4), x=20, y=0.2635, parse=T) +
scale_x_reverse()
```

```{r}
#Plot He vs Temperature 
R4 = round(summary(lm(y$He ~ y$temperature))$r.squared, 4)             
ggplot(y, aes(x = temperature, y = He)) + geom_point(aes(shape=pop_order, color=pop_order), size = 3) + 
  scale_shape_manual(values = c(15,18,17,16)) +
geom_smooth(method=lm) + 
ggtitle("Expected heterozygosity vs Temperature, Neutral SNPs") +
annotate(geom = "text", label=paste("italic(R^2)==",R4), x=23.25, y=0.2635, parse=T) +
scale_x_reverse()
```

```{r}
#Plot Ho vs Temperature 
R4 = round(summary(lm(y$Ho ~ y$temperature))$r.squared, 4)             
ggplot(y, aes(x = temperature, y = Ho)) + geom_point(aes(shape=pop_order, color=pop_order), size = 3) + 
  scale_shape_manual(values = c(15,18,17,16)) + 
geom_smooth(method=lm) + 
ggtitle("Observed heterozygosity vs Temperature, Neutral SNPs") +
annotate(geom = "text", label=paste("italic(R^2)==",R4), x=23.25, y=0.2635, parse=T) +
scale_x_reverse()
```

```{r}
#Plot He vs Salinity 
R4 = round(summary(lm(y$He ~ y$salinity))$r.squared, 4)             
ggplot(y, aes(x = salinity, y = He)) + geom_point(aes(shape=pop_order, color=pop_order), size = 3) + 
  scale_shape_manual(values = c(15,18,17,16)) + 
geom_smooth(method=lm) + 
ggtitle("Expected heterozygosity vs Salinity, Neutral SNPs") +
annotate(geom = "text", label=paste("italic(R^2)==",R4), x=23.25, y=0.2675, parse=T) +
scale_x_reverse()
```

```{r}
#Plot Ho vs Salinity 
R4 = round(summary(lm(y$Ho ~ y$salinity))$r.squared, 4)             
ggplot(y, aes(x = salinity, y = Ho)) + geom_point(aes(shape=pop_order, color=pop_order), size = 3) + 
  scale_shape_manual(values = c(15,18,17,16)) + 
geom_smooth(method=lm) + 
ggtitle("Observed heterozygosity vs Salinity, Neutral SNPs") +
annotate(geom = "text", label=paste("italic(R^2)==",R4), x=23.25, y=0.2635, parse=T) +
scale_x_reverse()
```

```{r}
#Plot He vs pH
R4 = round(summary(lm(y$He ~ y$pH))$r.squared, 4)             
ggplot(y, aes(x = pH, y = He)) + geom_point(aes(shape=pop_order, color=pop_order), size = 3) + 
  scale_shape_manual(values = c(15,18,17,16)) +
geom_smooth(method=lm) + 
ggtitle("Expected heterozygosity vs pH, Neutral SNPs") +
annotate(geom = "text", label=paste("italic(R^2)==",R4), x=7.45, y=0.2635, parse=T) +
scale_x_reverse()
```

```{r}
#Plot Ho vs pH
R4 = round(summary(lm(y$Ho ~ y$pH))$r.squared, 4)             
ggplot(y, aes(x = pH, y = Ho)) + geom_point(aes(shape=pop_order, color=pop_order), size = 3) + 
  scale_shape_manual(values = c(15,18,17,16)) + 
geom_smooth(method=lm) + 
ggtitle("Observed heterozygosity vs pH, Neutral SNPs") +
annotate(geom = "text", label=paste("italic(R^2)==",R4), x=7.45, y=0.2635, parse=T) +
scale_x_reverse()
```

```{r}
#Plot He vs Chlorophyll a 
R4 = round(summary(lm(y$He ~ y$Chlor_a))$r.squared, 4)             
ggplot(y, aes(x = Chlor_a, y = He)) + geom_point(aes(shape=pop_order, color=pop_order), size = 3) + 
  scale_shape_manual(values = c(15,18,17,16)) + 
geom_smooth(method=lm) + 
ggtitle("Expected heterozygosity vs Chlorophyll a, Neutral SNPs") +
annotate(geom = "text", label=paste("italic(R^2)==",R4), x=6, y=0.2665, parse=T) +
scale_x_reverse()
```

```{r}
#Plot Ho vs Chlorophyll a 
R4 = round(summary(lm(y$Ho ~ y$Chlor_a))$r.squared, 4)             
ggplot(y, aes(x = Chlor_a, y = Ho)) + geom_point(aes(shape=pop_order, color=pop_order), size = 3) + 
  scale_shape_manual(values = c(15,18,17,16)) + 
geom_smooth(method=lm) + 
ggtitle("Observed heterozygosity vs Chlorophyll a, Neutral SNPs") +
annotate(geom = "text", label=paste("italic(R^2)==",R4), x=6, y=0.2665, parse=T) +
scale_x_reverse()
```

```{r}
#Plot He vs Dissolved Oxygen 
R4 = round(summary(lm(y$He ~ y$DO))$r.squared, 4)             
ggplot(y, aes(x = DO, y = He)) + geom_point(aes(shape=pop_order, color=pop_order), size = 3) + 
  scale_shape_manual(values = c(15,18,17,16)) + 
geom_smooth(method=lm) + 
ggtitle("Expected heterozygosity vs Dissolved Oxygen, Neutral SNPs") +
annotate(geom = "text", label=paste("italic(R^2)==",R4), x=5.5, y=0.2635, parse=T) +
scale_x_reverse()
```

```{r}
#Plot Ho vs Dissolved Oxygen 
R4 = round(summary(lm(y$Ho ~ y$DO))$r.squared, 4)             
ggplot(y, aes(x = DO, y = Ho)) + geom_point(aes(shape=pop_order, color=pop_order), size = 3) + 
  scale_shape_manual(values = c(15,18,17,16)) + 
geom_smooth(method=lm) + 
ggtitle("Observed heterozygosity vs Dissolved Oxygen, Neutral SNPs") +
annotate(geom = "text", label=paste("italic(R^2)==",R4), x=5.5, y=0.2635, parse=T) +
scale_x_reverse()
```


## PCA 

```{r}
X <- tab(stratted.u, freq = TRUE, NA.method = "mean")
pca1 <- dudi.pca(X, scale = FALSE, scannf = FALSE, nf = 3)
barplot(pca1$eig[1:50], main = "PCA eigenvalues", col = heat.colors(50))
s.class(pca1$li, pop(stratted.filt))
title("PCA of neutral SNPs\naxes 1-2")
add.scatter.eig(pca1$eig[1:20], 3,1,2)
col <- funky(15) 
s.class(pca1$li, pop(stratted.filt),xax=1,yax=2, col=col, axesell=FALSE, cstar=0, cpoint=3, grid=FALSE)
```


## Discriminant Analysis of Principal Components (DAPC)

The first line of the code block below is interactive. In the console, it will tell you:  
1. Choose the number PCs to retain (>=1): 
    I picked 1 
2. Choose the number discriminant functions to retain (>=1):
    I picked 2

```{r}
grp <- find.clusters(stratted.u, n.pca = 1, choose.n.clust = FALSE)
table(pop(stratted.u), grp$grp)
table.value(table(pop(stratted.u), grp$grp), col.lab=paste("inf", 1:2), row.lab=paste("ori", 1:4))
```

Again, the first line of the code block below is interactive. In the console, it will tell you:  
1. Choose the number PCs to retain (>=1): 
    I picked 1 
2. Choose the number discriminant functions to retain (>=1):
    I picked 1

```{r}
dapc1 <- dapc(stratted.u, grp$grp, n.pca = 1, n.da= 1)
scatter(dapc1,col=col,bg="white", solid=1)
```

Again, the first line of the code block below is interactive. In the console, it will tell you:  
1. Choose the number PCs to retain (>=1): 
    I picked 1 
2. Choose the number discriminant functions to retain (>=1):
    I picked 1

```{r}
contrib <- loadingplot(dapc1$var.contr, axis=1, thres=.01, lab.jitter=1)
contrib
setPop(rad.filt) <- ~Library
dapc1 <- dapc(stratted.u, pop(stratted.u), n.pca = 1, n.da = 1)
contrib <- loadingplot(dapc1$var.contr, axis=1, thres=.05, lab.jitter=1)
```

```{r}
#Structure Like
compoplot(dapc1, posi="bottomright",txt.leg=paste("Cluster", 1:4), lab="", ncol=1, xlab="individuals")
temp <- which(apply(dapc1$posterior,1, function(e) all(e<0.9)))
compoplot(dapc1, subset=temp, posi="bottomright", txt.leg=paste("Cluster", 1:4), ncol=2)
```


### PCAviz

```{r}
NA.afDraw<- function(ind){
  ind.mat <- ind@tab
  new.mat <- ind.mat
  af = colSums(ind.mat[,seq(1,ncol(ind.mat)-1,2)],na.rm = TRUE)/
      (2*apply(ind.mat[,seq(1,ncol(ind.mat)-1,2)],2,function(x) sum(!is.na(x))))
  af.Draw <- function(geno, af){
     new <- function(geno,af){
        if(is.na(geno)){
        newA = rbinom(1,2,af)
        }
        else {newA <- geno}
        return(newA)
   }
  new.row <- mapply(geno,af,FUN = new)
  return(new.row)}
  
  new.mat[,seq(1,ncol(ind.mat)-1,2)] <- t(apply(ind.mat[,seq(1,ncol(ind.mat)-1,2)],1,af.Draw,af))
  new.mat[,seq(2,ncol(ind.mat),2)] <- 2-new.mat[,seq(1,ncol(ind.mat)-1,2)]
  new.ind <- ind
  new.ind@tab <- new.mat
  return(new.ind)
}
```

```{r}
u.na <- NA.afDraw(stratted.u)
```

```{r}
pca <- dudi.pca(u.na,center=TRUE,scale=TRUE,scannf = F, nf = 30)
```

```{r}
col18 <- funky(length(unique(u.na@strata$Population)))
#Colors that match the neutral Structure results
col6 <-  c("firebrick","dodgerblue","midnightblue","salmon")
par(mfrow=c(2,2))
s.class(pca$li, strata(u.na)$Population,xax=1,yax=2, 
        sub = "neutralloci,PC 1-2, 40 individuals, 14,880 loci, maf 5.0%",
        possub = "topleft",col=transp(col6,.6),axesell=FALSE,
        cstar=0, cpoint=3, grid=FALSE, cellipse = 0)
s.class(pca$li, strata(u.na)$Population,xax=1,yax=3, 
        sub = "neutralloci,PC 1-3, 40 individuals, 14,880 loci, maf 5.0%",
        possub = "topleft",col=transp(col6,.6),axesell=FALSE,
        cstar=0, cpoint=3, grid=FALSE, cellipse = 0)
s.class(pca$li, strata(u.na)$Population,xax=1,yax=4, 
        sub = "neutralloci,PC 1-4, 40 individuals, 14,880 loci, maf 5.0%",
        possub = "topleft",col=transp(col6,.6),axesell=FALSE,
        cstar=0, cpoint=3, grid=FALSE, cellipse = 0)
s.class(pca$li, strata(u.na)$Population,xax=1,yax=5, 
        sub = "neutralloci,PC 1-5, 40 individuals, 14,880 loci, maf 5.0%",
        possub = "topleft",col=transp(col6,.6),axesell=FALSE,
        cstar=0, cpoint=3, grid=FALSE, cellipse = 0)
```

```{r}
s.label(pca$li, xax=1,yax=5, 
        sub = "neutralloci,PC 1-2, 40 individuals, 14,880 loci, maf 5.0%",
        possub = "topleft")
```


```{r}
eig.perc <- 100*pca$eig/sum(pca$eig)
head(eig.perc)
```

```{r}
li <-pca$li
c1 <- pca$c1
#Create dataframe of info like latitude and population for each individual
info_mat <- as.data.frame(cbind(u.na@strata, u.na@other$Latitude, u.na@other$Longitude, u.na@other$Distance, u.na@other$Temperature, u.na@other$Salinity, u.na@other$pH, u.na@other$Chlorophylla, u.na@other$DO))
colnames(info_mat) <- c("Population","Latitude", "Longitude", "Distance","Temperature","Salinity","pH","Chlorophylla","DO")
colnames(c1) <- colnames(li)
#create pcaviz object
pviz <- pcaviz(x=li,rotation=c1,dat=info_mat)
```

```{r}
p = list(size=2) 
plot(pviz,color = "Distance", draw.points = T, group.summary.labels = F, draw.pc.axes = T, geom.point.params = p)
```

```{r}
# Plot against distance 
p = list(size=6)
plot(pviz, coords = c("Axis1","Distance"),
     show.legend = T,color = "Population",colors = col18,
     draw.points =T,group.summary.labels = F,geom.point.params= list(size = 6, alpha = 0.9),geom.point.summary.params = list(size =6,alpha = 0.4))
```

```{r}
plot(pviz,coords = c("Axis1","Axis2","Axis3","Axis4","Axis5","Distance"),group = NULL,color= "Population",colors = col18,draw.points = T,scale.pc.axes = 0.6, show.legend=F)
```

```{r}
p = list(size=6)
A2 <- plot(pviz,coords = c("Axis2","Distance"),group="Population",
     show.legend = F,color = "Population",colors = col6,draw.points =T,
     group.summary.labels = F,draw.linear.fit = F, coord_fixed=T)
A3 <- plot(pviz,coords = c("Axis3","Distance"),group="Population",
     show.legend = F,color = "Population",colors = col6,
     draw.points =T,group.summary.labels = F,draw.linear.fit = F)
A4 <- plot(pviz,coords = c("Axis4","Distance"),group="Population",
     show.legend = F,color = "Population",colors = col6,
     draw.points =T,group.summary.labels = F,draw.linear.fit = F)
A5 <- plot(pviz,coords = c("Axis5","Distance"),group="Population",
     show.legend = T,color = "Population",colors = col6,
     draw.points =T,group.summary.labels = F,draw.linear.fit = F)
plot_grid(A2,A3,A4,A5, labels = c('A','B','C','D'))
```



# Neutral Genic SNPs

****

## Making files

**Make genind object**

`neutralgene.recode.vcf` contains neutral SNPs from genic regions for populations PVD, GB, BIS, and NAR. Steps for generating this VCF file are located in [EecSeq_Cvirginica_NeutralLoci.md](https://github.com/amyzyck/EecSeq_NB_EasternOyster/blob/master/Analysis/PopGen_SeaGen_Analyses/NonHaplotigMasked_Genome/Neutral_SNPs/EecSeq_Cvirginica_NeutralLoci.md).

Population NIN was removed from the VCF file following steps in [EecSeq_Cvirginica_OutlierDetection.md](https://github.com/amyzyck/EecSeq_NB_EasternOyster/blob/master/Analysis/EecSeq_Cvirginica_OutlierDetection.md).

`strata` contains population, environmental, and library information for each sample - can be accessed [here](https://github.com/amyzyck/EecSeq_NB_EasternOyster/blob/master/Analysis/PopGen_SeaGen_Analyses/strata).

```{r}
my_vcf_gene <- read.vcfR("neutralgene.recode.vcf")
strata <- read.table("strata", header=TRUE)
```

```{r}
rad.filt_gene <- vcfR2genind(my_vcf_gene, strata = strata, pop = c(rep("BIS", 10),rep("GB", 10),rep("NAR", 10), rep("PVD",10)))
```

```{r}
rad.filt_gene
```

```{r}
#Providing population names for plotting
pop_order <- c("BIS","GB","NAR","PVD")
```

**Read in the other info from .strata file and extract information such as locality, latitude, and longitude.**
```{r}
info <- as.data.frame(read.table("strata",header = T,sep = "\t",stringsAsFactors = F))
mystrats_gene <- as.data.frame(matrix(nrow = length(indNames(rad.filt_gene)),ncol=10))
colnames(mystrats_gene) <- c("Population","Latitude","Longitude","Distance","Temperature","SE","Salinity","pH","Chlorophylla","DO")
just.strats <- select(info,c("Population"))
stratted.filt_gene <- strata(rad.filt_gene, formula= Population, combine = TRUE,just.strats)
stratted.filt_gene@other <- select(info, Latitude,Longitude,Distance,SE,Temperature,Salinity,pH,Chlorophylla,DO)
```

```{r}
stratted.filt_gene
```

**Repeat for quasi-independent set of SNPs**

Steps for filtering are documented for OutFLANK in [EecSeq_Cvirginica_OutlierDetection.md](https://github.com/amyzyck/EecSeq_NB_EasternOyster/blob/master/Analysis/EecSeq_Cvirginica_OutlierDetection.md). However, I am repeating the steps here. 

```{r}
my_vcf_gene <- read.vcfR("neutralgene.recode.vcf")

geno_gene <- extract.gt(my_vcf_gene) # Character matrix containing the genotypes
position_gene <- getPOS(my_vcf_gene) # Positions in bp
chromosome_gene <- getCHROM(my_vcf_gene) # Chromosome information

G_gene <- matrix(NA, nrow = nrow(geno_gene), ncol = ncol(geno_gene))

G_gene[geno_gene %in% c("0/0", "0|0")] <- 0
G_gene[geno_gene  %in% c("0/1", "1/0", "1|0", "0|1")] <- 1
G_gene[geno_gene %in% c("1/1", "1|1")] <- 2

# NA should be replaced with “9” 
G_gene[is.na(G_gene)] <- 9
```


Chromosomes need to be of class integer for this to work.
```{r}
# Visualizing chromosomes
chrom_unique_gene <- unique(chromosome_gene)
print(chrom_unique_gene)
```

```{r}
# Removing "NC_" from the chromosome name so it can be converted to an integer
chrom_new_gene <- chromosome_gene%>%str_replace("NC_","")

# Converting the character string into an integer
chrom1_gene <- as.integer(chrom_new_gene)
```

`chromosome` and `position` need to be sorted for imputation to work.

```{r}
chrom_sort_gene <- sort(chrom1_gene)
pos_sort_gene <- sort(position_gene)
```

_Note: This filtering program does not allow for missing genotype values._

### Remove missing genotypes

```{r}
# removing missing data 
G_gene_miss <- matrix(NA, nrow = nrow(geno_gene), ncol = ncol(geno_gene))

G_gene_miss[geno_gene %in% c("0/0", "0|0")] <- 0
G_gene_miss[geno_gene  %in% c("0/1", "1/0", "1|0", "0|1")] <- 1
G_gene_miss[geno_gene %in% c("1/1", "1|1")] <- 2

# removing missing data
G_gene_miss[na.omit(G_gene_miss)]
```


```{r}
# Converting the genotype matrix to class FBM.code256 using the matrix where missing data was removed
G1_gene_miss <- add_code256(big_copy(t(G_gene_miss),type="raw"),code=bigsnpr:::CODE_012)
```

```{r}
newpc_gene_miss <- snp_autoSVD(G1_gene_miss,infos.chr =chrom_sort_gene,infos.pos = pos_sort_gene)
which_pruned_gene_miss <- attr(newpc_gene_miss, which="subset") # Indexes of remaining SNPS after pruning
length(which_pruned_gene_miss)
```


```{r}
invisible(lapply(which_pruned_gene_miss, write, "pruned_data_gene.txt", append=TRUE))
```


**In terminal**

```
$ mawk '!/#/' neutralexon.recode.vcf | cut -f1,2 > totallociexon
$ NUM=(`cat totallociexon | wc -l`)
$ paste <(seq 1 $NUM) totallociexon > lociexon.plus.index
$ cat pruned_data_exon.txt | parallel "grep -w ^{} lociexon.plus.index" | cut -f2,3> pruned_data_exon.loci.txt

$ head pruned_data_exon.loci.txt

output:
NC_035780.1	573785
NC_035780.1	573788
NC_035780.1	574229
NC_035780.1	574241
NC_035780.1	574307
NC_035780.1	574313
NC_035780.1	574568
NC_035780.1	574580
NC_035780.1	574616
NC_035780.1	620575
```

Create VCF file with just the pruned_data loci 
```
$ vcftools --vcf neutralexon.recode.vcf --recode --recode-INFO-all --positions pruned_data_exon.loci.txt --out pruned_data_exon

output:
After filtering, kept 40 out of 40 Individuals
Outputting VCF file...
After filtering, kept 11377 out of a possible 50238 Sites
Run Time = 2.00 seconds
```

```{r}
my_vcf_u_gene <- read.vcfR("pruned_data_gene.recode.vcf")
strata <- read.table("strata", header=TRUE)
```

```{r}
rad.u_gene <- vcfR2genind(my_vcf_u_gene, strata = strata, pop = c(rep("BIS", 10),rep("GB", 10),rep("NAR", 10), rep("PVD",10)))
```

```{r}
rad.u_gene
```


```{r}
stratted.u_gene <- strata(rad.u_gene, formula= Population, combine = TRUE,just.strats)
stratted.u_gene@other <- select(info, Latitude,Longitude, Distance, SE,  Temperature,Salinity,pH,Chlorophylla,DO)
```

```{r}
stratted.u_gene
```

**Make hierfstat object**
```{r}
hf.filt_gene <- genind2hierfstat(rad.filt_gene, pop = c(rep("BIS", 10),rep("GB", 10),rep("NAR", 10), rep("PVD",10))) 
```

```{r}
hf.u_gene <- genind2hierfstat(rad.u_gene, pop = c(rep("BIS", 10),rep("GB", 10),rep("NAR", 10), rep("PVD",10)))
hf.u_gene <- hf.u_gene$hierfstat.no.imputation
```

## Estimating effective migration surfaces [(EEMS)](https://github.com/dipetkov/eems)

__1. datapath.diffs__

**Steps for generating `datapath.diffs` are completed in RStudio.**
```{r}
# V1 methtod to get diffs matrix, preferred
bed2diffs_v1 <- function(Geno) {
  nIndiv <- nrow(Geno)
  nSites <- ncol(Geno)
  Diffs <- matrix(0, nIndiv, nIndiv)
  
  for (i in seq(nIndiv - 1)) {
    for (j in seq(i + 1, nIndiv)) {
      x <- Geno[i, ]
      y <- Geno[j, ]
      Diffs[i, j] <- mean((x - y)^2, na.rm = TRUE)
      Diffs[j, i] <- Diffs[i, j]
    }
  }
  Diffs
}
```

```{r}
# V2 method to get .diffs matrix, only if V1 doesn't work
bed2diffs_v2 <- function(Geno) {
  nIndiv <- nrow(Geno)
  nSites <- ncol(Geno)
  Miss <- is.na(Geno)
  ## Impute NAs with the column means (= twice the allele frequencies)
  Mean <- matrix(colMeans(Geno, na.rm = TRUE), ## a row of means
                 nrow = nIndiv, ncol = nSites, byrow = TRUE) ## a matrix with nIndiv identical rows of means
  Mean[Miss == 0] <- 0 ## Set the means that correspond to observed genotypes to 0
  Geno[Miss == 1] <- 0 ## Set the missing genotypes to 0 (used to be NA) 
  Geno <- Geno + Mean
  ## Compute similarities
  Sim <- Geno %*% t(Geno) / nSites
  SelfSim <- diag(Sim) ## self-similarities
  vector1s <- rep(1, nIndiv) ## vector of 1s
  ## This chunk generates a `diffs` matrix
  Diffs <- SelfSim %*% t(vector1s) + vector1s %*% t(SelfSim) - 2 * Sim
  Diffs
}
```

```{r}
geno_gene <- stratted.filt_gene@tab
```

```{r}
# Get rid of non-biallelic loci
multi.loci_gene <- names(which(stratted.filt_gene@loc.n.all != 2))
multi.cols_gene <- which(grepl(paste0("^", multi.loci_gene, "\\.\\d+$", collapse = "|"), colnames(geno_gene)))
if (length(multi.cols_gene)) geno_gene <- geno_gene[, - multi.cols_gene]
nloci_gene <- dim(geno_gene)[2] / 2
dim(geno_gene)
```

```{r}
stopifnot(identical(stratted.filt_gene@type, 'codom'))
```

```{r}
# bed2diffs functions  
diffs.v1_gene <- bed2diffs_v1(geno_gene)
diffs.v2_gene <- bed2diffs_v2(geno_gene)
# Round to 6 digits
diffs.v1_gene <- round(diffs.v1_gene, digits = 6)
diffs.v2_gene <- round(diffs.v2_gene, digits = 6)
```

**Check that the dissimilarity matrix has one positive eigenvalue and nIndiv-1 negative eigenvalues, as required by a full-rank Euclidean distance matrix. If the V1 method does not make a Euclidean matrix, you must use V2.**

```{r}
tail(sort(round(eigen(diffs.v1_gene)$values, digits = 2)))
```

```{r}
tail(sort(round(eigen(diffs.v2_gene)$values, digits = 2)))
```

```{r}
# Set suffix for EEMS input files
suf_gene <- "neutralexondata-filt"
```

```{r}
# This saves the file to directory 
write.table(diffs.v1_gene, paste(suf_gene,".v1.diffs",sep=""), 
            col.names = FALSE, row.names = FALSE, quote = FALSE)
```


__2. datapath.coord__

`datapath.coord` are the sample coordinates, two coordinates per sample, one sample per line. The sampling locations should be given in the same order as the rows and columns of the dissimilarity matrix.

**Steps for generating `datapath.coord` are completed in RStudio.**
```{r}
## Get gps coordinates from previously created info matrix
xOR.info <- dplyr::filter(info)
gps_matrix <- select(xOR.info,c("Longitude","Latitude"))
```

```{r}
#write .coord file
write.table(gps_matrix, paste(suf_gene,".v1.coord",sep=""),col.names = FALSE, row.names = FALSE,quote = FALSE)
```

__3. datapath.outer__

`datapath.outer` are the habitat coordinates, as a sequence of vertices that form a closed polygon. The habitat vertices should be listed counterclockwise and the first vertex should also be the last vertex, so that the outline is a closed ring. Otherwise, EEMS attempts to "correct" the polygon and prints a warning message.

**`datapath.outer` is created manually in Excel,based on site coordinates gathered from Google Maps, copied into terminal using `nano`, and saved as the file with the appropriate extension.** 


**`runeems_snps` is then run in command-line following the steps documented in [NB_EEMS_OutlierHap.md](https://github.com/amyzyck/EecSeq_NB_EasternOyster/blob/master/Analysis/PopGen_SeaGen_Analyses/EEMS/NB_EEMS_OutlierHap.md)  

**Back in RStudio to plot `runeems_snps` outputs**

```{r}
# Install rEEMSplots
library(rEEMSplots)
```

```{r} 
# Plotting EEMS after running runeems_snps
path = "./NB_EEMS_Neutral/"
dirs = c(paste0(path,"neutralgenedata-D200-chain1"), paste0(path,"neutralgenedata-D300-chain1"), paste0(path,"neutralgenedata-D600-chain1"))
```

```{r}
eems.plots(mcmcpath = c(paste0(path,"./neutralgenedata-D200-chain1"), paste0(path,"neutralgenedata-D300-chain1"), paste0(path,"neutralgenedata-D600-chain1")), plotpath = paste0(path,"neutralgenedata-All-plots"),
           longlat = T,add.grid=F,add.outline = T,add.demes = T,
           projection.in = "+proj=longlat +datum=WGS84",projection.out = "+proj=merc +datum=WGS84",
           add.map = T,add.abline = T, add.r.squared = T)
```

## Pairwise Fst

```{r}
fst.mat_gene <- pairwise.WCfst(hf.filt_gene)
```

```{r}
gindF.fst.mat.triN_gene <- as.matrix(fst.mat_gene)
colnames(gindF.fst.mat.triN_gene) <- pop_order
rownames(gindF.fst.mat.triN_gene) <- pop_order
```

```{r}
meltedN_gene <- melt(gindF.fst.mat.triN_gene, na.rm =TRUE)
round(gindF.fst.mat.triN_gene,4)
```

```{r}
summary(meltedN_gene$value)
```


```{r}
#Plotting Pairwise fst
neutral <- ggplot(data = meltedN_gene, aes(Var2, Var1, fill = value))+ geom_tile(color = "white")+ 
  scale_fill_gradient(low = "white", high = "red", name="FST")  +
  ggtitle(expression(atop("Pairwise FST, WC (1984) Neutral Gene SNPs", atop(italic("N = 40, L = 66,939"), ""))))+
  labs( x = "Sampling Site", y = "Sampling Site") + 
  theme(axis.text.x = element_text(angle = 45, vjust = 1, size = 12, hjust = 1),axis.text.y = element_text(size = 13)) + 
  theme(axis.title = element_text(size = 16),legend.text = element_text(size =15), legend.title = element_text(size =16)) +
  theme(plot.title = element_text(size = 17)) +
coord_fixed()
neutral
```

## Genetic diversity (observed and expected heterozygosity)

```{r}
comb_gene <- summary(stratted.filt_gene)
names(comb_gene)
```

```{r}
plot(comb_gene$Hobs, xlab="Loci number", ylab="Observed Heterozygosity", 
     main="Observed heterozygosity per locus")
```


```{r}
plot(comb_gene$Hobs,comb_gene$Hexp, xlab="Hobs", ylab="Hexp", 
     main="Expected heterozygosity as a function of observed heterozygosity per locus")
```

```{r}
bartlett.test(list(comb_gene$Hexp, comb_gene$Hobs)) # a test : H0: Hexp = Hobs
```

**_Significant difference between Observed and expected heterzygosity_**.  

```{r}
basicstat_gene <- basic.stats(hf.filt_gene, diploid = TRUE, digits = 3)
```

```{r}
as.data.frame(basicstat_gene$overall)
```

```{r}
# get bootstrap confidence values for Fis
boot_gene <- boot.ppfis(hf.filt_gene,nboot = 1000)
boot5_gene <- boot.ppfis(hf.filt_gene,nboot = 1000,quant = 0.5)
```

```{r}
# combine all pop statistics
colnames(basicstat_gene$Ho) <- pop_order
Ho_gene <- colMeans(basicstat_gene$Ho,na.rm = T)
He_gene <- colMeans(basicstat_gene$Hs,na.rm = T)
Fis_gene <- boot5_gene$fis.ci$ll
y_gene <- cbind(pop_order,Ho_gene,He_gene,Fis_gene,boot_gene$fis.ci,latitude, longitude, distance, sewage, temperature, salinity, pH, Chlor_a, DO)
y_gene
```

```{r}
summary(He_gene)
```

```{r}
summary(Fis_gene)
```

```{r}
# Plot He vs Latitude
R2_gene = round(summary(lm(y_gene$He_gene ~ y_gene$latitude))$r.squared, 4)             
ggplot(y_gene, aes(x = latitude, y = He_gene)) + geom_point(aes(shape=pop_order, color=pop_order), size = 3) + 
  scale_shape_manual(values = c(15,18,17,16)) +
geom_smooth(method=lm) + 
ggtitle("Expected heterozygosity vs Latitude, Neutral Gene SNPs") +
annotate(geom = "text", label=paste("italic(R^2)==",R2_gene), x=41.55, y=0.2635, parse=T) +
scale_x_continuous()
```

```{r}
# Plot Ho vs Latitude
R2_gene = round(summary(lm(y_gene$Ho_gene ~ y_gene$latitude))$r.squared, 4)             
ggplot(y_gene, aes(x = latitude, y = Ho_gene)) + geom_point(aes(shape=pop_order, color=pop_order), size = 3) + 
  scale_shape_manual(values = c(15,18,17,16)) + 
geom_smooth(method=lm) + 
ggtitle("Observed heterozygosity vs Latitude, Neutral Gene SNPs") +
annotate(geom = "text", label=paste("italic(R^2)==",R2_gene), x=41.55, y=0.2635, parse=T) +
scale_x_continuous()
```

```{r}
#Plot He vs Longitude
R3_gene = round(summary(lm(y_gene$He_gene ~ y_gene$longitude))$r.squared, 4)             
ggplot(y_gene, aes(x = longitude, y = He_gene)) + geom_point(aes(shape=pop_order, color=pop_order), size = 3) + 
  scale_shape_manual(values = c(15,18,17,16)) +
geom_smooth(method=lm) + 
ggtitle("Expected heterozygosity vs Longitude, Neutral Gene SNPs") +
annotate(geom = "text", label=paste("italic(R^2)==",R3_gene), x=-71.45, y=0.2635, parse=T) +
scale_x_reverse()
```

```{r}
#Plot Ho vs Longitude
R3_gene = round(summary(lm(y_gene$Ho_gene ~ y_gene$longitude))$r.squared, 4)             
ggplot(y_gene, aes(x = longitude, y = Ho_gene)) + geom_point(aes(shape=pop_order, color=pop_order), size = 3) + 
  scale_shape_manual(values = c(15,18,17,16)) +
geom_smooth(method=lm) + 
ggtitle("Observed heterozygosity vs Longitude, Neutral Gene SNPs") +
annotate(geom = "text", label=paste("italic(R^2)==",R3_gene), x=-71.45, y=0.2635, parse=T) +
scale_x_reverse()
```

```{r}
#Plot He vs Distance from sewage outflow 
R4_gene = round(summary(lm(y_gene$He_gene ~ y_gene$distance))$r.squared, 4)             
ggplot(y_gene, aes(x = distance, y = He_gene)) + geom_point(aes(shape=pop_order, color=pop_order), size = 3) + 
  scale_shape_manual(values = c(15,18,17,16)) + 
geom_smooth(method=lm) + 
ggtitle("Expected heterozygosity vs Distance, Neutral Gene SNPs") +
annotate(geom = "text", label=paste("italic(R^2)==",R4_gene), x=2, y=0.2685, parse=T) +
scale_x_reverse()
```

```{r}
#Plot Ho vs Distance from sewage outflow 
R4_gene = round(summary(lm(y_gene$Ho_gene ~ y_gene$distance))$r.squared, 4)             
ggplot(y_gene, aes(x = distance, y = Ho_gene)) + geom_point(aes(shape=pop_order, color=pop_order), size = 3) + 
  scale_shape_manual(values = c(15,18,17,16)) + 
geom_smooth(method=lm) + 
ggtitle("Observed heterozygosity vs Distance, Neutral Gene SNPs") +
annotate(geom = "text", label=paste("italic(R^2)==",R4_gene), x=2, y=0.2635, parse=T) +
scale_x_reverse()
```

```{r}
#Plot He vs Sewage effluent  
R_gene = round(summary(lm(y_gene$He_gene ~ y_gene$sewage))$r.squared, 4)             
ggplot(y_gene, aes(x = sewage, y = He_gene)) + geom_point(aes(shape=pop_order, color=pop_order), size = 3) + 
  scale_shape_manual(values = c(15,18,17,16)) + 
geom_smooth(method=lm) + 
ggtitle("Expected heterozygosity vs Sewage Effluent, Neutral Gene SNPs") +
annotate(geom = "text", label=paste("italic(R^2)==",R_gene), x=7, y=0.265, parse=T) +
scale_x_reverse()
```

```{r}
#Plot Ho vs Sewage effluent  
R_gene = round(summary(lm(y_gene$Ho_gene ~ y_gene$sewage))$r.squared, 4)             
ggplot(y_gene, aes(x = sewage, y = Ho_gene)) + geom_point(aes(shape=pop_order, color=pop_order), size = 3) + 
  scale_shape_manual(values = c(15,18,17,16)) + 
geom_smooth(method=lm) + 
ggtitle("Observed heterozygosity vs Sewage Effluent, Neutral Gene SNPs") +
annotate(geom = "text", label=paste("italic(R^2)==",R_gene), x=7, y=0.26, parse=T) +
scale_x_reverse()
```

```{r}
#Plot He vs Temperature 
R5_gene = round(summary(lm(y_gene$He_gene ~ y_gene$temperature))$r.squared, 4)             
ggplot(y_gene, aes(x = temperature, y = He_gene)) + geom_point(aes(shape=pop_order, color=pop_order), size = 3) + 
  scale_shape_manual(values = c(15,18,17,16)) +
geom_smooth(method=lm) + 
ggtitle("Expected heterozygosity vs Temperature, Neutral Gene SNPs") +
annotate(geom = "text", label=paste("italic(R^2)==",R5_gene), x=23.25, y=0.2635, parse=T) +
scale_x_reverse()
```

```{r}
#Plot Ho vs Temperature 
R5_gene = round(summary(lm(y_gene$Ho_gene ~ y_gene$temperature))$r.squared, 4)             
ggplot(y_gene, aes(x = temperature, y = Ho_gene)) + geom_point(aes(shape=pop_order, color=pop_order), size = 3) + 
  scale_shape_manual(values = c(15,18,17,16)) + 
geom_smooth(method=lm) + 
ggtitle("Observed heterozygosity vs Temperature, Neutral Gene SNPs") +
annotate(geom = "text", label=paste("italic(R^2)==",R5_gene), x=23.25, y=0.2635, parse=T) +
scale_x_reverse()
```

```{r}
#Plot He vs Salinity 
R6_gene = round(summary(lm(y_gene$He_gene ~ y_gene$salinity))$r.squared, 4)             
ggplot(y_gene, aes(x = salinity, y = He_gene)) + geom_point(aes(shape=pop_order, color=pop_order), size = 3) + 
  scale_shape_manual(values = c(15,18,17,16)) + 
geom_smooth(method=lm) + 
ggtitle("Expected heterozygosity vs Salinity, Neutral Gene SNPs") +
annotate(geom = "text", label=paste("italic(R^2)==",R6_gene), x=23.25, y=0.2635, parse=T) +
scale_x_reverse()
```

```{r}
#Plot Ho vs Salinity 
R6_gene = round(summary(lm(y_gene$Ho_gene ~ y_gene$salinity))$r.squared, 4)             
ggplot(y_gene, aes(x = salinity, y = Ho_gene)) + geom_point(aes(shape=pop_order, color=pop_order), size = 3) + 
  scale_shape_manual(values = c(15,18,17,16)) + 
geom_smooth(method=lm) + 
ggtitle("Observed heterozygosity vs Salinity, Neutral Gene SNPs") +
annotate(geom = "text", label=paste("italic(R^2)==",R6_gene), x=23.25, y=0.2635, parse=T) +
scale_x_reverse()
```

```{r}
#Plot He vs pH
R7_gene = round(summary(lm(y_gene$He_gene ~ y_gene$pH))$r.squared, 4)             
ggplot(y_gene, aes(x = pH, y = He_gene)) + geom_point(aes(shape=pop_order, color=pop_order), size = 3) + 
  scale_shape_manual(values = c(15,18,17,16)) +
geom_smooth(method=lm) + 
ggtitle("Expected heterozygosity vs pH, Neutral Gene SNPs") +
annotate(geom = "text", label=paste("italic(R^2)==",R7_gene), x=7.45, y=0.2635, parse=T) +
scale_x_reverse()
```

```{r}
#Plot Ho vs pH
R7_gene = round(summary(lm(y_gene$Ho_gene ~ y_gene$pH))$r.squared, 4)             
ggplot(y_gene, aes(x = pH, y = Ho_gene)) + geom_point(aes(shape=pop_order, color=pop_order), size = 3) + 
  scale_shape_manual(values = c(15,18,17,16)) + 
geom_smooth(method=lm) + 
ggtitle("Observed heterozygosity vs pH, Neutral Gene SNPs") +
annotate(geom = "text", label=paste("italic(R^2)==",R7_gene), x=7.45, y=0.2635, parse=T) +
scale_x_reverse()
```

```{r}
#Plot He vs Chlorophyll a 
R8_gene = round(summary(lm(y_gene$He_gene ~ y_gene$Chlor_a))$r.squared, 4)             
ggplot(y_gene, aes(x = Chlor_a, y = He_gene)) + geom_point(aes(shape=pop_order, color=pop_order), size = 3) + 
  scale_shape_manual(values = c(15,18,17,16)) + 
geom_smooth(method=lm) + 
ggtitle("Expected heterozygosity vs Chlorophyll a, Neutral Gene SNPs") +
annotate(geom = "text", label=paste("italic(R^2)==",R8_gene), x=6, y=0.2665, parse=T) +
scale_x_reverse()
```

```{r}
#Plot Ho vs Chlorophyll a 
R8_gene = round(summary(lm(y_gene$Ho_gene ~ y_gene$Chlor_a))$r.squared, 4)             
ggplot(y_gene, aes(x = Chlor_a, y = Ho_gene)) + geom_point(aes(shape=pop_order, color=pop_order), size = 3) + 
  scale_shape_manual(values = c(15,18,17,16)) + 
geom_smooth(method=lm) + 
ggtitle("Observed heterozygosity vs Chlorophyll a, Neutral Gene SNPs") +
annotate(geom = "text", label=paste("italic(R^2)==",R8_gene), x=6, y=0.2665, parse=T) +
scale_x_reverse()
```

```{r}
#Plot He vs Dissolved Oxygen 
R9_gene = round(summary(lm(y_gene$He_gene ~ y_gene$DO))$r.squared, 4)             
ggplot(y_gene, aes(x = DO, y = He_gene)) + geom_point(aes(shape=pop_order, color=pop_order), size = 3) + 
  scale_shape_manual(values = c(15,18,17,16)) + 
geom_smooth(method=lm) + 
ggtitle("Expected heterozygosity vs Dissolved Oxygen, Neutral Gene SNPs") +
annotate(geom = "text", label=paste("italic(R^2)==",R9_gene), x=5.5, y=0.2635, parse=T) +
scale_x_reverse()
```

```{r}
#Plot Ho vs Dissolved Oxygen 
R9_gene = round(summary(lm(y_gene$Ho_gene ~ y_gene$DO))$r.squared, 4)             
ggplot(y_gene, aes(x = DO, y = Ho_gene)) + geom_point(aes(shape=pop_order, color=pop_order), size = 3) + 
  scale_shape_manual(values = c(15,18,17,16)) + 
geom_smooth(method=lm) + 
ggtitle("Observed heterozygosity vs Dissolved Oxygen, Neutral Gene SNPs") +
annotate(geom = "text", label=paste("italic(R^2)==",R9_gene), x=5.5, y=0.2635, parse=T) +
scale_x_reverse()
```


## PCA 

```{r}
X_gene <- tab(stratted.u_gene, freq = TRUE, NA.method = "mean")
pca1_gene <- dudi.pca(X_gene, scale = FALSE, scannf = FALSE, nf = 3)
barplot(pca1_gene$eig[1:50], main = "PCA eigenvalues", col = heat.colors(50))
s.class(pca1_gene$li, pop(stratted.filt_gene))
title("PCA of neutral gene SNPs\naxes 1-2")
add.scatter.eig(pca1_gene$eig[1:20], 3,1,2)
col <- funky(15) 
s.class(pca1_gene$li, pop(stratted.filt_gene),xax=1,yax=2, col=col, axesell=FALSE, cstar=0, cpoint=3, grid=FALSE)
```


## Discriminant Analysis of Principal Components (DAPC)

The first line of the code block below is interactive. In the console, it will tell you:  
1. Choose the number PCs to retain (>=1): 
    I picked 1 
2. Choose the number discriminant functions to retain (>=1):
    I picked 2

```{r}
grp_gene <- find.clusters(stratted.u_gene, n.pca = 1, choose.n.clust = FALSE)
table(pop(stratted.u_gene), grp_gene$grp)
table.value(table(pop(stratted.u_gene), grp_gene$grp), col.lab=paste("inf", 1:2), row.lab=paste("ori", 1:4))
```

Again, the first line of the code block below is interactive. In the console, it will tell you:  
1. Choose the number PCs to retain (>=1): 
    I picked 1 
2. Choose the number discriminant functions to retain (>=1):
    I picked 1

```{r}
dapc1_gene <- dapc(stratted.u_gene, grp_gene$grp, n.pca = 1, n.da= 1)
scatter(dapc1_gene,col=col,bg="white", solid=1)
```

Again, the first line of the code block below is interactive. In the console, it will tell you:  
1. Choose the number PCs to retain (>=1): 
    I picked 1 
2. Choose the number discriminant functions to retain (>=1):
    I picked 1

```{r}
contrib_gene <- loadingplot(dapc1_gene$var.contr, axis=1, thres=.01, lab.jitter=1)
contrib_gene
setPop(rad.filt_gene) <- ~Library
dapc1_gene <- dapc(stratted.u_gene, pop(stratted.u_gene), n.pca = 1, n.da = 1)
contrib_gene <- loadingplot(dapc1_gene$var.contr, axis=1, thres=.05, lab.jitter=1)
```

```{r}
#Structure Like
compoplot(dapc1_gene, posi="bottomright",txt.leg=paste("Cluster", 1:4), lab="", ncol=1, xlab="individuals")
temp_gene <- which(apply(dapc1_gene$posterior,1, function(e) all(e<0.9)))
compoplot(dapc1_gene, subset=temp_gene, posi="bottomright", txt.leg=paste("Cluster", 1:4), ncol=2)
```

### PCAviz

```{r}
NA.afDraw<- function(ind){
  ind.mat <- ind@tab
  new.mat <- ind.mat
  af = colSums(ind.mat[,seq(1,ncol(ind.mat)-1,2)],na.rm = TRUE)/
      (2*apply(ind.mat[,seq(1,ncol(ind.mat)-1,2)],2,function(x) sum(!is.na(x))))
  af.Draw <- function(geno, af){
     new <- function(geno,af){
        if(is.na(geno)){
        newA = rbinom(1,2,af)
        }
        else {newA <- geno}
        return(newA)
   }
  new.row <- mapply(geno,af,FUN = new)
  return(new.row)}
  
  new.mat[,seq(1,ncol(ind.mat)-1,2)] <- t(apply(ind.mat[,seq(1,ncol(ind.mat)-1,2)],1,af.Draw,af))
  new.mat[,seq(2,ncol(ind.mat),2)] <- 2-new.mat[,seq(1,ncol(ind.mat)-1,2)]
  new.ind <- ind
  new.ind@tab <- new.mat
  return(new.ind)
}
```

```{r}
u.na_gene <- NA.afDraw(stratted.u_gene)
```

```{r}
pca_gene <- dudi.pca(u.na_gene,center=TRUE,scale=TRUE,scannf = F, nf = 30)
```

```{r}
col18 <- funky(length(unique(u.na_gene@strata$Population)))
#Colors that match the neutral Structure results
col6 <-  c("firebrick","dodgerblue","midnightblue","salmon")
par(mfrow=c(2,2))
s.class(pca_gene$li, strata(u.na_gene)$Population,xax=1,yax=2, 
        sub = "neutralgeneloci,PC 1-2, 40 individuals, 11,620 loci, maf 5.0%",
        possub = "topleft",col=transp(col6,.6),axesell=FALSE,
        cstar=0, cpoint=3, grid=FALSE, cellipse = 0)
s.class(pca_gene$li, strata(u.na_gene)$Population,xax=1,yax=3, 
        sub = "neutralgeneloci,PC 1-3, 40 individuals, 11,620 loci, maf 5.0%",
        possub = "topleft",col=transp(col6,.6),axesell=FALSE,
        cstar=0, cpoint=3, grid=FALSE, cellipse = 0)
s.class(pca_gene$li, strata(u.na_gene)$Population,xax=1,yax=4, 
        sub = "neutralgeneloci,PC 1-4, 40 individuals, 11,620 loci, maf 5.0%",
        possub = "topleft",col=transp(col6,.6),axesell=FALSE,
        cstar=0, cpoint=3, grid=FALSE, cellipse = 0)
s.class(pca_gene$li, strata(u.na_gene)$Population,xax=1,yax=5, 
        sub = "neutralgeneloci,PC 1-5, 40 individuals, 11,620 loci, maf 5.0%",
        possub = "topleft",col=transp(col6,.6),axesell=FALSE,
        cstar=0, cpoint=3, grid=FALSE, cellipse = 0)
```



# Neutral Intergenic SNPs

****

## Making files

**Make genind object**

`neutralintergenic.recode.vcf` contains neutral SNPs from intergenic regions for populations PVD, GB, BIS, and NAR. Steps for generating this VCF file are located in [EecSeq_Cvirginica_NeutralLoci.md](https://github.com/amyzyck/EecSeq_NB_EasternOyster/blob/master/Analysis/PopGen_SeaGen_Analyses/NonHaplotigMasked_Genome/Neutral_SNPs/EecSeq_Cvirginica_NeutralLoci.md).

Population NIN was removed from the VCF file following steps in [EecSeq_Cvirginica_OutlierDetection.md](https://github.com/amyzyck/EecSeq_NB_EasternOyster/blob/master/Analysis/EecSeq_Cvirginica_OutlierDetection.md).

`strata` contains population, environmental, and library information for each sample - can be accessed [here](https://github.com/amyzyck/EecSeq_NB_EasternOyster/blob/master/Analysis/PopGen_SeaGen_Analyses/strata).

```{r}
my_vcf_inter <- read.vcfR("neutralintergenic.recode.vcf")
strata <- read.table("strata", header=TRUE)
```

```{r}
rad.filt_inter <- vcfR2genind(my_vcf_inter, strata = strata, pop = c(rep("BIS", 10),rep("GB", 10),rep("NAR", 10), rep("PVD",10)))
```

```{r}
rad.filt_inter
```

```{r}
#Providing population names for plotting
pop_order <- c("BIS","GB","NAR","PVD")
```

**Read in the other info from .strata file and extract information such as locality, latitude, and longitude.**
```{r}
info <- as.data.frame(read.table("strata",header = T,sep = "\t",stringsAsFactors = F))
mystrats_inter <- as.data.frame(matrix(nrow = length(indNames(rad.filt_inter)),ncol=10))
colnames(mystrats_inter) <- c("Population","Latitude","Longitude","Distance","Temperature","SE","Salinity","pH","Chlorophylla","DO")
just.strats <- select(info,c("Population"))
stratted.filt_inter <- strata(rad.filt_inter, formula= Population, combine = TRUE,just.strats)
stratted.filt_inter@other <- select(info, Latitude,Longitude,Distance,SE,Temperature,Salinity,pH,Chlorophylla,DO)
```

```{r}
stratted.filt_inter
```

**Repeat for quasi-independent set of SNPs**

Steps for filtering are documented for OutFLANK in [EecSeq_Cvirginica_OutlierDetection.md](https://github.com/amyzyck/EecSeq_NB_EasternOyster/blob/master/Analysis/EecSeq_Cvirginica_OutlierDetection.md). However, I am repeating the steps here. 

```{r}
my_vcf_inter <- read.vcfR("neutralintergenic.recode.vcf")

geno_inter <- extract.gt(my_vcf_inter) # Character matrix containing the genotypes
position_inter <- getPOS(my_vcf_inter) # Positions in bp
chromosome_inter <- getCHROM(my_vcf_inter) # Chromosome information

G_inter <- matrix(NA, nrow = nrow(geno_inter), ncol = ncol(geno_inter))

G_inter[geno_inter %in% c("0/0", "0|0")] <- 0
G_inter[geno_inter  %in% c("0/1", "1/0", "1|0", "0|1")] <- 1
G_inter[geno_inter %in% c("1/1", "1|1")] <- 2

# NA should be replaced with “9”
G_inter[is.na(G_inter)] <- 9
```


Chromosomes need to be of class integer for this to work.
```{r}
# Visualizing chromosomes
chrom_unique_inter <- unique(chromosome_inter)
print(chrom_unique_inter)
```

```{r}
# Removing "NC_" from the chromosome name so it can be converted to an integer
chrom_new_inter <- chromosome_inter%>%str_replace("NC_","")

# Converting the character string into an integer
chrom1_inter <- as.integer(chrom_new_inter)
```

`chromosome` and `position` need to be sorted for imputation to work.

```{r}
chrom_sort_inter <- sort(chrom1_inter)
pos_sort_inter <- sort(position_inter)
```

_Note: This filtering program does not allow for missing genotype values._


### Remove missing genotypes

```{r}
# removing missing data 
G_inter_miss <- matrix(NA, nrow = nrow(geno_inter), ncol = ncol(geno_inter))

G_inter_miss[geno_inter %in% c("0/0", "0|0")] <- 0
G_inter_miss[geno_inter  %in% c("0/1", "1/0", "1|0", "0|1")] <- 1
G_inter_miss[geno_inter %in% c("1/1", "1|1")] <- 2

# removing missing data
G_inter_miss[na.omit(G_inter_miss)]
```


```{r}
# Converting the genotype matrix to class FBM.code256 using the matrix where missing data was removed
G1_inter_miss <- add_code256(big_copy(t(G_inter_miss),type="raw"),code=bigsnpr:::CODE_012)
```

```{r}
newpc_inter_miss <- snp_autoSVD(G1_inter_miss,infos.chr =chrom_sort_inter,infos.pos = pos_sort_inter, roll.size = 35) # roll.size = Radius of rolling windows to smooth log-p-values
# maximum roll.size I could use is 35
which_pruned_inter_miss <- attr(newpc_inter_miss, which="subset") # Indexes of remaining SNPS after pruning
length(which_pruned_inter_miss)
```


```{r}
invisible(lapply(which_pruned_inter_miss, write, "pruned_data_inter.txt", append=TRUE))
```


**In terminal**

```
$ mawk '!/#/' neutralintergenic.recode.vcf | cut -f1,2 > totallociinter
$ NUM=(`cat totallociinter | wc -l`)
$ paste <(seq 1 $NUM) totallociinter > lociinter.plus.index
$ cat pruned_data_inter.txt | parallel "grep -w ^{} lociinter.plus.index" | cut -f2,3> pruned_data_inter.loci.txt

$ head pruned_data_inter.loci.txt

output:
NC_035780.1	623558
NC_035780.1	623903
NC_035780.1	623906
NC_035780.1	1490085
NC_035780.1	1490100
NC_035780.1	1695081
NC_035780.1	1695087
NC_035780.1	1756507
NC_035780.1	1882142
NC_035780.1	2425975
```

Create VCF file with just the pruned_data loci 
```
$ vcftools --vcf neutralintergenic.recode.vcf --recode --recode-INFO-all --positions pruned_data_inter.loci.txt --out pruned_data_inter

output:
After filtering, kept 40 out of 40 Individuals
Outputting VCF file...
After filtering, kept 2019 out of a possible 7581 Sites
Run Time = 1.00 seconds
```

```{r}
my_vcf_u_inter <- read.vcfR("pruned_data_inter.recode.vcf")
strata <- read.table("strata", header=TRUE)
```

```{r}
rad.u_inter <- vcfR2genind(my_vcf_u_inter, strata = strata, pop = c(rep("BIS", 10),rep("GB", 10),rep("NAR", 10), rep("PVD",10)))
```

```{r}
rad.u_inter
```


```{r}
stratted.u_inter <- strata(rad.u_inter, formula= Population, combine = TRUE,just.strats)
stratted.u_inter@other <- select(info, Latitude,Longitude, Distance, SE,  Temperature,Salinity,pH,Chlorophylla,DO)
```

```{r}
stratted.u_inter
```

**Make hierfstat object**
```{r}
hf.filt_inter <- genind2hierfstat(rad.filt_inter, pop = c(rep("BIS", 10),rep("GB", 10),rep("NAR", 10), rep("PVD",10))) 
```

```{r}
hf.u_inter <- genind2hierfstat(rad.u_inter, pop = c(rep("BIS", 10),rep("GB", 10),rep("NAR", 10), rep("PVD",10)))
hf.u_inter <- hf.u_inter$hierfstat.no.imputation
```

## Estimating effective migration surfaces [(EEMS)](https://github.com/dipetkov/eems)

__1. datapath.diffs__

**Steps for generating `datapath.diffs` are completed in RStudio.**
```{r}
# V1 methtod to get diffs matrix, preferred
bed2diffs_v1 <- function(Geno) {
  nIndiv <- nrow(Geno)
  nSites <- ncol(Geno)
  Diffs <- matrix(0, nIndiv, nIndiv)
  
  for (i in seq(nIndiv - 1)) {
    for (j in seq(i + 1, nIndiv)) {
      x <- Geno[i, ]
      y <- Geno[j, ]
      Diffs[i, j] <- mean((x - y)^2, na.rm = TRUE)
      Diffs[j, i] <- Diffs[i, j]
    }
  }
  Diffs
}
```

```{r}
# V2 method to get .diffs matrix, only if V1 doesn't work
bed2diffs_v2 <- function(Geno) {
  nIndiv <- nrow(Geno)
  nSites <- ncol(Geno)
  Miss <- is.na(Geno)
  ## Impute NAs with the column means (= twice the allele frequencies)
  Mean <- matrix(colMeans(Geno, na.rm = TRUE), ## a row of means
                 nrow = nIndiv, ncol = nSites, byrow = TRUE) ## a matrix with nIndiv identical rows of means
  Mean[Miss == 0] <- 0 ## Set the means that correspond to observed genotypes to 0
  Geno[Miss == 1] <- 0 ## Set the missing genotypes to 0 (used to be NA) 
  Geno <- Geno + Mean
  ## Compute similarities
  Sim <- Geno %*% t(Geno) / nSites
  SelfSim <- diag(Sim) ## self-similarities
  vector1s <- rep(1, nIndiv) ## vector of 1s
  ## This chunk generates a `diffs` matrix
  Diffs <- SelfSim %*% t(vector1s) + vector1s %*% t(SelfSim) - 2 * Sim
  Diffs
}
```

```{r}
geno_inter <- stratted.filt_inter@tab
```

```{r}
# Get rid of non-biallelic loci
multi.loci_inter <- names(which(stratted.filt_inter@loc.n.all != 2))
multi.cols_inter <- which(grepl(paste0("^", multi.loci_inter, "\\.\\d+$", collapse = "|"), colnames(geno_inter)))
if (length(multi.cols_inter)) geno_inter <- geno_inter[, - multi.cols_inter]
nloci_inter <- dim(geno_inter)[2] / 2
dim(geno_inter)
```

```{r}
stopifnot(identical(stratted.filt_inter@type, 'codom'))
```

```{r}
# bed2diffs functions  
diffs.v1_inter <- bed2diffs_v1(geno_inter)
diffs.v2_inter <- bed2diffs_v2(geno_inter)
# Round to 6 digits
diffs.v1_inter <- round(diffs.v1_inter, digits = 6)
diffs.v2_inter <- round(diffs.v2_inter, digits = 6)
```

**Check that the dissimilarity matrix has one positive eigenvalue and nIndiv-1 negative eigenvalues, as required by a full-rank Euclidean distance matrix. If the V1 method does not make a Euclidean matrix, you must use V2.**

```{r}
tail(sort(round(eigen(diffs.v1_inter)$values, digits = 2)))
```

```{r}
tail(sort(round(eigen(diffs.v2_inter)$values, digits = 2)))
```

```{r}
# Set suffix for EEMS input files
suf_inter <- "neutralinterdata-filt"
```

```{r}
# This saves the file to directory 
write.table(diffs.v1_inter, paste(suf_inter,".v1.diffs",sep=""), 
            col.names = FALSE, row.names = FALSE, quote = FALSE)
```


__2. datapath.coord__

`datapath.coord` are the sample coordinates, two coordinates per sample, one sample per line. The sampling locations should be given in the same order as the rows and columns of the dissimilarity matrix.

**Steps for generating `datapath.coord` are completed in RStudio.**
```{r}
## Get gps coordinates from previously created info matrix
xOR.info <- dplyr::filter(info)
gps_matrix <- select(xOR.info,c("Longitude","Latitude"))
```

```{r}
#write .coord file
write.table(gps_matrix, paste(suf_inter,".v1.coord",sep=""),col.names = FALSE, row.names = FALSE,quote = FALSE)
```

__3. datapath.outer__

`datapath.outer` are the habitat coordinates, as a sequence of vertices that form a closed polygon. The habitat vertices should be listed counterclockwise and the first vertex should also be the last vertex, so that the outline is a closed ring. Otherwise, EEMS attempts to "correct" the polygon and prints a warning message.

**`datapath.outer` is created manually in Excel,based on site coordinates gathered from Google Maps, copied into terminal using `nano`, and saved as the file with the appropriate extension.** 


**`runeems_snps` is then run in command-line following the steps documented in [NB_EEMS_OutlierHap.md](https://github.com/amyzyck/EecSeq_NB_EasternOyster/blob/master/Analysis/PopGen_SeaGen_Analyses/EEMS/NB_EEMS_OutlierHap.md)  

**Back in RStudio to plot `runeems_snps` outputs**

```{r}
# Install rEEMSplots
library(rEEMSplots)
```

```{r} 
# Plotting EEMS after running runeems_snps
path = "./NB_EEMS_Neutral/"
dirs = c(paste0(path,"neutralinterdata-D200-chain1"), paste0(path,"neutralinterdata-D300-chain1"), paste0(path,"neutralinterdata-D600-chain1"))
```

```{r}
eems.plots(mcmcpath = c(paste0(path,"./neutralinterdata-D200-chain1"), paste0(path,"neutralinterdata-D300-chain1"), paste0(path,"neutralinterdata-D600-chain1")), plotpath = paste0(path,"neutralinterdata-All-plots"),
           longlat = T,add.grid=F,add.outline = T,add.demes = T,
           projection.in = "+proj=longlat +datum=WGS84",projection.out = "+proj=merc +datum=WGS84",
           add.map = T,add.abline = T, add.r.squared = T)
```

## Pairwise Fst

```{r}
fst.mat_inter <- pairwise.WCfst(hf.filt_inter)
```

```{r}
gindF.fst.mat.triN_inter <- as.matrix(fst.mat_inter)
colnames(gindF.fst.mat.triN_inter) <- pop_order
rownames(gindF.fst.mat.triN_inter) <- pop_order
```

```{r}
meltedN_inter <- melt(gindF.fst.mat.triN_inter, na.rm =TRUE)
round(gindF.fst.mat.triN_inter,4)
```

```{r}
summary(meltedN_inter$value)
```


```{r}
#Plotting Pairwise fst
neutral_inter <- ggplot(data = meltedN_inter, aes(Var2, Var1, fill = value))+ geom_tile(color = "white")+ 
  scale_fill_gradient(low = "white", high = "red", name="FST")  +
  ggtitle(expression(atop("Pairwise FST, WC (1984) Neutral Intergenic SNPs", atop(italic("N = 40, L = 7,581"), ""))))+
  labs( x = "Sampling Site", y = "Sampling Site") + 
  theme(axis.text.x = element_text(angle = 45, vjust = 1, size = 12, hjust = 1),axis.text.y = element_text(size = 13)) + 
  theme(axis.title = element_text(size = 16),legend.text = element_text(size =15), legend.title = element_text(size =16)) +
  theme(plot.title = element_text(size = 17)) +
coord_fixed()
neutral_inter
```

## Genetic diversity (observed and expected heterozygosity)

```{r}
comb_inter <- summary(stratted.filt_inter)
names(comb_inter)
```

```{r}
plot(comb_inter$Hobs, xlab="Loci number", ylab="Observed Heterozygosity", 
     main="Observed heterozygosity per locus")
```


```{r}
plot(comb_inter$Hobs,comb_inter$Hexp, xlab="Hobs", ylab="Hexp", 
     main="Expected heterozygosity as a function of observed heterozygosity per locus")
```

```{r}
bartlett.test(list(comb_inter$Hexp, comb_inter$Hobs)) # a test : H0: Hexp = Hobs
```

**_Significant difference between Observed and expected heterzygosity_**.  

```{r}
basicstat_inter <- basic.stats(hf.filt_inter, diploid = TRUE, digits = 3)
```

```{r}
as.data.frame(basicstat_inter$overall)
```

```{r}
# get bootstrap confidence values for Fis
boot_inter <- boot.ppfis(hf.filt_inter,nboot = 1000)
boot5_inter <- boot.ppfis(hf.filt_inter,nboot = 1000,quant = 0.5)
```

```{r}
# combine all pop statistics
colnames(basicstat_inter$Ho) <- pop_order
Ho_inter <- colMeans(basicstat_inter$Ho,na.rm = T)
He_inter <- colMeans(basicstat_inter$Hs,na.rm = T)
Fis_inter <- boot5_inter$fis.ci$ll
y_inter <- cbind(pop_order,Ho_inter,He_inter,Fis_inter,boot_inter$fis.ci,latitude, longitude, distance, sewage, temperature, salinity, pH, Chlor_a, DO)
y_inter
```

```{r}
summary(He_inter)
```

```{r}
summary(Fis_inter)
```

```{r}
# Plot He vs Latitude
R2_inter = round(summary(lm(y_inter$He_inter ~ y_inter$latitude))$r.squared, 4)             
ggplot(y_inter, aes(x = latitude, y = He_inter)) + geom_point(aes(shape=pop_order, color=pop_order), size = 3) + 
  scale_shape_manual(values = c(15,18,17,16)) +
geom_smooth(method=lm) + 
ggtitle("Expected heterozygosity vs Latitude, Neutral Intergenic SNPs") +
annotate(geom = "text", label=paste("italic(R^2)==",R2_inter), x=41.55, y=0.2635, parse=T) +
scale_x_continuous()
```

```{r}
# Plot Ho vs Latitude
R2_inter = round(summary(lm(y_inter$Ho_inter ~ y_inter$latitude))$r.squared, 4)             
ggplot(y_inter, aes(x = latitude, y = Ho_inter)) + geom_point(aes(shape=pop_order, color=pop_order), size = 3) + 
  scale_shape_manual(values = c(15,18,17,16)) + 
geom_smooth(method=lm) + 
ggtitle("Observed heterozygosity vs Latitude, Neutral Intergenic SNPs") +
annotate(geom = "text", label=paste("italic(R^2)==",R2_inter), x=41.55, y=0.28, parse=T) +
scale_x_continuous()
```

```{r}
#Plot He vs Longitude
R3_inter = round(summary(lm(y_inter$He_inter ~ y_inter$longitude))$r.squared, 4)             
ggplot(y_inter, aes(x = longitude, y = He_inter)) + geom_point(aes(shape=pop_order, color=pop_order), size = 3) + 
  scale_shape_manual(values = c(15,18,17,16)) +
geom_smooth(method=lm) + 
ggtitle("Expected heterozygosity vs Longitude, Neutral Intergenic SNPs") +
annotate(geom = "text", label=paste("italic(R^2)==",R3_inter), x=-71.44, y=0.2635, parse=T) +
scale_x_reverse()
```

```{r}
#Plot Ho vs Longitude
R3_inter = round(summary(lm(y_inter$Ho_inter ~ y_inter$longitude))$r.squared, 4)             
ggplot(y_inter, aes(x = longitude, y = Ho_inter)) + geom_point(aes(shape=pop_order, color=pop_order), size = 3) + 
  scale_shape_manual(values = c(15,18,17,16)) +
geom_smooth(method=lm) + 
ggtitle("Observed heterozygosity vs Longitude, Neutral Integenic SNPs") +
annotate(geom = "text", label=paste("italic(R^2)==",R3_inter), x=-71.44, y=0.28, parse=T) +
scale_x_reverse()
```

```{r}
#Plot He vs Distance from sewage outflow 
R4_inter = round(summary(lm(y_inter$He_inter ~ y_inter$distance))$r.squared, 4)             
ggplot(y_inter, aes(x = distance, y = He_inter)) + geom_point(aes(shape=pop_order, color=pop_order), size = 3) + 
  scale_shape_manual(values = c(15,18,17,16)) + 
geom_smooth(method=lm) + 
ggtitle("Expected heterozygosity vs Distance, Neutral Intergenic SNPs") +
annotate(geom = "text", label=paste("italic(R^2)==",R4_inter), x=2, y=0.265, parse=T) +
scale_x_reverse()
```

```{r}
#Plot Ho vs Distance from sewage outflow 
R4_inter = round(summary(lm(y_inter$Ho_inter ~ y_inter$distance))$r.squared, 4)             
ggplot(y_inter, aes(x = distance, y = Ho_inter)) + geom_point(aes(shape=pop_order, color=pop_order), size = 3) + 
  scale_shape_manual(values = c(15,18,17,16)) + 
geom_smooth(method=lm) + 
ggtitle("Observed heterozygosity vs Distance, Neutral Intergenic SNPs") +
annotate(geom = "text", label=paste("italic(R^2)==",R4_inter), x=2, y=0.28, parse=T) +
scale_x_reverse()
```

```{r}
#Plot He vs Sewage effluent  
R_inter = round(summary(lm(y_inter$He_inter ~ y_inter$sewage))$r.squared, 4)             
ggplot(y_inter, aes(x = sewage, y = He_inter)) + geom_point(aes(shape=pop_order, color=pop_order), size = 3) + 
  scale_shape_manual(values = c(15,18,17,16)) + 
geom_smooth(method=lm) + 
ggtitle("Expected heterozygosity vs Sewage Effluent, Neutral Intergenic SNPs") +
annotate(geom = "text", label=paste("italic(R^2)==",R_inter), x=7, y=0.265, parse=T) +
scale_x_reverse()
```

```{r}
#Plot Ho vs Sewage effluent  
R_inter = round(summary(lm(y_inter$Ho_inter ~ y_inter$sewage))$r.squared, 4)             
ggplot(y_inter, aes(x = sewage, y = Ho_inter)) + geom_point(aes(shape=pop_order, color=pop_order), size = 3) + 
  scale_shape_manual(values = c(15,18,17,16)) + 
geom_smooth(method=lm) + 
ggtitle("Observed heterozygosity vs Sewage Effluent, Neutral Intergenic SNPs") +
annotate(geom = "text", label=paste("italic(R^2)==",R_inter), x=7, y=0.28, parse=T) +
scale_x_reverse()
```

```{r}
#Plot He vs Temperature 
R5_inter = round(summary(lm(y_inter$He_inter ~ y_inter$temperature))$r.squared, 4)             
ggplot(y_inter, aes(x = temperature, y = He_inter)) + geom_point(aes(shape=pop_order, color=pop_order), size = 3) + 
  scale_shape_manual(values = c(15,18,17,16)) +
geom_smooth(method=lm) + 
ggtitle("Expected heterozygosity vs Temperature, Neutral Intergenic SNPs") +
annotate(geom = "text", label=paste("italic(R^2)==",R5_inter), x=23.25, y=0.2635, parse=T) +
scale_x_reverse()
```

```{r}
#Plot Ho vs Temperature 
R5_inter = round(summary(lm(y_inter$Ho_inter ~ y_inter$temperature))$r.squared, 4)             
ggplot(y_inter, aes(x = temperature, y = Ho_inter)) + geom_point(aes(shape=pop_order, color=pop_order), size = 3) + 
  scale_shape_manual(values = c(15,18,17,16)) + 
geom_smooth(method=lm) + 
ggtitle("Observed heterozygosity vs Temperature, Neutral Intergenic SNPs") +
annotate(geom = "text", label=paste("italic(R^2)==",R5_inter), x=23.25, y=0.28, parse=T) +
scale_x_reverse()
```

```{r}
#Plot He vs Salinity 
R6_inter = round(summary(lm(y_inter$He_inter ~ y_inter$salinity))$r.squared, 4)             
ggplot(y_inter, aes(x = salinity, y = He_inter)) + geom_point(aes(shape=pop_order, color=pop_order), size = 3) + 
  scale_shape_manual(values = c(15,18,17,16)) + 
geom_smooth(method=lm) + 
ggtitle("Expected heterozygosity vs Salinity, Neutral Intergenic SNPs") +
annotate(geom = "text", label=paste("italic(R^2)==",R6_inter), x=23.25, y=0.2635, parse=T) +
scale_x_reverse()
```

```{r}
#Plot Ho vs Salinity 
R6_inter = round(summary(lm(y_inter$Ho_inter ~ y_inter$salinity))$r.squared, 4)             
ggplot(y_inter, aes(x = salinity, y = Ho_inter)) + geom_point(aes(shape=pop_order, color=pop_order), size = 3) + 
  scale_shape_manual(values = c(15,18,17,16)) + 
geom_smooth(method=lm) + 
ggtitle("Observed heterozygosity vs Salinity, Neutral Intergenic SNPs") +
annotate(geom = "text", label=paste("italic(R^2)==",R6_inter), x=23.25, y=0.28, parse=T) +
scale_x_reverse()
```

```{r}
#Plot He vs pH
R7_inter = round(summary(lm(y_inter$He_inter ~ y_inter$pH))$r.squared, 4)             
ggplot(y_inter, aes(x = pH, y = He_inter)) + geom_point(aes(shape=pop_order, color=pop_order), size = 3) + 
  scale_shape_manual(values = c(15,18,17,16)) +
geom_smooth(method=lm) + 
ggtitle("Expected heterozygosity vs pH, Neutral Intergenic SNPs") +
annotate(geom = "text", label=paste("italic(R^2)==",R7_inter), x=7.45, y=0.2635, parse=T) +
scale_x_reverse()
```

```{r}
#Plot Ho vs pH
R7_inter = round(summary(lm(y_inter$Ho_inter ~ y_inter$pH))$r.squared, 4)             
ggplot(y_inter, aes(x = pH, y = Ho_inter)) + geom_point(aes(shape=pop_order, color=pop_order), size = 3) + 
  scale_shape_manual(values = c(15,18,17,16)) + 
geom_smooth(method=lm) + 
ggtitle("Observed heterozygosity vs pH, Neutral Intergenic SNPs") +
annotate(geom = "text", label=paste("italic(R^2)==",R7_inter), x=7.45, y=0.28, parse=T) +
scale_x_reverse()
```

```{r}
#Plot He vs Chlorophyll a 
R8_inter = round(summary(lm(y_inter$He_inter ~ y_inter$Chlor_a))$r.squared, 4)             
ggplot(y_inter, aes(x = Chlor_a, y = He_inter)) + geom_point(aes(shape=pop_order, color=pop_order), size = 3) + 
  scale_shape_manual(values = c(15,18,17,16)) + 
geom_smooth(method=lm) + 
ggtitle("Expected heterozygosity vs Chlorophyll a, Neutral Intergenic SNPs") +
annotate(geom = "text", label=paste("italic(R^2)==",R8_inter), x=8, y=0.27, parse=T) +
scale_x_reverse()
```

```{r}
#Plot Ho vs Chlorophyll a 
R8_inter = round(summary(lm(y_inter$Ho_inter ~ y_inter$Chlor_a))$r.squared, 4)             
ggplot(y_inter, aes(x = Chlor_a, y = Ho_inter)) + geom_point(aes(shape=pop_order, color=pop_order), size = 3) + 
  scale_shape_manual(values = c(15,18,17,16)) + 
geom_smooth(method=lm) + 
ggtitle("Observed heterozygosity vs Chlorophyll a, Neutral Intergenic SNPs") +
annotate(geom = "text", label=paste("italic(R^2)==",R8_inter), x=6, y=0.285, parse=T) +
scale_x_reverse()
```

```{r}
#Plot He vs Dissolved Oxygen 
R9_inter = round(summary(lm(y_inter$He_inter ~ y_inter$DO))$r.squared, 4)             
ggplot(y_inter, aes(x = DO, y = He_inter)) + geom_point(aes(shape=pop_order, color=pop_order), size = 3) + 
  scale_shape_manual(values = c(15,18,17,16)) + 
geom_smooth(method=lm) + 
ggtitle("Expected heterozygosity vs Dissolved Oxygen, Neutral Intergenic SNPs") +
annotate(geom = "text", label=paste("italic(R^2)==",R9_inter), x=5.5, y=0.265, parse=T) +
scale_x_reverse()
```

```{r}
#Plot Ho vs Dissolved Oxygen 
R9_inter = round(summary(lm(y_inter$Ho_inter ~ y_inter$DO))$r.squared, 4)             
ggplot(y_inter, aes(x = DO, y = Ho_inter)) + geom_point(aes(shape=pop_order, color=pop_order), size = 3) + 
  scale_shape_manual(values = c(15,18,17,16)) + 
geom_smooth(method=lm) + 
ggtitle("Observed heterozygosity vs Dissolved Oxygen, Neutral Intergenic SNPs") +
annotate(geom = "text", label=paste("italic(R^2)==",R9_inter), x=5.5, y=0.28, parse=T) +
scale_x_reverse()
```


## PCA 

```{r}
X_inter <- tab(stratted.u_inter, freq = TRUE, NA.method = "mean")
pca1_inter <- dudi.pca(X_inter, scale = FALSE, scannf = FALSE, nf = 3)
barplot(pca1_inter$eig[1:50], main = "PCA eigenvalues", col = heat.colors(50))
s.class(pca1_inter$li, pop(stratted.filt_inter))
title("PCA of neutral intergenic SNPs\naxes 1-2")
add.scatter.eig(pca1_inter$eig[1:20], 3,1,2)
col <- funky(15) 
s.class(pca1_inter$li, pop(stratted.filt_inter),xax=1,yax=2, col=col, axesell=FALSE, cstar=0, cpoint=3, grid=FALSE)
```


## Discriminant Analysis of Principal Components (DAPC)

The first line of the code block below is interactive. In the console, it will tell you:  
1. Choose the number PCs to retain (>=1): 
    I picked 1 
2. Choose the number discriminant functions to retain (>=1):
    I picked 2

```{r}
grp_inter <- find.clusters(stratted.u_inter, n.pca = 1, choose.n.clust = FALSE)
table(pop(stratted.u_inter), grp_inter$grp)
table.value(table(pop(stratted.u_inter), grp_inter$grp), col.lab=paste("inf", 1:2), row.lab=paste("ori", 1:4))
```

Again, the first line of the code block below is interactive. In the console, it will tell you:  
1. Choose the number PCs to retain (>=1): 
    I picked 1 
2. Choose the number discriminant functions to retain (>=1):
    I picked 1

```{r}
dapc1_inter <- dapc(stratted.u_inter, grp_inter$grp, n.pca = 1, n.da= 1)
scatter(dapc1_inter,col=col,bg="white", solid=1)
```

Again, the first line of the code block below is interactive. In the console, it will tell you:  
1. Choose the number PCs to retain (>=1): 
    I picked 1 
2. Choose the number discriminant functions to retain (>=1):
    I picked 1

```{r}
contrib_inter <- loadingplot(dapc1_inter$var.contr, axis=1, thres=.01, lab.jitter=1)
contrib_inter
setPop(rad.filt_inter) <- ~Library
dapc1_inter <- dapc(stratted.u_inter, pop(stratted.u_inter), n.pca = 1, n.da = 1)
contrib_inter <- loadingplot(dapc1_inter$var.contr, axis=1, thres=.05, lab.jitter=1)
```

```{r}
#Structure Like
compoplot(dapc1_inter, posi="bottomright",txt.leg=paste("Cluster", 1:4), lab="", ncol=1, xlab="individuals")
temp_inter <- which(apply(dapc1_inter$posterior,1, function(e) all(e<0.9)))
compoplot(dapc1_inter, subset=temp_inter, posi="bottomright", txt.leg=paste("Cluster", 1:4), ncol=2)
```

### PCAviz

```{r}
NA.afDraw<- function(ind){
  ind.mat <- ind@tab
  new.mat <- ind.mat
  af = colSums(ind.mat[,seq(1,ncol(ind.mat)-1,2)],na.rm = TRUE)/
      (2*apply(ind.mat[,seq(1,ncol(ind.mat)-1,2)],2,function(x) sum(!is.na(x))))
  af.Draw <- function(geno, af){
     new <- function(geno,af){
        if(is.na(geno)){
        newA = rbinom(1,2,af)
        }
        else {newA <- geno}
        return(newA)
   }
  new.row <- mapply(geno,af,FUN = new)
  return(new.row)}
  
  new.mat[,seq(1,ncol(ind.mat)-1,2)] <- t(apply(ind.mat[,seq(1,ncol(ind.mat)-1,2)],1,af.Draw,af))
  new.mat[,seq(2,ncol(ind.mat),2)] <- 2-new.mat[,seq(1,ncol(ind.mat)-1,2)]
  new.ind <- ind
  new.ind@tab <- new.mat
  return(new.ind)
}
```

```{r}
u.na_inter <- NA.afDraw(stratted.u_inter)
```

```{r}
pca_inter <- dudi.pca(u.na_inter,center=TRUE,scale=TRUE,scannf = F, nf = 30)
```

```{r}
col18 <- funky(length(unique(u.na_inter@strata$Population)))
#Colors that match the neutral Structure results
col6 <-  c("firebrick","dodgerblue","midnightblue","salmon")
par(mfrow=c(2,2))
s.class(pca_inter$li, strata(u.na_inter)$Population,xax=1,yax=2, 
        sub = "neutralintergenicloci,PC 1-2, 40 individuals, 2,019 loci, maf 5.0%",
        possub = "topleft",col=transp(col6,.6),axesell=FALSE,
        cstar=0, cpoint=3, grid=FALSE, cellipse = 0)
s.class(pca_inter$li, strata(u.na_inter)$Population,xax=1,yax=3, 
        sub = "neutralintergenicloci,PC 1-3, 40 individuals, 2,019 loci, maf 5.0%",
        possub = "topleft",col=transp(col6,.6),axesell=FALSE,
        cstar=0, cpoint=3, grid=FALSE, cellipse = 0)
s.class(pca_inter$li, strata(u.na_inter)$Population,xax=1,yax=4, 
        sub = "neutralintergenicloci,PC 1-4, 40 individuals, 2,019 loci, maf 5.0%",
        possub = "topleft",col=transp(col6,.6),axesell=FALSE,
        cstar=0, cpoint=3, grid=FALSE, cellipse = 0)
s.class(pca_inter$li, strata(u.na_inter)$Population,xax=1,yax=5, 
        sub = "neutralintergenicloci,PC 1-5, 40 individuals, 2,019 loci, maf 5.0%",
        possub = "topleft",col=transp(col6,.6),axesell=FALSE,
        cstar=0, cpoint=3, grid=FALSE, cellipse = 0)
```


# Neutral Intron SNPs

****

## Making files

**Make genind object**

`neutralintron.recode.vcf` contains neutral SNPs from intron regions for populations PVD, GB, BIS, and NAR. Steps for generating this VCF file are located in [EecSeq_Cvirginica_NeutralLoci.md](https://github.com/amyzyck/EecSeq_NB_EasternOyster/blob/master/Analysis/PopGen_SeaGen_Analyses/NonHaplotigMasked_Genome/Neutral_SNPs/EecSeq_Cvirginica_NeutralLoci.md).

Population NIN was removed from the VCF file following steps in [EecSeq_Cvirginica_OutlierDetection.md](https://github.com/amyzyck/EecSeq_NB_EasternOyster/blob/master/Analysis/EecSeq_Cvirginica_OutlierDetection.md).

`strata` contains population, environmental, and library information for each sample - can be accessed [here](https://github.com/amyzyck/EecSeq_NB_EasternOyster/blob/master/Analysis/PopGen_SeaGen_Analyses/strata).

```{r}
my_vcf_intron <- read.vcfR("neutralintron.recode.vcf")
strata <- read.table("strata", header=TRUE)
```

```{r}
rad.filt_intron <- vcfR2genind(my_vcf_intron, strata = strata, pop = c(rep("BIS", 10),rep("GB", 10),rep("NAR", 10), rep("PVD",10)))
```

```{r}
rad.filt_intron
```

```{r}
#Providing population names for plotting
pop_order <- c("BIS","GB","NAR","PVD")
```

**Read in the other info from .strata file and extract information such as locality, latitude, and longitude.**
```{r}
info <- as.data.frame(read.table("strata",header = T,sep = "\t",stringsAsFactors = F))
mystrats_intron <- as.data.frame(matrix(nrow = length(indNames(rad.filt_intron)),ncol=10))
colnames(mystrats_intron) <- c("Population","Latitude","Longitude","Distance","Temperature","SE","Salinity","pH","Chlorophylla","DO")
just.strats <- select(info,c("Population"))
stratted.filt_intron <- strata(rad.filt_intron, formula= Population, combine = TRUE,just.strats)
stratted.filt_intron@other <- select(info, Latitude,Longitude,Distance,SE,Temperature,Salinity,pH,Chlorophylla,DO)
```

```{r}
stratted.filt_intron
```

**Repeat for quasi-independent set of SNPs**

Steps for filtering are documented for OutFLANK in [EecSeq_Cvirginica_OutlierDetection.md](https://github.com/amyzyck/EecSeq_NB_EasternOyster/blob/master/Analysis/EecSeq_Cvirginica_OutlierDetection.md). However, I am repeating the steps here. 

```{r}
my_vcf_intron <- read.vcfR("neutralintron.recode.vcf")

geno_intron <- extract.gt(my_vcf_intron) # Character matrix containing the genotypes
position_intron <- getPOS(my_vcf_intron) # Positions in bp
chromosome_intron <- getCHROM(my_vcf_intron) # Chromosome information

G_intron <- matrix(NA, nrow = nrow(geno_intron), ncol = ncol(geno_intron))

G_intron[geno_intron %in% c("0/0", "0|0")] <- 0
G_intron[geno_intron  %in% c("0/1", "1/0", "1|0", "0|1")] <- 1
G_intron[geno_intron %in% c("1/1", "1|1")] <- 2

# NA should be replaced with “9” 
G_intron[is.na(G_intron)] <- 9
```


Chromosomes need to be of class integer for this to work.
```{r}
# Visualizing chromosomes
chrom_unique_intron <- unique(chromosome_intron)
print(chrom_unique_intron)
```

```{r}
# Removing "NC_" from the chromosome name so it can be converted to an integer
chrom_new_intron <- chromosome_intron%>%str_replace("NC_","")

# Converting the character string into an integer
chrom1_intron <- as.integer(chrom_new_intron)
```

`chromosome` and `position` need to be sorted for imputation to work.

```{r}
chrom_sort_intron <- sort(chrom1_intron)
pos_sort_intron <- sort(position_intron)
```

_Note: This filtering program does not allow for missing genotype values._


### Remove missing genotypes

```{r}
# removing missing data 
G_intron_miss <- matrix(NA, nrow = nrow(geno_intron), ncol = ncol(geno_intron))

G_intron_miss[geno_intron %in% c("0/0", "0|0")] <- 0
G_intron_miss[geno_intron  %in% c("0/1", "1/0", "1|0", "0|1")] <- 1
G_intron_miss[geno_intron %in% c("1/1", "1|1")] <- 2

# removing missing data
G_intron_miss[na.omit(G_intron_miss)]
```


```{r}
# Converting the genotype matrix to class FBM.code256 using the matrix where missing data was removed
G1_intron_miss <- add_code256(big_copy(t(G_intron_miss),type="raw"),code=bigsnpr:::CODE_012)
```

```{r}
newpc_intron_miss <- snp_autoSVD(G1_intron_miss,infos.chr =chrom_sort_intron,infos.pos = pos_sort_intron, roll.size = 32) # roll.size = Radius of rolling windows to smooth log-p-values
# maximum roll.size I could use is 32
which_pruned_intron_miss <- attr(newpc_intron_miss, which="subset") # Indexes of remaining SNPS after pruning
length(which_pruned_intron_miss)
```


```{r}
invisible(lapply(which_pruned_intron_miss, write, "pruned_data_intron.txt", append=TRUE))
```


**In terminal**

```
$ mawk '!/#/' neutralintron.recode.vcf | cut -f1,2 > totallociintron
$ NUM=(`cat totallociintron | wc -l`)
$ paste <(seq 1 $NUM) totallociintron > lociintron.plus.index
$ cat pruned_data_intron.txt | parallel "grep -w ^{} lociintron.plus.index" | cut -f2,3> pruned_data_intron.loci.txt

$ head pruned_data_intron.loci.txt

output:
NC_035780.1	574321
NC_035780.1	574444
NC_035780.1	574657
NC_035780.1	612658
NC_035780.1	618706
NC_035780.1	722826
NC_035780.1	771520
NC_035780.1	771800
NC_035780.1	771812
NC_035780.1	780771
```

Create VCF file with just the pruned_data loci 
```
$ vcftools --vcf neutralintron.recode.vcf --recode --recode-INFO-all --positions pruned_data_intron.loci.txt --out pruned_data_intron

output:
After filtering, kept 40 out of 40 Individuals
Outputting VCF file...
After filtering, kept 3766 out of a possible 16701 Sites
Run Time = 1.00 seconds
```

```{r}
my_vcf_u_intron <- read.vcfR("pruned_data_intron.recode.vcf")
strata <- read.table("strata", header=TRUE)
```

```{r}
rad.u_intron <- vcfR2genind(my_vcf_u_intron, strata = strata, pop = c(rep("BIS", 10),rep("GB", 10),rep("NAR", 10), rep("PVD",10)))
```

```{r}
rad.u_intron
```


```{r}
stratted.u_intron <- strata(rad.u_intron, formula= Population, combine = TRUE,just.strats)
stratted.u_intron@other <- select(info, Latitude,Longitude, Distance, SE,  Temperature,Salinity,pH,Chlorophylla,DO)
```

```{r}
stratted.u_intron
```

**Make hierfstat object**
```{r}
hf.filt_intron <- genind2hierfstat(rad.filt_intron, pop = c(rep("BIS", 10),rep("GB", 10),rep("NAR", 10), rep("PVD",10))) 
```

```{r}
hf.u_intron <- genind2hierfstat(rad.u_intron, pop = c(rep("BIS", 10),rep("GB", 10),rep("NAR", 10), rep("PVD",10)))
hf.u_intron <- hf.u_intron$hierfstat.no.imputation
```

## Estimating effective migration surfaces [(EEMS)](https://github.com/dipetkov/eems)

__1. datapath.diffs__

**Steps for generating `datapath.diffs` are completed in RStudio.**
```{r}
# V1 methtod to get diffs matrix, preferred
bed2diffs_v1 <- function(Geno) {
  nIndiv <- nrow(Geno)
  nSites <- ncol(Geno)
  Diffs <- matrix(0, nIndiv, nIndiv)
  
  for (i in seq(nIndiv - 1)) {
    for (j in seq(i + 1, nIndiv)) {
      x <- Geno[i, ]
      y <- Geno[j, ]
      Diffs[i, j] <- mean((x - y)^2, na.rm = TRUE)
      Diffs[j, i] <- Diffs[i, j]
    }
  }
  Diffs
}
```

```{r}
# V2 method to get .diffs matrix, only if V1 doesn't work
bed2diffs_v2 <- function(Geno) {
  nIndiv <- nrow(Geno)
  nSites <- ncol(Geno)
  Miss <- is.na(Geno)
  ## Impute NAs with the column means (= twice the allele frequencies)
  Mean <- matrix(colMeans(Geno, na.rm = TRUE), ## a row of means
                 nrow = nIndiv, ncol = nSites, byrow = TRUE) ## a matrix with nIndiv identical rows of means
  Mean[Miss == 0] <- 0 ## Set the means that correspond to observed genotypes to 0
  Geno[Miss == 1] <- 0 ## Set the missing genotypes to 0 (used to be NA) 
  Geno <- Geno + Mean
  ## Compute similarities
  Sim <- Geno %*% t(Geno) / nSites
  SelfSim <- diag(Sim) ## self-similarities
  vector1s <- rep(1, nIndiv) ## vector of 1s
  ## This chunk generates a `diffs` matrix
  Diffs <- SelfSim %*% t(vector1s) + vector1s %*% t(SelfSim) - 2 * Sim
  Diffs
}
```

```{r}
geno_intron <- stratted.filt_intron@tab
```

```{r}
# Get rid of non-biallelic loci
multi.loci_intron <- names(which(stratted.filt_intron@loc.n.all != 2))
multi.cols_intron <- which(grepl(paste0("^", multi.loci_intron, "\\.\\d+$", collapse = "|"), colnames(geno_intron)))
if (length(multi.cols_intron)) geno_intron <- geno_intron[, - multi.cols_intron]
nloci_intron <- dim(geno_intron)[2] / 2
dim(geno_intron)
```

```{r}
stopifnot(identical(stratted.filt_intron@type, 'codom'))
```

```{r}
# bed2diffs functions  
diffs.v1_intron <- bed2diffs_v1(geno_intron)
diffs.v2_intron <- bed2diffs_v2(geno_intron)
# Round to 6 digits
diffs.v1_intron <- round(diffs.v1_intron, digits = 6)
diffs.v2_intron <- round(diffs.v2_intron, digits = 6)
```

**Check that the dissimilarity matrix has one positive eigenvalue and nIndiv-1 negative eigenvalues, as required by a full-rank Euclidean distance matrix. If the V1 method does not make a Euclidean matrix, you must use V2.**

```{r}
tail(sort(round(eigen(diffs.v1_intron)$values, digits = 2)))
```

```{r}
tail(sort(round(eigen(diffs.v2_intron)$values, digits = 2)))
```

```{r}
# Set suffix for EEMS input files
suf_intron <- "neutralintrondata-filt"
```

```{r}
# This saves the file to directory 
write.table(diffs.v1_intron, paste(suf_intron,".v1.diffs",sep=""), 
            col.names = FALSE, row.names = FALSE, quote = FALSE)
```


__2. datapath.coord__

`datapath.coord` are the sample coordinates, two coordinates per sample, one sample per line. The sampling locations should be given in the same order as the rows and columns of the dissimilarity matrix.

**Steps for generating `datapath.coord` are completed in RStudio.**
```{r}
## Get gps coordinates from previously created info matrix
xOR.info <- dplyr::filter(info)
gps_matrix <- select(xOR.info,c("Longitude","Latitude"))
```

```{r}
#write .coord file
write.table(gps_matrix, paste(suf_intron,".v1.coord",sep=""),col.names = FALSE, row.names = FALSE,quote = FALSE)
```

__3. datapath.outer__

`datapath.outer` are the habitat coordinates, as a sequence of vertices that form a closed polygon. The habitat vertices should be listed counterclockwise and the first vertex should also be the last vertex, so that the outline is a closed ring. Otherwise, EEMS attempts to "correct" the polygon and prints a warning message.

**`datapath.outer` is created manually in Excel,based on site coordinates gathered from Google Maps, copied into terminal using `nano`, and saved as the file with the appropriate extension.** 


**`runeems_snps` is then run in command-line following the steps documented in [NB_EEMS_OutlierHap.md](https://github.com/amyzyck/EecSeq_NB_EasternOyster/blob/master/Analysis/PopGen_SeaGen_Analyses/EEMS/NB_EEMS_OutlierHap.md) 

**Back in RStudio to plot `runeems_snps` outputs**

```{r}
# Install rEEMSplots
library(rEEMSplots)
```

```{r} 
# Plotting EEMS after running runeems_snps
path = "./NB_EEMS_Neutral/"
dirs = c(paste0(path,"neutralintrondata-D200-chain1"), paste0(path,"neutralintrondata-D300-chain1"), paste0(path,"neutralintrondata-D600-chain1"))
```

```{r}
eems.plots(mcmcpath = c(paste0(path,"./neutralintrondata-D200-chain1"), paste0(path,"neutralintrondata-D300-chain1"), paste0(path,"neutralintrondata-D600-chain1")), plotpath = paste0(path,"neutralintrondata-All-plots"),
           longlat = T,add.grid=F,add.outline = T,add.demes = T,
           projection.in = "+proj=longlat +datum=WGS84",projection.out = "+proj=merc +datum=WGS84",
           add.map = T,add.abline = T, add.r.squared = T)
```

## Pairwise Fst

```{r}
fst.mat_intron <- pairwise.WCfst(hf.filt_intron)
```

```{r}
gindF.fst.mat.triN_intron <- as.matrix(fst.mat_intron)
colnames(gindF.fst.mat.triN_intron) <- pop_order
rownames(gindF.fst.mat.triN_intron) <- pop_order
```

```{r}
meltedN_intron <- melt(gindF.fst.mat.triN_intron, na.rm =TRUE)
round(gindF.fst.mat.triN_intron,4)
```

```{r}
summary(meltedN_intron$value)
```


```{r}
#Plotting Pairwise fst
neutral_intron <- ggplot(data = meltedN_intron, aes(Var2, Var1, fill = value))+ geom_tile(color = "white")+ 
  scale_fill_gradient(low = "white", high = "red", name="FST")  +
  ggtitle(expression(atop("Pairwise FST, WC (1984) Neutral Intron SNPs", atop(italic("N = 40, L = 16,701"), ""))))+
  labs( x = "Sampling Site", y = "Sampling Site") + 
  theme(axis.text.x = element_text(angle = 45, vjust = 1, size = 12, hjust = 1),axis.text.y = element_text(size = 13)) + 
  theme(axis.title = element_text(size = 16),legend.text = element_text(size =15), legend.title = element_text(size =16)) +
  theme(plot.title = element_text(size = 17)) +
coord_fixed()
neutral_intron
```

## Genetic diversity (observed and expected heterozygosity)

```{r}
comb_intron <- summary(stratted.filt_intron)
names(comb_intron)
```

```{r}
plot(comb_intron$Hobs, xlab="Loci number", ylab="Observed Heterozygosity", 
     main="Observed heterozygosity per locus")
```


```{r}
plot(comb_intron$Hobs,comb_intron$Hexp, xlab="Hobs", ylab="Hexp", 
     main="Expected heterozygosity as a function of observed heterozygosity per locus")
```

```{r}
bartlett.test(list(comb_intron$Hexp, comb_intron$Hobs)) # a test : H0: Hexp = Hobs
```

**_Significant difference between Observed and expected heterzygosity_**.  

```{r}
basicstat_intron <- basic.stats(hf.filt_intron, diploid = TRUE, digits = 3)
```

```{r}
as.data.frame(basicstat_intron$overall)
```

```{r}
# get bootstrap confidence values for Fis
boot_intron <- boot.ppfis(hf.filt_intron,nboot = 1000)
boot5_intron <- boot.ppfis(hf.filt_intron,nboot = 1000,quant = 0.5)
```

```{r}
# combine all pop statistics
colnames(basicstat_intron$Ho) <- pop_order
Ho_intron <- colMeans(basicstat_intron$Ho,na.rm = T)
He_intron <- colMeans(basicstat_intron$Hs,na.rm = T)
Fis_intron <- boot5_intron$fis.ci$ll
y_intron <- cbind(pop_order,Ho_intron,He_intron,Fis_intron,boot_intron$fis.ci,latitude, longitude, distance, sewage, temperature, salinity, pH, Chlor_a, DO)
y_intron
```

```{r}
summary(He_intron)
```

```{r}
summary(Fis_intron)
```

```{r}
# Plot He vs Latitude
R2_intron = round(summary(lm(y_intron$He_intron ~ y_intron$latitude))$r.squared, 4)             
ggplot(y_intron, aes(x = latitude, y = He_intron)) + geom_point(aes(shape=pop_order, color=pop_order), size = 3) + 
  scale_shape_manual(values = c(15,18,17,16)) +
geom_smooth(method=lm) + 
ggtitle("Expected heterozygosity vs Latitude, Neutral Intron SNPs") +
annotate(geom = "text", label=paste("italic(R^2)==",R2_intron), x=41.55, y=0.2635, parse=T) +
scale_x_continuous()
```

```{r}
# Plot Ho vs Latitude
R2_intron = round(summary(lm(y_intron$Ho_intron ~ y_intron$latitude))$r.squared, 4)             
ggplot(y_intron, aes(x = latitude, y = Ho_intron)) + geom_point(aes(shape=pop_order, color=pop_order), size = 3) + 
  scale_shape_manual(values = c(15,18,17,16)) + 
geom_smooth(method=lm) + 
ggtitle("Observed heterozygosity vs Latitude, Neutral Intron SNPs") +
annotate(geom = "text", label=paste("italic(R^2)==",R2_intron), x=41.55, y=0.26, parse=T) +
scale_x_continuous()
```

```{r}
#Plot He vs Longitude
R3_intron = round(summary(lm(y_intron$He_intron ~ y_intron$longitude))$r.squared, 4)             
ggplot(y_intron, aes(x = longitude, y = He_intron)) + geom_point(aes(shape=pop_order, color=pop_order), size = 3) + 
  scale_shape_manual(values = c(15,18,17,16)) +
geom_smooth(method=lm) + 
ggtitle("Expected heterozygosity vs Longitude, Neutral Intron SNPs") +
annotate(geom = "text", label=paste("italic(R^2)==",R3_intron), x=-71.44, y=0.2635, parse=T) +
scale_x_reverse()
```

```{r}
#Plot Ho vs Longitude
R3_intron = round(summary(lm(y_intron$Ho_intron ~ y_intron$longitude))$r.squared, 4)             
ggplot(y_intron, aes(x = longitude, y = Ho_intron)) + geom_point(aes(shape=pop_order, color=pop_order), size = 3) + 
  scale_shape_manual(values = c(15,18,17,16)) +
geom_smooth(method=lm) + 
ggtitle("Observed heterozygosity vs Longitude, Neutral Intron SNPs") +
annotate(geom = "text", label=paste("italic(R^2)==",R3_intron), x=-71.44, y=0.26, parse=T) +
scale_x_reverse()
```

```{r}
#Plot He vs Distance from sewage outflow 
R4_intron = round(summary(lm(y_intron$He_intron ~ y_intron$distance))$r.squared, 4)             
ggplot(y_intron, aes(x = distance, y = He_intron)) + geom_point(aes(shape=pop_order, color=pop_order), size = 3) + 
  scale_shape_manual(values = c(15,18,17,16)) + 
geom_smooth(method=lm) + 
ggtitle("Expected heterozygosity vs Distance, Neutral Intron SNPs") +
annotate(geom = "text", label=paste("italic(R^2)==",R4_intron), x=2, y=0.265, parse=T) +
scale_x_reverse()
```

```{r}
#Plot Ho vs Distance from sewage outflow 
R4_intron = round(summary(lm(y_intron$Ho_intron ~ y_intron$distance))$r.squared, 4)             
ggplot(y_intron, aes(x = distance, y = Ho_intron)) + geom_point(aes(shape=pop_order, color=pop_order), size = 3) + 
  scale_shape_manual(values = c(15,18,17,16)) + 
geom_smooth(method=lm) + 
ggtitle("Observed heterozygosity vs Distance, Neutral Intron SNPs") +
annotate(geom = "text", label=paste("italic(R^2)==",R4_intron), x=2, y=0.255, parse=T) +
scale_x_reverse()
```

```{r}
#Plot He vs Sewage effluent  
R_intron = round(summary(lm(y_intron$He_intron ~ y_intron$sewage))$r.squared, 4)             
ggplot(y_intron, aes(x = sewage, y = He_intron)) + geom_point(aes(shape=pop_order, color=pop_order), size = 3) + 
  scale_shape_manual(values = c(15,18,17,16)) + 
geom_smooth(method=lm) + 
ggtitle("Expected heterozygosity vs Sewage Effluent, Neutral Intron SNPs") +
annotate(geom = "text", label=paste("italic(R^2)==",R_intron), x=7, y=0.263, parse=T) +
scale_x_reverse()
```

```{r}
#Plot Ho vs Sewage effluent  
R_intron = round(summary(lm(y_intron$Ho_intron ~ y_intron$sewage))$r.squared, 4)             
ggplot(y_intron, aes(x = sewage, y = Ho_intron)) + geom_point(aes(shape=pop_order, color=pop_order), size = 3) + 
  scale_shape_manual(values = c(15,18,17,16)) + 
geom_smooth(method=lm) + 
ggtitle("Observed heterozygosity vs Sewage Effluent, Neutral Intron SNPs") +
annotate(geom = "text", label=paste("italic(R^2)==",R_intron), x=7, y=0.26, parse=T) +
scale_x_reverse()
```

```{r}
#Plot He vs Temperature 
R5_intron = round(summary(lm(y_intron$He_intron ~ y_intron$temperature))$r.squared, 4)             
ggplot(y_intron, aes(x = temperature, y = He_intron)) + geom_point(aes(shape=pop_order, color=pop_order), size = 3) + 
  scale_shape_manual(values = c(15,18,17,16)) +
geom_smooth(method=lm) + 
ggtitle("Expected heterozygosity vs Temperature, Neutral Intron SNPs") +
annotate(geom = "text", label=paste("italic(R^2)==",R5_intron), x=23.25, y=0.2635, parse=T) +
scale_x_reverse()
```

```{r}
#Plot Ho vs Temperature 
R5_intron = round(summary(lm(y_intron$Ho_intron ~ y_intron$temperature))$r.squared, 4)             
ggplot(y_intron, aes(x = temperature, y = Ho_intron)) + geom_point(aes(shape=pop_order, color=pop_order), size = 3) + 
  scale_shape_manual(values = c(15,18,17,16)) + 
geom_smooth(method=lm) + 
ggtitle("Observed heterozygosity vs Temperature, Neutral Intron SNPs") +
annotate(geom = "text", label=paste("italic(R^2)==",R5_intron), x=23.25, y=0.26, parse=T) +
scale_x_reverse()
```

```{r}
#Plot He vs Salinity 
R6_intron = round(summary(lm(y_intron$He_intron ~ y_intron$salinity))$r.squared, 4)             
ggplot(y_intron, aes(x = salinity, y = He_intron)) + geom_point(aes(shape=pop_order, color=pop_order), size = 3) + 
  scale_shape_manual(values = c(15,18,17,16)) + 
geom_smooth(method=lm) + 
ggtitle("Expected heterozygosity vs Salinity, Neutral Intron SNPs") +
annotate(geom = "text", label=paste("italic(R^2)==",R6_intron), x=23.25, y=0.267, parse=T) +
scale_x_reverse()
```

```{r}
#Plot Ho vs Salinity 
R6_intron = round(summary(lm(y_intron$Ho_intron ~ y_intron$salinity))$r.squared, 4)             
ggplot(y_intron, aes(x = salinity, y = Ho_intron)) + geom_point(aes(shape=pop_order, color=pop_order), size = 3) + 
  scale_shape_manual(values = c(15,18,17,16)) + 
geom_smooth(method=lm) + 
ggtitle("Observed heterozygosity vs Salinity, Neutral Intron SNPs") +
annotate(geom = "text", label=paste("italic(R^2)==",R6_intron), x=23.25, y=0.255, parse=T) +
scale_x_reverse()
```

```{r}
#Plot He vs pH
R7_intron = round(summary(lm(y_intron$He_intron ~ y_intron$pH))$r.squared, 4)             
ggplot(y_intron, aes(x = pH, y = He_intron)) + geom_point(aes(shape=pop_order, color=pop_order), size = 3) + 
  scale_shape_manual(values = c(15,18,17,16)) +
geom_smooth(method=lm) + 
ggtitle("Expected heterozygosity vs pH, Neutral Intron SNPs") +
annotate(geom = "text", label=paste("italic(R^2)==",R7_intron), x=7.45, y=0.2635, parse=T) +
scale_x_reverse()
```

```{r}
#Plot Ho vs pH
R7_intron = round(summary(lm(y_intron$Ho_intron ~ y_intron$pH))$r.squared, 4)             
ggplot(y_intron, aes(x = pH, y = Ho_intron)) + geom_point(aes(shape=pop_order, color=pop_order), size = 3) + 
  scale_shape_manual(values = c(15,18,17,16)) + 
geom_smooth(method=lm) + 
ggtitle("Observed heterozygosity vs pH, Neutral Intron SNPs") +
annotate(geom = "text", label=paste("italic(R^2)==",R7_intron), x=7.45, y=0.255, parse=T) +
scale_x_reverse()
```

```{r}
#Plot He vs Chlorophyll a 
R8_intron = round(summary(lm(y_intron$He_intron ~ y_intron$Chlor_a))$r.squared, 4)             
ggplot(y_intron, aes(x = Chlor_a, y = He_intron)) + geom_point(aes(shape=pop_order, color=pop_order), size = 3) + 
  scale_shape_manual(values = c(15,18,17,16)) + 
geom_smooth(method=lm) + 
ggtitle("Expected heterozygosity vs Chlorophyll a, Neutral Intron SNPs") +
annotate(geom = "text", label=paste("italic(R^2)==",R8_intron), x=8, y=0.263, parse=T) +
scale_x_reverse()
```

```{r}
#Plot Ho vs Chlorophyll a 
R8_intron = round(summary(lm(y_intron$Ho_intron ~ y_intron$Chlor_a))$r.squared, 4)             
ggplot(y_intron, aes(x = Chlor_a, y = Ho_intron)) + geom_point(aes(shape=pop_order, color=pop_order), size = 3) + 
  scale_shape_manual(values = c(15,18,17,16)) + 
geom_smooth(method=lm) + 
ggtitle("Observed heterozygosity vs Chlorophyll a, Neutral Intron SNPs") +
annotate(geom = "text", label=paste("italic(R^2)==",R8_intron), x=6, y=0.26, parse=T) +
scale_x_reverse()
```

```{r}
#Plot He vs Dissolved Oxygen 
R9_intron = round(summary(lm(y_intron$He_intron ~ y_intron$DO))$r.squared, 4)             
ggplot(y_intron, aes(x = DO, y = He_intron)) + geom_point(aes(shape=pop_order, color=pop_order), size = 3) + 
  scale_shape_manual(values = c(15,18,17,16)) + 
geom_smooth(method=lm) + 
ggtitle("Expected heterozygosity vs Dissolved Oxygen, Neutral Intron SNPs") +
annotate(geom = "text", label=paste("italic(R^2)==",R9_intron), x=6, y=0.265, parse=T) +
scale_x_reverse()
```

```{r}
#Plot Ho vs Dissolved Oxygen 
R9_intron = round(summary(lm(y_intron$Ho_intron ~ y_intron$DO))$r.squared, 4)             
ggplot(y_intron, aes(x = DO, y = Ho_intron)) + geom_point(aes(shape=pop_order, color=pop_order), size = 3) + 
  scale_shape_manual(values = c(15,18,17,16)) + 
geom_smooth(method=lm) + 
ggtitle("Observed heterozygosity vs Dissolved Oxygen, Neutral Intron SNPs") +
annotate(geom = "text", label=paste("italic(R^2)==",R9_intron), x=6, y=0.26, parse=T) +
scale_x_reverse()
```


## PCA 

```{r}
X_intron <- tab(stratted.u_intron, freq = TRUE, NA.method = "mean")
pca1_intron <- dudi.pca(X_intron, scale = FALSE, scannf = FALSE, nf = 3)
barplot(pca1_intron$eig[1:50], main = "PCA eigenvalues", col = heat.colors(50))
s.class(pca1_intron$li, pop(stratted.filt_intron))
title("PCA of neutral intron SNPs\naxes 1-2")
add.scatter.eig(pca1_intron$eig[1:20], 3,1,2)
col <- funky(15) 
s.class(pca1_intron$li, pop(stratted.filt_intron),xax=1,yax=2, col=col, axesell=FALSE, cstar=0, cpoint=3, grid=FALSE)
```


## Discriminant Analysis of Principal Components (DAPC)

The first line of the code block below is interactive. In the console, it will tell you:  
1. Choose the number PCs to retain (>=1): 
    I picked 1 
2. Choose the number discriminant functions to retain (>=1):
    I picked 2

```{r}
grp_intron <- find.clusters(stratted.u_intron, n.pca = 1, choose.n.clust = FALSE)
table(pop(stratted.u_intron), grp_intron$grp)
table.value(table(pop(stratted.u_intron), grp_intron$grp), col.lab=paste("inf", 1:2), row.lab=paste("ori", 1:4))
```

Again, the first line of the code block below is interactive. In the console, it will tell you:  
1. Choose the number PCs to retain (>=1): 
    I picked 1 
2. Choose the number discriminant functions to retain (>=1):
    I picked 1

```{r}
dapc1_intron <- dapc(stratted.u_intron, grp_intron$grp, n.pca = 1, n.da= 1)
scatter(dapc1_intron,col=col,bg="white", solid=1)
```

Again, the first line of the code block below is interactive. In the console, it will tell you:  
1. Choose the number PCs to retain (>=1): 
    I picked 1 
2. Choose the number discriminant functions to retain (>=1):
    I picked 1

```{r}
contrib_intron <- loadingplot(dapc1_intron$var.contr, axis=1, thres=.01, lab.jitter=1)
contrib_intron
setPop(rad.filt_intron) <- ~Library
dapc1_intron <- dapc(stratted.u_intron, pop(stratted.u_intron), n.pca = 1, n.da = 1)
contrib_intron <- loadingplot(dapc1_intron$var.contr, axis=1, thres=.05, lab.jitter=1)
```

```{r}
#Structure Like
compoplot(dapc1_intron, posi="bottomright",txt.leg=paste("Cluster", 1:4), lab="", ncol=1, xlab="individuals")
temp_intron <- which(apply(dapc1_intron$posterior,1, function(e) all(e<0.9)))
compoplot(dapc1_intron, subset=temp_intron, posi="bottomright", txt.leg=paste("Cluster", 1:4), ncol=2)
```

### PCAviz

```{r}
NA.afDraw<- function(ind){
  ind.mat <- ind@tab
  new.mat <- ind.mat
  af = colSums(ind.mat[,seq(1,ncol(ind.mat)-1,2)],na.rm = TRUE)/
      (2*apply(ind.mat[,seq(1,ncol(ind.mat)-1,2)],2,function(x) sum(!is.na(x))))
  af.Draw <- function(geno, af){
     new <- function(geno,af){
        if(is.na(geno)){
        newA = rbinom(1,2,af)
        }
        else {newA <- geno}
        return(newA)
   }
  new.row <- mapply(geno,af,FUN = new)
  return(new.row)}
  
  new.mat[,seq(1,ncol(ind.mat)-1,2)] <- t(apply(ind.mat[,seq(1,ncol(ind.mat)-1,2)],1,af.Draw,af))
  new.mat[,seq(2,ncol(ind.mat),2)] <- 2-new.mat[,seq(1,ncol(ind.mat)-1,2)]
  new.ind <- ind
  new.ind@tab <- new.mat
  return(new.ind)
}
```

```{r}
u.na_intron <- NA.afDraw(stratted.u_intron)
```

```{r}
pca_intron <- dudi.pca(u.na_intron,center=TRUE,scale=TRUE,scannf = F, nf = 30)
```

```{r}
col18 <- funky(length(unique(u.na_intron@strata$Population)))
#Colors that match the neutral Structure results
col6 <-  c("firebrick","dodgerblue","midnightblue","salmon")
par(mfrow=c(2,2))
s.class(pca_intron$li, strata(u.na_intron)$Population,xax=1,yax=2, 
        sub = "neutralintronloci,PC 1-2, 40 individuals, 3,766 loci, maf 5.0%",
        possub = "topleft",col=transp(col6,.6),axesell=FALSE,
        cstar=0, cpoint=3, grid=FALSE, cellipse = 0)
s.class(pca_intron$li, strata(u.na_intron)$Population,xax=1,yax=3, 
        sub = "neutralintronloci,PC 1-3, 40 individuals, 3,766 loci, maf 5.0%",
        possub = "topleft",col=transp(col6,.6),axesell=FALSE,
        cstar=0, cpoint=3, grid=FALSE, cellipse = 0)
s.class(pca_intron$li, strata(u.na_intron)$Population,xax=1,yax=4, 
        sub = "neutralintronloci,PC 1-4, 40 individuals, 3,766 loci, maf 5.0%",
        possub = "topleft",col=transp(col6,.6),axesell=FALSE,
        cstar=0, cpoint=3, grid=FALSE, cellipse = 0)
s.class(pca_intron$li, strata(u.na_intron)$Population,xax=1,yax=5, 
        sub = "neutralintronloci,PC 1-5, 40 individuals, 3,766 loci, maf 5.0%",
        possub = "topleft",col=transp(col6,.6),axesell=FALSE,
        cstar=0, cpoint=3, grid=FALSE, cellipse = 0)
```


# Seascape Redundancy Analysis

This code follows that documented by [Tom Jenkins](https://github.com/Tom-Jenkins/seascape_rda_tutorial).

## [Prepare genetic data for redundancy analysis](https://github.com/Tom-Jenkins/seascape_rda_tutorial/blob/master/1.Prepare_genetic_data/1.prepare_genetic_data.R).

#### Notes before execution:
1. Make sure all required R packages are installed.
2. Set working directory to the location of this R script.

```{r}
# Load packages
library(adegenet)
library(poppr)
library(dplyr)
library(reshape2)
library(ggplot2)
library(vcfR)
```

```{r}
# Explore data
rad.filt
nLoc(rad.filt) # number of loci
nPop(rad.filt) # number of sites
nInd(rad.filt) # number of individuals
summary(rad.filt$pop) # sample size
```

```{r}
# Calculate allele frequencies for each site
allele_freqs = data.frame(rraf(rad.filt, by_pop = TRUE, correction = FALSE), check.names = FALSE)
```

```{r}
# Keep only the first of the two alleles for each SNP (since p=1-q).
allele_freqs = allele_freqs[, seq(1, dim(allele_freqs)[2], 2)]
```

```{r}
# Export allele frequencies
write.csv(allele_freqs, file = "all_allele_freqs.csv", row.names = TRUE)
```

#--------------#
#
# Calculate minor allele frequencies
#
#--------------#

```{r}
# Separate genind object by site
site_list = seppop(rad.filt)
names(site_list)
```

```{r}
# Calculate the minor allele frequency for each site
maf_list = lapply(site_list, FUN = minorAllele)
```

```{r}
# Convert list to dataframe
maf = as.data.frame(maf_list) %>% t() %>% as.data.frame()
```

```{r}
# Export minor allele frequencies
write.csv(maf, file = "minor_allele_freqs.csv", row.names = TRUE)
```

## [Prepare environmental data for redundancy analysis](https://github.com/Tom-Jenkins/seascape_rda_tutorial/blob/master/3.Prepare_environmental_data/3.prepare_env_data.R).

#### Environmental variables:
- Distance from sewage effluent source (km)
- Sewage Effluent (PW stats)
- Mean temperature (deg C)
- Mean Salinity (psu)
- Mean pH
- Mean Chlorophyll-a (ug/L)
- Mean Dissolved Oxygen (mg/L)

Environmental data for each population is saved in a `strata_pop` file that can be accessed [here](https://github.com/amyzyck/EecSeq_NB_EasternOyster/blob/master/Analysis/PopGen_SeaGen_Analyses/strata_pop)

```{r}
# All environmental data was previously saved in strata file 
strata_pop <- read.table("strata_pop", header=TRUE)
strata_pop
```

```{r}
# Export data as a csv file
write.csv(strata_pop, file="environmental_data.csv", row.names = FALSE)
```

I also prepared spatial data for the redundancy analysis which is documented [here](https://github.com/amyzyck/EecSeq_NB_EasternOyster/blob/master/Analysis/PopGen_SeaGen_Analyses/RedundancyAnalysis/PrepSpatialData.Rmd).

Allele frequency, environmental, and spatial csv files are saved to your working directory and must be imported into the Rscript to run the redundancy analysis. Documentation of the RDA can be accessed [here](https://github.com/amyzyck/EecSeq_NB_EasternOyster/blob/master/Analysis/PopGen_SeaGen_Analyses/RedundancyAnalysis/RDA_Outlier_Hap.Rmd). 

RDA and pRDA output plots can be viewed [here](https://github.com/amyzyck/EecSeq_NB_EasternOyster/tree/master/Output/RedundancyAnalysis/NonHapMasked_Neutral).