---
title: "Redundancy Analysis - Outlier SNPs Hap"
author: "Amy Zyck"
date: "2/23/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Redundancy Analysis

This code follows that documented by [Tom Jenkins](https://github.com/Tom-Jenkins/seascape_rda_tutorial).

Steps for completing the redundancy analysis for Outlier SNPs. Steps for preparing the allele frequency data and environmental data can be found [here](https://github.com/amyzyck/EecSeq_NB_EasternOyster/blob/master/Analysis/PopGen_SeaGen_Analyses/HaplotigMasked_Genome/Outlier_SNPs/NB_PopGen_OutlierHap.Rmd), and steps for preparing spatial data can be found [here](https://github.com/amyzyck/EecSeq_NB_EasternOyster/blob/master/Analysis/PopGen_SeaGen_Analyses/HaplotigMasked_Genome/Outlier_SNPs/RedundancyAnalysis/PrepSpatialData.Rmd). 

```{r}
# Load packages
library(psych)
library(dplyr)
library(adespatial)
library(vegan)
```

```{r}
# Import genetic data
allele_freqs_outlier = read.csv("all_allele_freqs.csv", row.names = 1, check.names = FALSE)
allele_freqs_outlier_BayEnv = read.csv("all_allele_freqs_BayEnv.csv", row.names = 1, check.names = FALSE)
```

```{r}
# Import environmental data
env.raw = read.csv("environmental_data.csv", row.names = 1)
```

```{r}
# Import spatial data
dbmem.raw = read.csv("dbmems.csv")
```

```{r}
# Set seed
set.seed(123)
```


#--------------#
#
# Multicollinearity checks
#
#--------------#

```{r}
# Plot and run correlation test on environmental variables
pairs.panels(env.raw, scale = TRUE)
```

```{r}
# Remove correlated variables
env.data = subset(env.raw, select = -c(Latitude,Longitude))
pairs.panels(env.data, scale = TRUE)
```

For now, continue without removing correlated variables - replace env.data with env.raw

#--------------#
#
# Identify significant variables
#
#--------------#

```{r}
# Use forward selection to identify significant environmental variables
Y = allele_freqs_outlier
X = env.raw
env.for = forward.sel(Y, X, alpha = TRUE)
env.for
```

Only one dbMEM 

```{r}
# Use forward selection to identify significant dbmems
dbmem.for = forward.sel(Y = allele_freqs_outlier, X = dbmem.raw, alpha = TRUE)
dbmem.for
```

```{r}
# Subset only significant independent variables to include in the RDA
env.sig = subset(env.data, select = env.for$variables)
str(env.sig)
dbmem.sig = subset(dbmem.raw, select = dbmem.for$variables)
str(dbmem.sig)
```

```{r}
# Combine environmental variables and dbmems
env.dbmems = cbind(env.sig, dbmem.sig)
str(env.dbmems)
```


#--------------#
#
# Redundancy analysis
#
#--------------#

```{r}
# Perform RDA with all variables
rda1 = rda(allele_freqs_outlier ~ ., data = env.sig, scale = FALSE)
rda1
```

```{r}
# Model summaries
RsquareAdj(rda1) # adjusted Rsquared 
vif.cca(rda1) # variance inflation factor (<10 OK)
anova.cca(rda1, permutations = 1000) # full model
anova.cca(rda1, permutations = 1000, by="margin") # per variable 
```

```{r}
# Variance explained by each canonical axis
summary(eigenvals(rda1, model = "constrained"))
screeplot(rda1)
```

```{r}
# Create a dataframe to correctly colour regions
col_dframe = data.frame("site" = rownames(allele_freqs_outlier))
```

```{r}
# Function to add regional labels to dataframe
addregion = function(x){
  # If pop label is present function will output the region
  if(x=="BIS") y = " Bissel Cove (BIS) "
  if(x=="GB") y = " Greenwich Bay (GB) "
  if(x=="NAR") y = " Narrow River (NAR) "
  if(x=="PVD") y = " Providence (PVD) "
  return(y)
}
```

```{r}
# Add regional labels
col_dframe$region = sapply(col_dframe$site, addregion)
```

```{r}
# Add factor levels
region_order = c(" Bissel Cove (BIS) "," Greenwich Bay (GB) "," Narrow River (NAR) ", " Providence (PVD) ")
col_dframe$region = factor(col_dframe$region, levels = region_order)
```

```{r}
# Create colour scheme
# blue=#377EB8, green=#7FC97F, orange=#FDB462, red=#E31A1C
cols = c("#7FC97F","#377EB8","#FDB462","#E31A1C")
```

```{r}
# Visualise results of RDA
png("rda.png", width = 8, height = 7, units = "in", res = 600)
plot(rda1, type="n", scaling = 3)
title("Seascape redundancy analysis - Outlier SNPs")
# SITES
points(rda1, display="sites", pch=21, scaling=3, cex=2, col="black",
       bg=cols[col_dframe$region]) # sites
# text(rda1, display="sites", scaling = 3, col="black", font=2, pos=4)
# PREDICTORS
text(rda1, display="bp", scaling=3, col="red1", cex=1.4, lwd=2)
# SNPS
# text(rda1, display="species", scaling = 3, col="blue", cex=0.7, pos=4) # SNPs
# LEGEND
legend("topright", legend=levels(col_dframe$region), bty="n", col="black",
       pch=21, cex=1.3, pt.bg=cols)
# OTHER LABELS
adj.R2 = round(RsquareAdj(rda1)$adj.r.squared, 3)
mtext(bquote(italic("Adjusted R")^"2"~"= "~.(adj.R2)), side = 3, adj = 0.5)
dev.off()
```

#--------------#
#
# Partial redundancy analysis
#
#--------------#

```{r}
# Perform RDA while controlling for geographical location
pRDA = rda(allele_freqs_outlier_BayEnv ~ Salinity + DO + Condition(MEM1),
           data = env.dbmems, scale = FALSE)
pRDA
RsquareAdj(pRDA) # adjusted Rsquared 
vif.cca(pRDA) # variance inflation factor (<10 OK)
anova.cca(pRDA, permutations = 1000) # full model
anova.cca(pRDA, permutations = 1000, by = "margin") # per variable
```

```{r}
# Visualise results of RDA
png("partial_rda.png", width = 9, height = 7, units = "in", res = 600)
plot(pRDA, type="n", scaling = 3)
title("pRDA - Outlier SNPs")
# SITES
points(pRDA, display="sites", pch=21, scaling=3, cex=2, col="black",
       bg=cols[col_dframe$region]) # sites
text(pRDA, display="sites", scaling = 3, col="black", font=2, pos=4)
# PREDICTORS
text(pRDA, display="bp", scaling=3, col="red1", cex=1.4, lwd=2)
# SNPS
# text(pRDA, display="species", scaling = 3, col="blue", cex=0.7, pos=4) # SNPs
# LEGEND
legend("topright", legend=levels(col_dframe$region), bty="n", col="black",
       pch=21, cex=1.3, pt.bg=cols)
# OTHER LABELS
R2 = round(RsquareAdj(pRDA)$r.squared, 3)
mtext(bquote(italic("R")^"2"~"= "~.(R2)), side = 3, adj = 0.5)
dev.off()
```

# --------------#
#
# Candidate SNPs for local adaptation
#
# --------------#

```{r}
# Which axes are significant?
anova.cca(pRDA, permutations = 1000, by = "axis")
```

```{r}
# Extract SNP loadings for sig. axes
snp.load = scores(pRDA, choices = 1, display = "species")
```

```{r}
# Plot histograms of SNP loadings
hist(snp.load, main = "SNP loadings on RDA1")
```

```{r}
# Identify SNPs in the tails of the distribution
# Function from https://popgen.nescent.org/2018-03-27_RDA_GEA.html
outliers = function(x,z){
  lims = mean(x) + c(-1, 1) * z * sd(x) # find loadings +/-z sd from mean loading     
  x[x < lims[1] | x > lims[2]]          # locus names in these tails
}

# x = loadings vector, z = number of standard deviations to use
candidates = outliers(x = snp.load, z = 3)
```

```{r}
# Convert matric to dataframe
snp.load.df = snp.load %>% as.data.frame
snp.load.df$SNP_ID = rownames(snp.load.df)
str(snp.load.df)
```

```{r}
# Extract locus ID
snp.load.df %>% filter(RDA1 == candidates)
```